### 2015-10-07 Paper.tex
Import current state

\documentclass{llncs}
\usepackage{graphicx}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amssymb}
\usepackage{etoolbox}
\usepackage{color}
\usepackage{changepage}
\usepackage{csquotes}
\usepackage{tabularx,booktabs,multirow}
\usepackage{longtable}
\usepackage{framed}

\graphicspath{{Graphics/}}

\newcommand{\sym}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\svm}{SVM$^{rank}$}

\setlength{\tabcolsep}{6pt}

\begin{document}

\title{Learning Constraints and Optimization Criteria}
\author{Samuel Kolb}
\authorrunning{Samuel Kolb} % abbreviated author list (for running head)
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Samuel Kolb}
\institute{KULeuven, Leuven, Belgium\\samuel.kolb@me.com}

%\email{samuel.kolb@me.com}

\maketitle              % typeset the title of the contribution

\begin{abstract}
This paper presents a system that is able to automatically learn both constraints from examples and optimization criteria from rankings.
Both hard and soft constraints can be learned and constraints are represented as first order logical clauses.
Choosing such clauses as representations enables the use of ILP techniques such as clausal discovery.
In this research, optimization criteria are represented as weighted clauses.
For all examples, the clause learning system could find the essential constraints efficiently.
The learning system for optimization criteria is shown to achieve accurate results and can learn criteria that identify the correct optimal solution even for few examples and noisy rankings.
Finally, a procedure is outlined that allows weighted first order logic clauses to be used for optimization.
\end{abstract}

\section{Introduction}
% People use computers to solve complex problems, which are hard to solve by hand.
% Traditionally, programs or algorithms are written to tell the computer \emph{how} to solve a particular problem.
% These programs are essentially precise, step-by-step instructions.
% Alternatively, problems can be described \emph{declaratively} by formulating them in a high level language.
% Such a language describes a problem without specifying how a solution is to be found.
% Generic programs, called solvers, are used to find solutions to problems formulated in this language.

% A popular movement within declarative programming is constraint programming.
% In constraint programming users specify variables, their domains and constraints that act upon these variables.
% The constraint solver assigns values to the variables such that all the constraints hold.
% This solver has a large degree of freedom in choosing a search strategy and ideally the user does not need to be familiar with the underlying algorithms.

% Aside from being popular in academia, it is also being used in industrial applications\cite{Simonis:IndustrialApplicationsCP}.
% Much of the research on constraint programming is focused on improving solvers.
% In practice, however, it can be challenging for non-experts to formulate constraints to describe their problem\cite{Wallace:PrinciplesCP}.

This research tackles the learning of constraints and optimization criteria.
Non-experts often experience difficulties when modeling a problem using constraints\cite{Wallace:PrinciplesCP}.
The first goal of this research is to automate the step of modeling a problem, in order to, amongst others, make constraint solving more accessible.
In this approach a few examples of what a user considers valid solutions are used by a computer system to find constraints.
These constraints can then be used by an existing constraint solver.

Constraints are represented as logical clauses.
This representation allows for the use of techniques from the field of Inductive Logic Programming (ILP).
Specifically, a new implementation of the Clausal Discovery algorithm\cite{DeRaedt:ClausalDiscovery} is used.
Research on constraint learning is often done independent of ILP.
By exploiting the parallels between both fields, this research follows the same approach as Lallouet et al.\cite{Lallouet:LearningCP}.

Hard constraints are used to distinguish solutions to a problem.
So called soft constraints can be used to express properties, which do not have to be fulfilled by every solution.
Weights can be assigned to such soft constraints to express their importance.
Such weighted soft constraints can be used as optimization criteria.
Optimal solutions are found by maximizing the sum of the weights of soft constraints that are fulfilled by the solution.

The second goal of this research is to learn from user preference and to formulate corresponding optimization criteria.
By using rankings over a few examples, weighted soft constraints are found.
Both the learning of constraints as well as the learning of optimization criteria aim to bridge the gap between problems and their formal representation.

The use of weighted soft constraints is inspired by the research of Campigotto et al.\cite{campigotto2011active}, amongst others.
Their work examines attempts to obtain similar results in an interactive context for propositional constraints.
In this research, first-order logical clauses will be used.

The weighted clauses cannot be directly used in a constraint solver.
Therefore, an approach is outlined that describes how these constraints can be used in an existing solver to obtain an optimal solution in practice.

This research has been conducted during the course of my masters thesis~\cite{kolb2015thesis}.
\section{Background}
In this section a brief overview of the important concepts and relevant literature will be provided

\subsection{Clauses}
First-order logical clauses consist of a disjunction of literals.
Literals are defined as being either an atom or a negated atom.
Clauses are often expressed using a body and a head: $\mathit{head} \leftarrow \mathit{body}$.
All literals containing a negation are grouped in the body and the other literals form the head of the clause.
For example, $\lnot a_1 \lor \lnot a_2 \lor a_3 \lor a_4$ corresponds with $a_3, a_4 \leftarrow a_1, a_2$.

Atoms are logical concepts that can be assigned a truth value.
In this paper, atoms will be restricted to either the symbols $\mathit{true}$ or $\mathit{false}$ or a predicate.
Predicates consist of a predicate symbol and a set of terms.
Terms are either constants or variables and describe objects in a domain.
Predicates express relations between these objects: $\mathit{symbol}(\mathit{term_1}, ... \mathit{term_n})$.
Clauses are universally quantified and will, in this research, only contain variables.

\subsection{Learning constraints}
Constraint learning is a hard problem.
The search space of constraints is usually very large.
Furthermore, there are usually only a few examples to learn from, unlike in typical machine learning problems.

Different systems attempt to learn constraints using different methods.
The problem of having little examples can be alleviated by interactively generating data.
Both the systems, Conacq2\cite{bessiere2007query} en QuAcq\cite{bessiere2013constraint}, generate (partial) examples and query the user about their validity.

ModelSeeker\cite{Beldiceanu:ModelSeeker} attempts to structure the problem variables in different ways.
The system then tests for different global constraints if they hold within this structure.
This approach allows for constraint learning from a few examples.

As mentioned earlier, the research of Lallouet et al.\cite{Lallouet:LearningCP} already tried to use ILP techniques to learn constraints.
In their approach, the search space of constraints is explored in a bidirectional manner, primarily using negative examples.

\subsection{Clausal Discovery}
The clausal discovery algorithm proposed by De Raedt et al.\cite{DeRaedt:ClausalDiscovery} attempts to learn first order logic clauses from positive examples.
By using a refinement operator, the search space is traversed by incrementally generalizing clauses until they cover all examples, starting from the empty clause~($\square$).
The result set consists of the most specific clauses covering all examples.
This research uses the clausal discovery algorithm on a high level but provides a very different implementation than the original clausal discovery engine.

\subsection{Logic Solver}
The logic system IDP\cite{de2013prototype,wittocx2008idp} is used to calculate whether examples satisfy clauses and is used as a theorem prover.
It supports an extension of first order logic ($FO(\cdot)$) and can also be used to generate new solutions.
Background knowledge is given as an IDP theory.

\subsection{Optimization}
The Boolean Satisfiability Problem (SAT) attempts to determine whether a propositional formula can be satisfied.
Formulas can be rewritten into a special form called the conjunctive normal form (CNF), which consists of a conjunction of clauses (disjunctions).
An extension of this problem is the Maximum Satisfiability Problem (MAX-SAT).
In MAX-SAT the amount of satisfied clauses is maximized.
By assigning positive integer weights to clauses and maximizing the sum of the weights of satisfied clauses, one obtains the weighted MAX-SAT problem.
The work by Campigotto et al.\cite{campigotto2011active} attempts to learn propositional formulas en weights automatically.
Inspired by their results, this research tries to learn first-order logic optimization criteria.
The resulting optimization problem can be seen as an extension of weighted MAX-SAT in which the clauses are in first-order logic.

\section{Problem Statement}
This research has two main objectives: learning constraints and learning optimization criteria.
The system designed to solve these problems attempts to simplify the process of generating formal representations.
Therefore, it should be easy to use and take into account what kind of information a user can provide.

\begin{framed}
  \noindent
  \begin{minipage}{\textwidth}
    \paragraph*{Goal 1}
    Given a set of examples and a limit $t$, the goal of the constraint learning system is to find maximally specific clauses that are satisfied by at least $t$ of the given examples.
  \end{minipage}
\end{framed}

Examples describe specific solutions.
They contain a set of objects and exhaustively describe all relations (predicates) that hold on these objects.
The learned clauses are independent of the specific objects within the examples and contain only variables.

\begin{example}
  Consider sudoku as an example of a problem, where hard constraints are to be found.
  Given one or more solved sudokus, the learning system has to identify the rules of the game.
  Since the learned clauses only contain universally quantified variables, for example $4 \times 4$ sudokus can be used to find constraints.
  The learned constraints are independent of the size of the sudoku and can also be used to solve $9 \times 9$ sudokus, for example.

  In some problems, one is interested in regularities that  often but not always occur.
  An example would be regularities in certain types of buildings.
  While certain regularities would occur in many buildings, there are often exceptions and different styles.
  In this case, one could attempt to find constraints on the structure of the floor plans that are satisfied by a given amount of examples.
\end{example}

\begin{framed}
  \noindent
  \begin{minipage}{\textwidth}
    \paragraph*{Goal 2}
    Given a set of examples and a set of partial rankings over these examples, the goal of the optimization system is to learn soft constraints and weights such that the order described by these optimization criteria maximally corresponds with the given rankings.
  \end{minipage}
\end{framed}

The soft constraints consist of clauses that are satisfied by some of the examples.
Rankings are partial orders over the examples: $e_x = e_y > ... > e_a = e_b$.
The weighted clauses can be used to calculate a score for every example
\begin{eqnarray*}
  \sum\limits_{\mathit{weight}, \mathit{c}} \mathit{weight} \cdot v_{c}
\end{eqnarray*}
With $v_c = 1$ if the clause $c$ is satisfied by the example and $v_c = 0$ if the $c$ is not satisfied.
These scores allow for a total order over all examples to be defined.

\begin{example}
  Consider the scenario of moving to a new city.
  Multiple job offers are available, different accommodations are to be considered and there are different schools in the city.
  Learning weighted soft constraints can make user preferences explicit.
  For example, it could be desirable that the school is in a low crime area or that the locations of the school and work place are close.
\end{example}

A third goal of this research is to show how these constraints and especially the optimization criteria can be used in practice to find an optimal solution.

\section{Approach}
\begin{figure}

  \centering
    \includegraphics[width=0.8\linewidth]{ApproachOverview.pdf}
  \caption{Overview approach}
  \label{fig:struktuur}

\end{figure}

In this research, both constrain learning and learning optimization criteria were implemented.
An overview is presented in figure~\ref{fig:struktuur}.
Clauses are learned using examples.
To find weighted clauses, first the constraint learning system is used to find soft constraints.
Then the weights for these soft constraints are determined based on the rankings.
In order to determine the weights, the \svm system is used.

\subsection{Input}
The input consists of examples and definitions.
Every example contains a set of constants that are part of the model and an exhaustive list of the relations that hold on these constants.
Definitions describe either typing information or predicate definitions.
Typing information relates directly to the problem domain, improves the accuracy and efficiency of the learned clauses, and is usually easy to provide for the user.
Predicate definitions describe the types of the terms occurring in a predicate.
Two special types of predicates are allowed: calculated predicates and symmetric predicates.
Calculated predicates are not listed in examples, instead background knowledge is used to generate them.
Symmetric predicates are predicates for which the order of the terms does not matter.
If optimization criteria are to be learned, a set of rankings over the given examples is also provided.

\begin{example}
  Imagine problem that describes humans with a type \textit{human} and predicates human(\textit{human}), male(\textit{human}) and female(\textit{human}).
  An example could contain the constants S and A, both of type \textit{human}, and the relations: human(S), human(A), male(S), female(A).
  Clauses to be learned would be: human(x) $\leftarrow$ male(x) and human(x) $\leftarrow$ female(x),  female(x) $\lor$ male(x) $\leftarrow$ human(x) and \textit{false} $\leftarrow$ female(x) $\land$ male(x).
\end{example}
\subsection{Clausal Discovery}
The clause learning system is based on the clausal discovery algorithm.
The logical system IDP is used for two essential functions: $\mathit{covers}$ and $\mathit{entails}$.
The $\mathit{covers}$ function calculates which examples are covered by a clause.
The $\mathit{entails}$ function calculates whether an accepted clause is logically entailed by background knowledge or clauses already in the result set.

Every clause is subjected to two additional tests that attempt to eliminate redundant clauses early on.
The subset-test determines whether the clause is a superset of another clause that has been accepted earlier.
If such a clause exists and that clause covers the same examples as the new clause, the new clause is discarded.
Otherwise it will be tested whether the clause covers enough examples.
Clauses that do not cover enough examples are refined by the refinement operator.
The refined clauses will be added to the working queue.
If a clause offers enough examples, the $\mathit{entails}$ functions is calculated to decide whether the accepted clause should be added to the result set.

Refinement is an important step.
Clauses are extended to be more general and potentially cover more examples.
The refinement operator uses a list of atoms that can potentially be added to a clause.
This list is calculated in advance, given a maximal amount of variables.
Within the body and head of a clause, atoms may only be added in a specific order.
Additionally, several restrictions are in place.
Clauses must be connected, i.e. new atoms must always contain a variable that has already occurred.
Secondly, clauses must be range-restricted, which means that no new variables may be introduced in the head of the clause.
By looking at the positions of the atoms in the list, clauses can be reduced to a unique form.
This reduces the amount of duplicate clauses generated.

\subsection{Optimization}
The first step in finding weighted soft constraints is to identify soft constraints by using the constraint learning system and a low threshold.
Examples are characterized by the soft constraints (clauses) which it satisfies.
Therefore, every example can be translated to a vector of boolean features.
Each feature $f_i$ corresponds with one of the clauses and is \emph{1} if the clause covers the example and \emph{0} if it does not.
Existing machine learning software can be used to learn a linear \emph{scoring} function $\sum_i w_i \cdot f_i$ over these features based on the given rankings.
Hereby, positive weights are assigned to every feature to express its importance.
For an unseen example, the feature vector is computed by calculating what clauses cover the new example.
The scoring function can then be used to calculate a score for that example.

\begin{example}
  Consider examples~$e_1, e_2, e_3$, rankings $e_1 > e_2, e_2 > e_3$ and clauses $c_1, c_2, c_3$.
  If the clauses cover examples according to table~\ref{tbl:cover_examples}, then the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.
  An example $e_4$ that is covered by~$c_1$ and~$c_2$, is then assigned a score of $1$.
  This score has no value in the absolute sense, it can only be used to compare it to other examples ranked by the same function.
  In this example clause~$c_1$ represents a desirable property, clause~$c_2$ is ignored because it does not influence the ranking and clause~$c_3$ represents an undesirable property.

  \begin{table}[!htp]
  \caption{Clause coverage}
  \label{tbl:cover_examples}
  \begin{tabularx}{\textwidth}{c|ccc|X}
    \textbf{Clauses}  &$c_1$    & $c_2$   & $c_3$   & \textbf{Feature vector}\\
    \toprule
    $e_1$         & Covers  &       &       & 1, 0, 0\\
    $e_2$         &       & Covers  &       & 0, 1, 0\\
    $e_3$         & Covers  & Covers  & Covers  & 1, 1, 1\\
  \end{tabularx}
  \end{table}

  Considering an extended form of the human example, the optimization function could be:
  1 $\cdot$ (funny(x)~$\leftarrow$~human(x)) + (-2) $\cdot$ (lazy(x)~$\leftarrow$~human(x)), for example.
  Clauses with a weight of zero are ignored.
\end{example}

The \svm{} software was chosen to find the scoring function since it is a robust and efficient system.
Contrary to many other ranking systems, it uses a linear model.
The weights that it assigns to the features can be used directly as weights for the optimization criteria.
The input format that is used by \svm{} is also supported by many other learn-to-rank implementations.
Therefore, other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be plugged into the implementation if so desired.

\subsection{Optimal Solution}
Solvers, such as IDP, can use clauses directly as hard constraints to generate a solution.
There is no solver that can use the chosen optimization criteria directly.
Weighted MAX-SAT solvers use propositional clauses and only allow for positive weights.

This optimization task can be solved directly in IDP, using inductive definitions, aggregates and minimization.
The only limitation is that the weights must be integer values.
Therefore, the weights of the clauses are divided by the smallest absolute weight, multiplied by a constant and rounded to the closest integer if necessary.
Removing this restriction for IDP is currently being researched and future releases might be able to deal with rational numbers directly.

In order to model the optimization problem in IDP, every clause $c_i$ with variables $v_1, ..., v_n$ is represented by a number $i$. For every clause a predicate $t(i)$ is added to capture the truth value of the clause.
A function $\mathit{cost}(i)$ specifies the cost of not satisfying the clause, which is equal to the integer weight of the clause.
\begin{eqnarray*}
  t(i) \Leftrightarrow \forall v_1, ..., v_n : c_i. \\
  cost(i) = w_i.
\end{eqnarray*}

Using $t$ and $\mathit{cost}$, a function $\mathit{actual}(i)$ is then defined in IDP as 0 if $t(i)$ and $\mathit{cost}(i)$ if $\lnot t(i)$.
This function is used in the term $\sum_i actual(i)$ to be minimized, which will allow IDP to search for an optimal solution.

\section{Evaluation}
Several experiments aim to measure the accuracy and efficiency of the learning systems for constraints and optimization criteria.

\subsection{Constraints}
Four problems have been used to evaluate constraint learning.
Two are existing problems: map coloring and sudoku.
Map coloring is a problem where countries are assigned colors and neighboring countries may not have the same color.
The examples to learn from each consist of a set of three countries that are correctly colored.
For sudoku a $4 \times 4$ solved sudoku is given.
Furthermore, there is the elevator problem with a soft constraint that holds on 2 of the 3 given examples.
The last problem is the co-housing problem with 4 hard constraints and 5 given examples.

\paragraph{Accuracy}
The essential constraints were found for all problems.
Additionally, constraints were found that describe some structure in the problems, such as that countries are never their own neighbor.
These kind of constraints may help a constraint solver to work more efficiently.

If the limits on the number of variables or literals per clause are too large, constraints are found that over-fit the training data.
These constraints are too specific and will exclude valid solutions.
On the other hand, if the limits chosen are too small, the necessary constraints will not be found.

\begin{table}
  \caption{Overview of execution times (average over 8 runs)}
  \begin{tabularx}{\linewidth}{rl|ll}

\textbf{Omitted}  & \textbf{Problem}    & \textbf{Time (baseline)}  & \textbf{Average time (s)}\\
\toprule
Nothing          & Map coloring  & 1.000       & 1.581   ($\pm$ 0.117)        \\
   (baseline)                   & Sudoku    & 1.000               & 4.787   ($\pm$ 0.062)    \\
                      & Elevator    & 1.000             & 3.182   ($\pm$ 0.073)  \\
                      & Co-housing  & 1.000             & 25.903  ($\pm$ 0.446)  \\
\midrule
Range      & Map coloring  & 2.928           & 4.629   ($\pm$ 0.199)\\
      restriction                  & Sudoku    & 3.367             & 16.118  ($\pm$ 0.154)  \\
                        & Elevator    & 12.713          & 40.453  ($\pm$ 0.319)\\
                        & Co-housing  & 8.021           & 207.768 ($\pm$ 0.330)\\
\midrule
Connected    & Map coloring  & 1.005           & 1.589   ($\pm$ 0.110)\\
    clauses       & Sudoku    & 1.476                          & 7.068   ($\pm$ 0.150)    \\
                        & Elevator    & 1.935           & 6.157   ($\pm$ 0.114)\\
                        & Co-housing  & 4.001           & 103.633 ($\pm$ 0.131)\\
  \end{tabularx}
  \label{tbl:uitvoering}
\end{table}

\paragraph{Efficiency}
Table~\ref{tbl:uitvoering} shows the execution time for several experiments demonstrating the effect of the syntactical restrictions.
Moreover, all efficiency measures (i.e. symmetric predicates, range restriction and connected clauses) are able to improve the execution time.

Further experiments show that increasing the number of variables or literals per clause impacts the efficiency.
Especially the combination of more variables \emph{and} literals can steeply increase the execution time.
Therefore it would be useful to adapt these parameters dynamically.
Adding additional examples only increases the execution time by a constant factor.
Since adding more examples can help improve the accuracy, this trade-off is often worthwhile.

\paragraph{Compared to humans}
Human programmed theories for map coloring and sudoku are available on the website of the IDP system.
These theories usually focus on being compact and contain only the essential constraints.
Table~\ref{tbl:mens} shows the results of two experiments that measure the time to find a solution for a new problem.
This time is measured for the learned theories as well as for hand made theories.
The learned theories are slightly adapted to be able to solve the same problems and corrected if they contains an unfair advantage.
  \begin{table}
    \caption{CPU times human vs. learned theory}
    \begin{tabularx}{\linewidth}{lr|X}
      \textbf{Problem} & \textbf{Type} & \textbf{Mean CPU time (s)} \\
      \toprule
      Map coloring & Human & $0.968$  ($\pm 0.023$) \\
      & Learned & $0.403$       ($\pm 0.015$) \\
      \midrule
      Sudoku & Human & $1.453$    ($\pm 0.018$) \\ 
      & Learned & $0.310$       ($\pm 0.012$)
    \end{tabularx}
    \label{tbl:mens}
  \end{table}

Especially for non-experts, a learning system can be useful to assist them during the modeling process.
Additionally, the learning system can function in an automatic setting.
These experiments show that learned constraints can be used to solve problems efficiently and even faster than hand programmed constraints for the examined cases.

\subsection{Optimization}
The efficiency of the learning system for optimization criteria depends mainly on the efficiency of learning soft constraints.
Therefore, the experiments in this section are focused on the accuracy of the optimization criteria and the influence of different factors on the accuracy.

The previously mentioned problem of selecting a place to live, a workplace and a school is used as an example.
In the chosen scenario, there are 18 possible options which form the available examples.
Two approaches are used for evaluation.
In the first approach, the examples are split into disjoint train and test sets.
The second approach uses all examples as train as well as test set.
An underlying model of four weighted soft constraints is used to calculate a score for each example.
All possible pairs of examples are generated and translated to pairwise inequalities using the calculated scores.
These rankings are given as input for learning.
Experiments use different sizes of training sets and different fractions of rankings.
Noise is simulated by flipping a fraction of the selected rankings.

The experiments are run 8 times, and the examples and rankings to be used as well as the rankings to be inverted are selected randomly.
Optimization criteria are searched based on the training set and the selected inequalities.
These criteria are used to predict the better example out of every pair of examples.
Experiments are scored by comparing the amount of correctly predicted pairs (1.0 indicating full agreement).

\begin{figure}
  \begin{minipage}{0.5\textwidth}

  \centering
    \includegraphics[width=1.1\linewidth]{rankings}
  \caption{Influence fraction of inequalities}
  \label{fig:fractie}

  \end{minipage}
  \begin{minipage}{0.5\textwidth}

      \centering
        \includegraphics[width=1.1\linewidth]{noise}
      \caption{Influence of noise}
      \label{fig:ruis}

  \end{minipage}
    \end{figure}

Figure~\ref{fig:fractie} shows how the scores improve as the amount of examples and the fraction of included inequalities increases.
In all cases, more than half the pairs of examples are correctly predicted and high scores can be obtained even for small datasets.
Other experiments have shown that even for lower scores the optimization criteria are often capable of finding an optimal solution.

The underlying model can be directly expressed using the soft constraints that can be learned.
However, similar scores are obtained for models which cannot be directly modeled (e.g. the use of weighted soft constraints which are disconnected).
While there are limits to the expressibility, the learned criteria are shown to be robust with respect to the exact formulation.

The influence of noise is shown in figure~\ref{fig:ruis}.
Hereby, the second approach of testing is used, where all the examples are used as test set.
The algorithm obtains high scores despite significant levels of noise.
Figure~\ref{fig:ruis} also shows that providing more rankings improves the robustness of the algorithm, even if the relative amount of noise remains unchanged.
Using the first approach with a training set of 40\%, scores over 0.7 can be obtained for noise levels of 20\%.

  \begin{table}
    \caption{Scores for different thresholds ($t$)}
    \begin{tabularx}{\linewidth}{XXXX}
      $t = 1$ & $t = 2$ & $t = 3$ & $t = 4$ \\
      \toprule
     0.823 & 0.740 & 0.788 & 0.735 \\
     ($\pm$ 0.073)&
($\pm$ 0.078)&
($\pm$ 0.074)&
($\pm$ 0.063)
    \end{tabularx}
    \label{tbl:limiet}
  \end{table}

Finally, table~\ref{tbl:limiet} shows that increasing the threshold used for finding soft constraints does not improve the score.
This experiment used a 40\% training set and 40\% of the available rankings.
However, if the size of the problem and examples is increased, higher threshold will likely be appropriate.

\section{Conclusion}
The research in this thesis has focused on solving two problems: 1) Learning constraints; and 2) Learning optimization criteria.
Implementation for both clause learning and clausal optimization have been provided that accomplish these tasks using first order logic clauses.

The constraint learning system has been able to learn the relevant hard and soft constraints in all experiments.
As such, the first goal of this research has been accomplished.
All problems have been solved in limited time (less than a minute) and often even within a few seconds.
For each problem, only a small number of examples was given to learn from.
The system requires only a minimal amount of information from the user.
However, the system also allows for the use of expressive background knowledge.
The learned constraints are domain independent, which facilitates the construction of positive examples.

The second goal of this research, learning optimization criteria, has been accomplished as well.
Optimization criteria are discovered that enable an optimal solution to be found.
Even for small datasets and noise on the rankings, constraints are found that enable most examples to be ranked correctly.

Aside from learning formal representations automatically from examples, this research shows how these representations can be used in practice.
This also forms an important step to enable the learning of optimization criteria in an interactive setting.

\paragraph{Future work}
This research offers multiple opportunities for future work.
It would be interesting to adapt the number of variables and literals in a clause dynamically.
This could be accomplished, for example, by using negative examples.
Learned clauses must be specific enough to not cover any negative examples.

Additionally, it would be interesting to add interactivity to the learning system for the generation of examples or rankings.
Rankings expressing that examples are equally ranked are currently ignored.
However, it could be interesting to incorporate this information into the algorithm, whenever it is provided explicitly.

As mentioned earlier, the domain independence of the clauses has several advantages.
In some cases, however, specific objects are inherently present in the problem.
The constraint learning system could be enhanced to include such global constants.

Finally, it would desirable to improve the implementation in order to tackle larger sized problems.
All the experiments were conducted with problems of limited size and the computation time increases rapidly if there are more predicates, variables and literals to be used in clauses.

\section*{Acknowledgments}
The author thanks his promoters Dr. Luc De Raedt en Dr. ir. Anton Dries.
Furthermore, he expresses his gratitude for the help provided by Bart Bogaerts, Dr. Jesse Davis, Dr. Marc Denecker and Vladimir Dzyuba.

%
% ---- Bibliography ----
%
\bibliographystyle{plain}
\bibliography{Bibliography}
\end{document}

### 2015-10-19 Paper.tex
Transition to AAAI style

@@
-\documentclass{llncs}
+\def\year{2015}
+\documentclass[letterpaper]{article}
+\usepackage{aaai}
 \usepackage{graphicx}
-
 \usepackage{algorithm}
 \usepackage{algpseudocode}
+\usepackage{amsmath}
 \usepackage{amssymb}
+\usepackage{amsthm}
 \usepackage{etoolbox}
@@
 
+\pdfinfo{
+/Title (Learning Constraints and Optimization Criteria)
+/Author (Samuel Kolb)}
+\setcounter{secnumdepth}{0}
+
 \graphicspath{{Graphics/}}
@@
 \newcommand{\svm}{SVM$^{rank}$}
+
+\newtheorem{definition}{Definition}[section]
+\newtheorem{question}{Question}
+\AfterEndEnvironment{definition}{\noindent\ignorespaces}
+
+\theoremstyle{definition}
+\newtheorem{example}{Example}
+\newtheorem{experiment}{Experiment}[question]
+\newtheorem{observation}{Observation}[question]
 
@@
 \title{Learning Constraints and Optimization Criteria}
-\author{Samuel Kolb}
-\authorrunning{Samuel Kolb} % abbreviated author list (for running head)
-%%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Samuel Kolb}
-\institute{KULeuven, Leuven, Belgium\\samuel.kolb@me.com}
+\author{
+  Samuel Kolb\\
+  KULeuven, Leuven, Belgium\\
+  samuel.kolb@cs.kuleuven.be
+}
 
-%\email{samuel.kolb@me.com}
-
-\maketitle              % typeset the title of the contribution
+\maketitle
 
@@
 \section{Introduction}
-% People use computers to solve complex problems, which are hard to solve by hand.
-% Traditionally, programs or algorithms are written to tell the computer \emph{how} to solve a particular problem.
-% These programs are essentially precise, step-by-step instructions.
-% Alternatively, problems can be described \emph{declaratively} by formulating them in a high level language.
-% Such a language describes a problem without specifying how a solution is to be found.
-% Generic programs, called solvers, are used to find solutions to problems formulated in this language.
-
-% A popular movement within declarative programming is constraint programming.
-% In constraint programming users specify variables, their domains and constraints that act upon these variables.
-% The constraint solver assigns values to the variables such that all the constraints hold.
-% This solver has a large degree of freedom in choosing a search strategy and ideally the user does not need to be familiar with the underlying algorithms.
-
-% Aside from being popular in academia, it is also being used in industrial applications\cite{Simonis:IndustrialApplicationsCP}.
-% Much of the research on constraint programming is focused on improving solvers.
-% In practice, however, it can be challenging for non-experts to formulate constraints to describe their problem\cite{Wallace:PrinciplesCP}.
-
 This research tackles the learning of constraints and optimization criteria.
-Non-experts often experience difficulties when modeling a problem using constraints\cite{Wallace:PrinciplesCP}.
+Non-experts often experience difficulties when modeling a problem using constraints \cite{Wallace:PrinciplesCP}.
 The first goal of this research is to automate the step of modeling a problem, in order to, amongst others, make constraint solving more accessible.
@@
 This representation allows for the use of techniques from the field of Inductive Logic Programming (ILP).
-Specifically, a new implementation of the Clausal Discovery algorithm\cite{DeRaedt:ClausalDiscovery} is used.
+Specifically, a new implementation of the Clausal Discovery algorithm \cite{DeRaedt:ClausalDiscovery} is used.
 Research on constraint learning is often done independent of ILP.
-By exploiting the parallels between both fields, this research follows the same approach as Lallouet et al.\cite{Lallouet:LearningCP}.
+By exploiting the parallels between both fields, this research follows the same approach as \cite{Lallouet:LearningCP}.
 
 Hard constraints are used to distinguish solutions to a problem.
-So called soft constraints can be used to express properties, which do not have to be fulfilled by every solution.
-Weights can be assigned to such soft constraints to express their importance.
+So called soft constraints can be used to express properties which do not have to be fulfilled by every solution.
+Soft constraints are assigned weights that express their importance.
 Such weighted soft constraints can be used as optimization criteria.
-Optimal solutions are found by maximizing the sum of the weights of soft constraints that are fulfilled by the solution.
+Optimal solutions are found by maximizing the sum of the weights of soft constraints that are fulfilled by a solution.
 
 The second goal of this research is to learn from user preference and to formulate corresponding optimization criteria.
-By using rankings over a few examples, weighted soft constraints are found.
+By using rankings over a few examples, weighted soft constraints can be found.
 Both the learning of constraints as well as the learning of optimization criteria aim to bridge the gap between problems and their formal representation.
 
-The use of weighted soft constraints is inspired by the research of Campigotto et al.\cite{campigotto2011active}, amongst others.
+The use of weighted soft constraints is inspired by the research of \cite{campigotto2011active}, amongst others.
 Their work examines attempts to obtain similar results in an interactive context for propositional constraints.
@@
 
-The weighted clauses cannot be directly used in a constraint solver.
+The weighted clauses cannot be directly used by a constraint solver.
 Therefore, an approach is outlined that describes how these constraints can be used in an existing solver to obtain an optimal solution in practice.
@@
 \section{Background}
-In this section a brief overview of the important concepts and relevant literature will be provided
+In this section a brief overview of the important concepts and relevant literature will be provided.
 
@@
 All literals containing a negation are grouped in the body and the other literals form the head of the clause.
-For example, $\lnot a_1 \lor \lnot a_2 \lor a_3 \lor a_4$ corresponds with $a_3, a_4 \leftarrow a_1, a_2$.
+For example, $\lnot a_1 \lor \lnot a_2 \lor a_3 \lor a_4$ corresponds to the clause $a_3, a_4 \leftarrow a_1, a_2$.
 
@@
 Constraint learning is a hard problem.
-The search space of constraints is usually very large.
+The search space of constraints is typically very large.
 Furthermore, there are usually only a few examples to learn from, unlike in typical machine learning problems.
@@
 The problem of having little examples can be alleviated by interactively generating data.
-Both the systems, Conacq2\cite{bessiere2007query} en QuAcq\cite{bessiere2013constraint}, generate (partial) examples and query the user about their validity.
+Both the systems Conacq2 \cite{bessiere2007query} en QuAcq \cite{bessiere2013constraint} generate (partial) examples and query the user about their validity.
 
-ModelSeeker\cite{Beldiceanu:ModelSeeker} attempts to structure the problem variables in different ways.
-The system then tests for different global constraints if they hold within this structure.
+ModelSeeker \cite{Beldiceanu:ModelSeeker} attempts to structure the problem variables in different ways.
+The system then tests for different global constraints whether they hold within this structure.
 This approach allows for constraint learning from a few examples.
 
-As mentioned earlier, the research of Lallouet et al.\cite{Lallouet:LearningCP} already tried to use ILP techniques to learn constraints.
+As mentioned earlier, the research of \cite{Lallouet:LearningCP} already tried to use ILP techniques to learn constraints.
 In their approach, the search space of constraints is explored in a bidirectional manner, primarily using negative examples.
@@
 \subsection{Clausal Discovery}
-The clausal discovery algorithm proposed by De Raedt et al.\cite{DeRaedt:ClausalDiscovery} attempts to learn first order logic clauses from positive examples.
-By using a refinement operator, the search space is traversed by incrementally generalizing clauses until they cover all examples, starting from the empty clause~($\square$).
+The clausal discovery algorithm proposed by \cite{DeRaedt:ClausalDiscovery} attempts to learn first order logic clauses from positive examples.
+Using a refinement operator, the search space is traversed by incrementally generalizing clauses until they cover all examples, starting from the empty clause~($\square$).
 The result set consists of the most specific clauses covering all examples.
@@
 \subsection{Logic Solver}
-The logic system IDP\cite{de2013prototype,wittocx2008idp} is used to calculate whether examples satisfy clauses and is used as a theorem prover.
+The logic system IDP \cite{de2013prototype,wittocx2008idp} is used to calculate whether examples satisfy clauses and is used as a theorem prover.
 It supports an extension of first order logic ($FO(\cdot)$) and can also be used to generate new solutions.
@@
 By assigning positive integer weights to clauses and maximizing the sum of the weights of satisfied clauses, one obtains the weighted MAX-SAT problem.
-The work by Campigotto et al.\cite{campigotto2011active} attempts to learn propositional formulas en weights automatically.
+The work by \cite{campigotto2011active} attempts to learn propositional formulas en weights automatically.
 Inspired by their results, this research tries to learn first-order logic optimization criteria.
@@
 \begin{example}
-  Consider sudoku as an example of a problem, where hard constraints are to be found.
+  Consider sudoku as an example of a problem where hard constraints are to be found.
   Given one or more solved sudokus, the learning system has to identify the rules of the game.
-  Since the learned clauses only contain universally quantified variables, for example $4 \times 4$ sudokus can be used to find constraints.
+  Since the learned clauses only contain universally quantified variables, for example $4 \times 4$ sudokus could be used to find constraints.
   The learned constraints are independent of the size of the sudoku and can also be used to solve $9 \times 9$ sudokus, for example.
@@
 \end{eqnarray*}
-With $v_c = 1$ if the clause $c$ is satisfied by the example and $v_c = 0$ if the $c$ is not satisfied.
+With $v_c = 1$ if the clause $c$ is satisfied by the example and $v_c = 0$ if $c$ is not satisfied.
 These scores allow for a total order over all examples to be defined.
@@
 
-In this research, both constrain learning and learning optimization criteria were implemented.
+In this research, both constraint learning and learning optimization criteria were implemented.
 An overview is presented in figure~\ref{fig:struktuur}.
@@
 \begin{example}
-  Imagine problem that describes humans with a type \textit{human} and predicates human(\textit{human}), male(\textit{human}) and female(\textit{human}).
+  Imagine a problem that describes humans with a type \textit{human} and predicates human(\textit{human}), male(\textit{human}) and female(\textit{human}).
   An example could contain the constants S and A, both of type \textit{human}, and the relations: human(S), human(A), male(S), female(A).
-  Clauses to be learned would be: human(x) $\leftarrow$ male(x) and human(x) $\leftarrow$ female(x),  female(x) $\lor$ male(x) $\leftarrow$ human(x) and \textit{false} $\leftarrow$ female(x) $\land$ male(x).
+  Clauses to be learned would be: human(x) $\leftarrow$ male(x), human(x) $\leftarrow$ female(x),  female(x) $\lor$ male(x) $\leftarrow$ human(x) and \textit{false} $\leftarrow$ female(x) $\land$ male(x).
 \end{example}
@@
 The $\mathit{covers}$ function calculates which examples are covered by a clause.
-The $\mathit{entails}$ function calculates whether an accepted clause is logically entailed by background knowledge or clauses already in the result set.
+The $\mathit{entails}$ function calculates whether an accepted clause is logically entailed by background knowledge or clauses that are already in the result set.
 
@@
 The refined clauses will be added to the working queue.
-If a clause offers enough examples, the $\mathit{entails}$ functions is calculated to decide whether the accepted clause should be added to the result set.
+If a clause covers enough examples, the $\mathit{entails}$ functions is calculated to decide whether the accepted clause should be added to the result set.
 
@@
 Existing machine learning software can be used to learn a linear \emph{scoring} function $\sum_i w_i \cdot f_i$ over these features based on the given rankings.
-Hereby, positive weights are assigned to every feature to express its importance.
+Hereby, weights are assigned to every feature to express its importance.
 For an unseen example, the feature vector is computed by calculating what clauses cover the new example.
@@
 
-This optimization task can be solved directly in IDP, using inductive definitions, aggregates and minimization.
+This optimization task can be solved in IDP, using inductive definitions, aggregates and minimization.
 The only limitation is that the weights must be integer values.
@@
 These constraints are too specific and will exclude valid solutions.
-On the other hand, if the limits chosen are too small, the necessary constraints will not be found.
+On the other hand, if the chosen limits are too small, the necessary constraints will not be found.
 
 \begin{table}
-  \caption{Overview of execution times (average over 8 runs)}
+  \caption{Execution times overview (average over 8 runs)}
   \begin{tabularx}{\linewidth}{rl|ll}
 
-\textbf{Omitted}  & \textbf{Problem}    & \textbf{Time (baseline)}  & \textbf{Average time (s)}\\
+\textbf{Omitted}  & \textbf{Problem}    & \textbf{Average time (s)} \\ % & \textbf{Time (baseline)}  
 \toprule
-Nothing          & Map coloring  & 1.000       & 1.581   ($\pm$ 0.117)        \\
-   (baseline)                   & Sudoku    & 1.000               & 4.787   ($\pm$ 0.062)    \\
-                      & Elevator    & 1.000             & 3.182   ($\pm$ 0.073)  \\
-                      & Co-housing  & 1.000             & 25.903  ($\pm$ 0.446)  \\
+Nothing           & Map coloring        & 1.581   ($\pm$ 0.117)     \\ % & 1.000                     
+   (baseline)     & Sudoku              & 4.787   ($\pm$ 0.062)     \\ % & 1.000                     
+                  & Elevator            & 3.182   ($\pm$ 0.073)     \\ % & 1.000                     
+                  & Co-housing          & 25.903  ($\pm$ 0.446)     \\ % & 1.000                     
 \midrule
-Range      & Map coloring  & 2.928           & 4.629   ($\pm$ 0.199)\\
-      restriction                  & Sudoku    & 3.367             & 16.118  ($\pm$ 0.154)  \\
-                        & Elevator    & 12.713          & 40.453  ($\pm$ 0.319)\\
-                        & Co-housing  & 8.021           & 207.768 ($\pm$ 0.330)\\
+Range             & Map coloring        & 4.629   ($\pm$ 0.199)     \\ % & 2.928                     
+restriction       & Sudoku              & 16.118  ($\pm$ 0.154)     \\ % & 3.367                     
+                  & Elevator            & 40.453  ($\pm$ 0.319)     \\ % & 12.713                    
+                  & Co-housing          & 207.768 ($\pm$ 0.330)     \\ % & 8.021                     
 \midrule
-Connected    & Map coloring  & 1.005           & 1.589   ($\pm$ 0.110)\\
-    clauses       & Sudoku    & 1.476                          & 7.068   ($\pm$ 0.150)    \\
-                        & Elevator    & 1.935           & 6.157   ($\pm$ 0.114)\\
-                        & Co-housing  & 4.001           & 103.633 ($\pm$ 0.131)\\
+Connected         & Map coloring        & 1.589   ($\pm$ 0.110)     \\ % & 1.005                     
+    clauses       & Sudoku              & 7.068   ($\pm$ 0.150)     \\ % & 1.476                     
+                  & Elevator            & 6.157   ($\pm$ 0.114)     \\ % & 1.935                     
+                  & Co-housing          & 103.633 ($\pm$ 0.131)     \\ % & 4.001                     
   \end{tabularx}
@@
 \paragraph{Efficiency}
-Table~\ref{tbl:uitvoering} shows the execution time for several experiments demonstrating the effect of the syntactical restrictions.
+Table~\ref{tbl:uitvoering} shows the execution time for several experiments, demonstrating the effect of the syntactical restrictions.
 Moreover, all efficiency measures (i.e. symmetric predicates, range restriction and connected clauses) are able to improve the execution time.
@@
 All possible pairs of examples are generated and translated to pairwise inequalities using the calculated scores.
-These rankings are given as input for learning.
+A fraction of these rankings are given as input for learning.
 Experiments use different sizes of training sets and different fractions of rankings.
-Noise is simulated by flipping a fraction of the selected rankings.
+Noise is simulated by flipping the inequalities for a fraction of the selected rankings.
 
@@
 
+\newpage
+
 %
@@
 %
-\bibliographystyle{plain}
+\bibliographystyle{aaai}
 \bibliography{Bibliography}

### 2015-10-19 Reviews.tex
Transition to AAAI style

\documentclass[a4paper]{article}

% Essentials
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage[table]{xcolor}
%\usepackage{latexsym,graphicx,graphics,epsfig}

% Hyperlinks
\hypersetup{
    colorlinks,
    linkcolor={red},
    citecolor={blue},
    urlcolor={blue}
}

% Optional but very useful
\usepackage[tmargin=3.5cm, bmargin=3.5cm, lmargin=3cm, rmargin=3cm]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{float}
\usepackage{enumitem}
\usepackage{tabularx,booktabs,multirow}
\usepackage{longtable}
\usepackage{url}
\usepackage{enumitem}
\usepackage{todonotes}
\usepackage{caption}

% Title
\title{Reviews Thesis-Paper}
\author{\textbf{Samuel Kolb}}
\date{\today}

% Indentation
\setlength\parindent{0pt}

\begin{document}

\maketitle

\section{Review 1}
The author introduces a framework for learning both constraints and optimization criteria.
The paper is a cut and paste of the author's thesis and it's hard to understand. 
In this respect, the overall paper organization needs improvements and the text needs an accurate rewriting to allow a better understanding of the content.
Moreover, it is not really clear the difference with the state-of-the-art solutions given the lack of comparison.

\section{Review 2}
To me, the paper presents the following weak points, for which it cannot be accepted:

- It is stated that the paper was born from a thesis and I believe few has been done to convert it
into a proper research paper. The Introduction doesn’t talk about state of the art, possible applications fields
of the presented approach, etc.

- section 2.1 explains FOL in a top-down manner, while usually one starts from atoms, terms until clauses. Moreover no reference is included
- section 2.2 should be moved in the introduction, rather than in the background, since it gives the reader
an idea of the existing methods
- in section 2.3 it is stated that a different implementation has been used for the clausal discovery algorithm, but
it is described nowhere
- section 2.4 doesn’t explain what an IDP theory is
- section 2.5 doesn’t cite any work about MAXSAT

- Section 3 and 4 titles should be made more specific and content-related
- in sect. 3 it is not shown the format (interpretations?clauses?) of the examples.
I think a complete example should represent the KB, the learned constraints, the weights, ecc..
- in sect. 3 the order over the examples is unclear: which is the criterion that allows one to say that an example
comes before another?

- Example 3: why are variables lowercase and constants uppercase? Don’t you use FOL syntax?
- In example 4 how is the function $(1 \cdot c1) + (0 \cdot c2) + (-2 \cdot c3)$ obtained from Table 1?
- No reference for SVMrank is given in sect. 4.3

- section 5.1: the examples should be illustrated in much more detail (what are the learnt constraints for the different domains?)
- section 5.1 - you name a paragraph “accuracy” but no accuracy values are reported on the experiments.
- in Table 2 what do you mean by Time(baseline)?
- “compared to humans” paragraph: the IDP system URL should be given
- section 5.2: what do you mean by “All possible pairs of examples are generated”?

How is this work positioned w.r.t. the state of the art - experimental results do not report such a comparison.

\section{Review 3}
This paper proposes an ILP-based approach to automatically learn hard and soft constraints from examples for use in a Constraint Satisfaction based problem solver.
The paper is more of an informal introduction to the proposed approach than a technical presentation thereof. It lacks justification of the choices, formal description of the methodology, detailed description of the experimental settings, and insight of the experimental results.

There are some formal issues. E.g.:
- in the basic definitions, "an atom is either true or false or a predicate" is incorrect. An atom is by definition a predicate applied to its arguments (if any), that can be true or false depending on an interpretation. true is a predicate without arguments that is always true; false is a predicate without arguments that is always false.
- in Sec. 3, concerning soft constraints, what do the equalities $e_x = e_y$ and $e_a = e_b$ in the example of ranking mean?
- in Sec. 3, concerning soft constraints, in the summation formula, what is 'weight'? it has no index, so it seems that it is independent of the clauses, so it should be moved to the left of the sum operator.

The background information is quite informal, and does not discuss related work that is in competition with the proposed approach (it just discusses related work the proposed approach builds upon).
Also, no explanation is given as to why exactly those systems are chosen to build upon. There are also other systems in the literature that might be used, why the proposed ones are better?

Sec. 4.3: where do the clauses used as constraints come from? I suppose they are not those learned in Sec. 4.2, because they were the hard constraints. Otherwise, if those were the soft constraints, where is the learning of hard constraints? Or, if the clauses learned according to Sec. 4.2 can be either hard or soft constraints, how and where is the distinction made?

\end{document}

### 2015-10-20 Paper.tex
Without context and review processing

@@
 \usepackage{algorithm}
-\usepackage{algpseudocode}
+\usepackage[noend]{algpseudocode}
 \usepackage{amsmath}
@@
   Samuel Kolb\\
-  KULeuven, Leuven, Belgium\\
+  KU Leuven, Leuven, Belgium\\
   samuel.kolb@cs.kuleuven.be
@@
 The learning system for optimization criteria is shown to achieve accurate results and can learn criteria that identify the correct optimal solution even for few examples and noisy rankings.
-Finally, a procedure is outlined that allows weighted first order logic clauses to be used for optimization.
+Finally, a procedure is outlined that allows weighted first order logic clauses to be used for optimization in practice.
 \end{abstract}
 
+% -------------------------------------------------
+% Introduction
+% -------------------------------------------------
+
 \section{Introduction}
-This research tackles the learning of constraints and optimization criteria.
-Non-experts often experience difficulties when modeling a problem using constraints \cite{Wallace:PrinciplesCP}.
+This paper introduces an approach for learning constraints and optimization criteria.
+Using constraints and optimization criteria such as the ones proposed in this paper allows problems to be modeled \emph{declaratively}.
+This means that the representation is decoupled from the process of finding a (optimal) solution for the problem.
+However, non-experts often experience difficulties when modeling a problem using constraints \cite{Wallace:PrinciplesCP}.
 The first goal of this research is to automate the step of modeling a problem, in order to, amongst others, make constraint solving more accessible.
@@
 
-Constraints are represented as logical clauses.
-This representation allows for the use of techniques from the field of Inductive Logic Programming (ILP).
-Specifically, a new implementation of the Clausal Discovery algorithm \cite{DeRaedt:ClausalDiscovery} is used.
-Research on constraint learning is often done independent of ILP.
-By exploiting the parallels between both fields, this research follows the same approach as \cite{Lallouet:LearningCP}.
+Constraint learning attempts to find constraints automatically given examples.
+By using first-order logical clauses as representation for constraints, techniques from Inductive Logic Programming (ILP) can be used for learning.
+Hereby, this research follows the same approach as \cite{Lallouet:LearningCP}, exploiting parallels between ILP and constraint learning.
+In this case, the clausal discovery algorithm \cite{DeRaedt:ClausalDiscovery} is used to find constraints.
 
-Hard constraints are used to distinguish solutions to a problem.
-So called soft constraints can be used to express properties which do not have to be fulfilled by every solution.
-Soft constraints are assigned weights that express their importance.
-Such weighted soft constraints can be used as optimization criteria.
-Optimal solutions are found by maximizing the sum of the weights of soft constraints that are fulfilled by a solution.
+There has been increasing interest in formulating declarative models for constraint satisfaction as well as optimization problems.
+By identifying soft constraints that are only satisfied by some examples and assigning weights to them based on their importance, an optimal solution can be sought.
+The use of weighted soft constraints to model user preferences is inspired, amongst others, by the research of \cite{campigotto2011active}.
+Their work examines attempts to obtain similar results in an interactive context for propositional constraints.
+In this research, first-order constraints are used and optimization criteria are modeled using a first-order version of weighted MAX-SAT.
+This paper also outlines how these constraints can be used in an existing solver to obtain an optimal solution in practice.
 
-The second goal of this research is to learn from user preference and to formulate corresponding optimization criteria.
-By using rankings over a few examples, weighted soft constraints can be found.
-Both the learning of constraints as well as the learning of optimization criteria aim to bridge the gap between problems and their formal representation.
+Since the underlying goal of this research is to facilitate the generation of a formal representation, the input format aims to be easy to use.
+It has been a key focus to make reasonable assumptions concerning the knowledge a user can provide.
+One important difference compared to similar research (including \cite{campigotto2011active}) is that users provide partial rankings over examples instead of absolute scores.
 
-The use of weighted soft constraints is inspired by the research of \cite{campigotto2011active}, amongst others.
-Their work examines attempts to obtain similar results in an interactive context for propositional constraints.
-In this research, first-order logical clauses will be used.
+This paper is based on research that has been conducted during the course of my masters thesis~\cite{kolb2015thesis}.
+The key contribution presented in this paper is a novel approach for learning weighted first order clauses from rankings to capture user preferences.
+In this context an implementation of clausal discovery has been realized that supports soft constraints.
+Moreover, an approach to obtaining optimal solutions for the weighted first order clauses is presented.
 
-The weighted clauses cannot be directly used by a constraint solver.
-Therefore, an approach is outlined that describes how these constraints can be used in an existing solver to obtain an optimal solution in practice.
+% -------------------------------------------------
+% Background
+% -------------------------------------------------
 
-This research has been conducted during the course of my masters thesis~\cite{kolb2015thesis}.
 \section{Background}
@@
 
-\subsection{Clauses}
+\paragraph{Clauses}
 First-order logical clauses consist of a disjunction of literals.
-Literals are defined as being either an atom or a negated atom.
+Literals are defined as being either an atom (positive literal) or a negated atom (negative literal).
 Clauses are often expressed using a body and a head: $\mathit{head} \leftarrow \mathit{body}$.
-All literals containing a negation are grouped in the body and the other literals form the head of the clause.
-For example, $\lnot a_1 \lor \lnot a_2 \lor a_3 \lor a_4$ corresponds to the clause $a_3, a_4 \leftarrow a_1, a_2$.
+Hereby, positive literals are grouped in the head and negative literals are grouped in the body.
+%For example, $\lnot a_1 \lor \lnot a_2 \lor a_3 \lor a_4$ corresponds to the clause $a_3, a_4 \leftarrow a_1, a_2$.
 
-Atoms are logical concepts that can be assigned a truth value.
-In this paper, atoms will be restricted to either the symbols $\mathit{true}$ or $\mathit{false}$ or a predicate.
-Predicates consist of a predicate symbol and a set of terms.
-Terms are either constants or variables and describe objects in a domain.
-Predicates express relations between these objects: $\mathit{symbol}(\mathit{term_1}, ... \mathit{term_n})$.
-Clauses are universally quantified and will, in this research, only contain variables.
+Terms are used to reason about objects and consist of either constants denoting specific objects or variables that can represent any object.
+In this paper constants will be written in uppercase ($\mathit{John}$) and variables in lowercase ($\mathit{x}$).
+Relations between logical objects are represented by predicates, consisting of a predicate symbol and a set of terms (e.g. $\mathit{married(John, x)}$).
+Predicates are atoms and can be assigned a truth value.
 
-\subsection{Learning constraints}
+\paragraph{Learning constraints}
 Constraint learning is a hard problem.
-The search space of constraints is typically very large.
-Furthermore, there are usually only a few examples to learn from, unlike in typical machine learning problems.
+The search space of constraints is typically very large and, unlike in typical machine learning problems, there are usually only few examples to learn from.
 
@@
 
-\subsection{Clausal Discovery}
+\paragraph{Clausal Discovery}
 The clausal discovery algorithm proposed by \cite{DeRaedt:ClausalDiscovery} attempts to learn first order logic clauses from positive examples.
@@
 The result set consists of the most specific clauses covering all examples.
-This research uses the clausal discovery algorithm on a high level but provides a very different implementation than the original clausal discovery engine.
 
-\subsection{Logic Solver}
+\paragraph{Logic Solver}
 The logic system IDP \cite{de2013prototype,wittocx2008idp} is used to calculate whether examples satisfy clauses and is used as a theorem prover.
@@
 
-\subsection{Optimization}
+\paragraph{Optimization}
 The Boolean Satisfiability Problem (SAT) attempts to determine whether a propositional formula can be satisfied.
@@
 By assigning positive integer weights to clauses and maximizing the sum of the weights of satisfied clauses, one obtains the weighted MAX-SAT problem.
-The work by \cite{campigotto2011active} attempts to learn propositional formulas en weights automatically.
-Inspired by their results, this research tries to learn first-order logic optimization criteria.
+The work by \cite{campigotto2011active} attempts to learn propositional formulas and weights automatically using examples and absolute scores.
+Inspired by their results, this research tries to learn first-order logic clauses and weights using examples and rankings.
 The resulting optimization problem can be seen as an extension of weighted MAX-SAT in which the clauses are in first-order logic.
 
+% -------------------------------------------------
+% Problem Statement
+% -------------------------------------------------
+
 \section{Problem Statement}
-This research has two main objectives: learning constraints and learning optimization criteria.
-The system designed to solve these problems attempts to simplify the process of generating formal representations.
-Therefore, it should be easy to use and take into account what kind of information a user can provide.
+The research presented in this paper aims to learn constraints and optimization criteria based on positive examples and user preferences.
+Since it aims to simplify the process of generating formal representations it should be easy to use and require input that a user can reasonably be expected to provide.
 
-\begin{framed}
-  \noindent
-  \begin{minipage}{\textwidth}
-    \paragraph*{Goal 1}
-    Given a set of examples and a limit $t$, the goal of the constraint learning system is to find maximally specific clauses that are satisfied by at least $t$ of the given examples.
-  \end{minipage}
-\end{framed}
+\subsection{Constraint learning}
+\emph{Given a set of examples and a limit $t$, the goal is to find maximally specific clauses that are satisfied by at least $t$ of the given examples.}
+\\\\
+Examples consist of a domain and an exhaustive description of all relations that hold on the domain objects.
+The learned clauses are domain-independent and contain only variables.
 
-Examples describe specific solutions.
-They contain a set of objects and exhaustively describe all relations (predicates) that hold on these objects.
-The learned clauses are independent of the specific objects within the examples and contain only variables.
+% \begin{example} " TODO - necessary?"
+%   Consider sudoku as an example of a problem where hard constraints are to be found.
+%   Given one or more solved sudokus, the learning system has to identify the rules of the game.
+%   Since the learned clauses only contain universally quantified variables, for example $4 \times 4$ sudokus could be used to find constraints.
+%   The learned constraints are independent of the size of the sudoku and can also be used to solve $9 \times 9$ sudokus, for example.
+
+%   In some problems, one is interested in regularities that  often but not always occur.
+%   An example would be regularities in certain types of buildings.
+%   While certain regularities would occur in many buildings, there are often exceptions and different styles.
+%   In this case, one could attempt to find constraints on the structure of the floor plans that are satisfied by a given amount of examples.
+% \end{example}
+
+\subsection{Learning optimization criteria}
+\emph{Given a set of examples and a set of partial rankings over these examples, the goal is to identify soft constraints and weights learn such that the order described by these optimization criteria maximally corresponds with the given rankings.}
+\\\\
+The soft constraints consist of clauses that are satisfied by some but not all of the examples.
+They can be identified using a system that implements the constraint learning task.
+Partial rankings reason over sets of examples and describe an order over them (e.g. the ranking $e_i > e_j$ expresses the users preference of example $e_i$ over example $e_j$).
+The weighted clauses (i.e. tuples $(weight, clause)$) can be used to calculate a score for every example:
+\begin{eqnarray*}
+  score(e) = \sum\limits_{(\mathit{w}, \mathit{c})} \mathit{w} \cdot \mathbf{1}_{c \text{ covers } e}
+\end{eqnarray*}
+
+\begin{example}[Moving]
+  \label{ex:moving}
+  Consider the scenario moving to a new city.
+  The user has to choose a job, a place to live and a school.
+  There are multiple job offers, houses and schools available across different areas of the city.
+  User provided examples include a choice of where to work, where to live and what school is chosen.
+  By comparing and ranking some of the examples, the user can express his preferences.
+  Weighted soft-constraints can now be learned and used to identify the best choice.
+  In this case, a possible soft constraint could be that the school and work place are in the same area.
+  The weights that have been learned indicate how desirable or how undesirable a constraint is.
+  Since the constraints are domain independent they can be used for any (unseen) city.
+\end{example}
+
+Aside from learning optimization criteria, this paper presents an approach to use these optimization criteria to find an optimal solution in practice.
+
+% -------------------------------------------------
+% Approach
+% -------------------------------------------------
+
+\section{Approach}
+In this research, both constraint learning and learning optimization criteria were implemented.
+Clauses are learned using examples.
+The clause learning implementation uses the IDP system to test if clauses satisfy the examples and identify redundant clauses.
+When learning weighted clauses, soft constraints are first identified in the examples using the clause learning implementation.
+These constraints are assigned weights based on the given rankings.
+
+\subsection{Input}
+The input consists of global definitions and a set of examples.
+Definitions consist of either typing information or predicate definitions.
+Typing information relates directly to the problem domain, improves the accuracy and efficiency of the learned clauses and is usually easy to provide for the user.
+Predicate definitions define predicate symbols and the types of their terms.
+Two special types of predicates are allowed: calculated predicates and symmetric predicates.
+Calculated predicates are not included in examples, instead background knowledge is used to generate them.
+This background knowledge is passed directly to the IDP system.
+Symmetric predicates are predicates for which the order of the terms does not matter.
+Every example contains a domain consisting of a set of constants and an exhaustive list of the relations (predicates) that hold on these constants.
+If optimization criteria are to be learned, a set of pairwise over the given examples is also provided.
+Rankings are used instead of absolute scores since rankings only concern few examples at a time.
+Because rankings only ask for relative positioning of a few examples, they are easier to provide.
 
 \begin{example}
-  Consider sudoku as an example of a problem where hard constraints are to be found.
-  Given one or more solved sudokus, the learning system has to identify the rules of the game.
-  Since the learned clauses only contain universally quantified variables, for example $4 \times 4$ sudokus could be used to find constraints.
-  The learned constraints are independent of the size of the sudoku and can also be used to solve $9 \times 9$ sudokus, for example.
-
-  In some problems, one is interested in regularities that  often but not always occur.
-  An example would be regularities in certain types of buildings.
-  While certain regularities would occur in many buildings, there are often exceptions and different styles.
-  In this case, one could attempt to find constraints on the structure of the floor plans that are satisfied by a given amount of examples.
+  \label{ex:moving-input}
+  Consider the moving problem (example~\ref{ex:moving-input}).
+  A model could, for example, use a type \textit{Area} and predicate definitions low\_crime(\textit{Area}), cheap(\textit{Area}), school\_in(\textit{Area}), work\_in(\textit{Area}) and live\_in(\textit{Area}).
+  Examples would that specify various areas $A_i$ and describe the conditions (e.g.
+cheap($A_1$), low\_crime($A_2$)) as well as the users choices (e.g.
+school\_in($A_2$), work\_in($A_2$), live\_in($A_1$)).
+  % Imagine a problem that describes humans with a type \textit{human} and predicates human(\textit{human}), male(\textit{human}) and female(\textit{human}).
+  % An example could contain the constants S and A, both of type \textit{human}, and the relations: human(S), human(A), male(S), female(A).
+  % Clauses to be learned would be: human(x) $\leftarrow$ male(x), human(x) $\leftarrow$ female(x),  female(x) $\lor$ male(x) $\leftarrow$ human(x) and \textit{false} $\leftarrow$ female(x) $\land$ male(x).
 \end{example}
 
-\begin{framed}
-  \noindent
-  \begin{minipage}{\textwidth}
-    \paragraph*{Goal 2}
-    Given a set of examples and a set of partial rankings over these examples, the goal of the optimization system is to learn soft constraints and weights such that the order described by these optimization criteria maximally corresponds with the given rankings.
-  \end{minipage}
-\end{framed}
+\begin{algorithm}
+  \caption{The clausal discovery algorithm}
+  \label{alg:cd}
 
-The soft constraints consist of clauses that are satisfied by some of the examples.
-Rankings are partial orders over the examples: $e_x = e_y > ... > e_a = e_b$.
-The weighted clauses can be used to calculate a score for every example
-\begin{eqnarray*}
-  \sum\limits_{\mathit{weight}, \mathit{c}} \mathit{weight} \cdot v_{c}
-\end{eqnarray*}
-With $v_c = 1$ if the clause $c$ is satisfied by the example and $v_c = 0$ if $c$ is not satisfied.
-These scores allow for a total order over all examples to be defined.
+  \begin{algorithmic}
+  \State $Q \gets \{\square\}$, $\sym{T} \gets \{\}$
+  \While{$\lnot isempty(Q)$}
+    \State $c \gets next(Q)$
+    \If{$covers(c, \textit{Examples})$}
+      \If{$\lnot entails(\sym{T}, c)$}
+        \State $\sym{T} = \sym{T} \cup c$
+      \EndIf
+    \Else
+      \State $Q \gets Q \cup \rho(c)$
+    \EndIf
+  \EndWhile
+  \State $\sym{T} \gets prune(\sym{T})$
+  \State \Return \sym{T}
+  \end{algorithmic}
+\end{algorithm}
 
-\begin{example}
-  Consider the scenario of moving to a new city.
-  Multiple job offers are available, different accommodations are to be considered and there are different schools in the city.
-  Learning weighted soft constraints can make user preferences explicit.
-  For example, it could be desirable that the school is in a low crime area or that the locations of the school and work place are close.
-\end{example}
-
-A third goal of this research is to show how these constraints and especially the optimization criteria can be used in practice to find an optimal solution.
-
-\section{Approach}
-\begin{figure}
-
-  \centering
-    \includegraphics[width=0.8\linewidth]{ApproachOverview.pdf}
-  \caption{Overview approach}
-  \label{fig:struktuur}
-
-\end{figure}
-
-In this research, both constraint learning and learning optimization criteria were implemented.
-An overview is presented in figure~\ref{fig:struktuur}.
-Clauses are learned using examples.
-To find weighted clauses, first the constraint learning system is used to find soft constraints.
-Then the weights for these soft constraints are determined based on the rankings.
-In order to determine the weights, the \svm system is used.
-
-\subsection{Input}
-The input consists of examples and definitions.
-Every example contains a set of constants that are part of the model and an exhaustive list of the relations that hold on these constants.
-Definitions describe either typing information or predicate definitions.
-Typing information relates directly to the problem domain, improves the accuracy and efficiency of the learned clauses, and is usually easy to provide for the user.
-Predicate definitions describe the types of the terms occurring in a predicate.
-Two special types of predicates are allowed: calculated predicates and symmetric predicates.
-Calculated predicates are not listed in examples, instead background knowledge is used to generate them.
-Symmetric predicates are predicates for which the order of the terms does not matter.
-If optimization criteria are to be learned, a set of rankings over the given examples is also provided.
-
-\begin{example}
-  Imagine a problem that describes humans with a type \textit{human} and predicates human(\textit{human}), male(\textit{human}) and female(\textit{human}).
-  An example could contain the constants S and A, both of type \textit{human}, and the relations: human(S), human(A), male(S), female(A).
-  Clauses to be learned would be: human(x) $\leftarrow$ male(x), human(x) $\leftarrow$ female(x),  female(x) $\lor$ male(x) $\leftarrow$ human(x) and \textit{false} $\leftarrow$ female(x) $\land$ male(x).
-\end{example}
 \subsection{Clausal Discovery}
-The clause learning system is based on the clausal discovery algorithm.
+The clause learning system is based on the clausal discovery algorithm (alg.~\ref{alg:cd}).
 The logical system IDP is used for two essential functions: $\mathit{covers}$ and $\mathit{entails}$.
@@
 
-Every clause is subjected to two additional tests that attempt to eliminate redundant clauses early on.
-The subset-test determines whether the clause is a superset of another clause that has been accepted earlier.
-If such a clause exists and that clause covers the same examples as the new clause, the new clause is discarded.
-Otherwise it will be tested whether the clause covers enough examples.
-Clauses that do not cover enough examples are refined by the refinement operator.
-The refined clauses will be added to the working queue.
-If a clause covers enough examples, the $\mathit{entails}$ functions is calculated to decide whether the accepted clause should be added to the result set.
+Starting with the empty clause, clauses are refined until they cover enough examples.
+If a clause covers enough examples it is added to result set, provided it is not entailed by background knowledge or other clauses in the result set.
+To avoid extensive calls to $\mathit{covers}$ and $\mathit{entails}$, two additional tests are used.
+The subset-test rejects clauses that are supersets of a clause that covers the same or all examples and has been accepted earlier.
+Because clauses can be formulated in multiple ways, a representative-test is used to remove clauses that are not in canonical form.
 
-Refinement is an important step.
-Clauses are extended to be more general and potentially cover more examples.
-The refinement operator uses a list of atoms that can potentially be added to a clause.
+In the refinement step, clauses are extended to be more general and potentially cover more examples.
+The refinement operator uses a list of atoms that may be added to a clause.
 This list is calculated in advance, given a maximal amount of variables.
-Within the body and head of a clause, atoms may only be added in a specific order.
-Additionally, several restrictions are in place.
-Clauses must be connected, i.e. new atoms must always contain a variable that has already occurred.
-Secondly, clauses must be range-restricted, which means that no new variables may be introduced in the head of the clause.
-By looking at the positions of the atoms in the list, clauses can be reduced to a unique form.
-This reduces the amount of duplicate clauses generated.
+Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.
+Object Identity is used to specify that variables with different names must denote different objects.
+Additionally there are two syntactic restrictions.
+Clauses must be connected (i.e. new atoms must always contain a variable that has already occurred) and range-restricted (i.e. no new variables may be introduced in the head of the clause).
 
 \subsection{Optimization}
-The first step in finding weighted soft constraints is to identify soft constraints by using the constraint learning system and a low threshold.
-Examples are characterized by the soft constraints (clauses) which it satisfies.
-Therefore, every example can be translated to a vector of boolean features.
-Each feature $f_i$ corresponds with one of the clauses and is \emph{1} if the clause covers the example and \emph{0} if it does not.
-Existing machine learning software can be used to learn a linear \emph{scoring} function $\sum_i w_i \cdot f_i$ over these features based on the given rankings.
-Hereby, weights are assigned to every feature to express its importance.
+The first step in finding weighted soft constraints is to identify soft constraints by using the constraint learning system with a (low) threshold.
+Examples are characterized by the soft constraints (clauses) which they satisfy.
+Therefore, every example~$e$ can be translated to a vector of boolean features by introducing a feature~$f_i$ for every clause~$c_i$, with $f_i = \mathbf{1}_{c_i \text{ covers } e}$.
+Existing software can be used to learn a linear scoring function $\sum_i w_i \cdot f_i$ over these features based on the given rankings.
 For an unseen example, the feature vector is computed by calculating what clauses cover the new example.
@@
   Consider examples~$e_1, e_2, e_3$, rankings $e_1 > e_2, e_2 > e_3$ and clauses $c_1, c_2, c_3$.
-  If the clauses cover examples according to table~\ref{tbl:cover_examples}, then the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.
-  An example $e_4$ that is covered by~$c_1$ and~$c_2$, is then assigned a score of $1$.
+  If the clauses cover examples according to table~\ref{tbl:cover_examples}, then, for example, the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.
+  A new example that is covered by~$c_1$ and~$c_2$, would be assigned a score of $1$, according to this function.
   This score has no value in the absolute sense, it can only be used to compare it to other examples ranked by the same function.
@@
 
-  \begin{table}[!htp]
+  \begin{table}
   \caption{Clause coverage}
   \label{tbl:cover_examples}
-  \begin{tabularx}{\textwidth}{c|ccc|X}
-    \textbf{Clauses}  &$c_1$    & $c_2$   & $c_3$   & \textbf{Feature vector}\\
+  \begin{tabularx}{\linewidth}{c|l|X}
+    \textbf{Example} & \textbf{Covered by} & \textbf{Feature vector}\\
     \toprule
-    $e_1$         & Covers  &       &       & 1, 0, 0\\
-    $e_2$         &       & Covers  &       & 0, 1, 0\\
-    $e_3$         & Covers  & Covers  & Covers  & 1, 1, 1\\
+    $e_1$             & $c_1$               & (1, 0, 0) \\
+    $e_2$             & $c_2$               & (0, 1, 0) \\
+    $e_3$             & $c_1$, $c_2$, $c_3$ & (1, 1, 1) \\
   \end{tabularx}
   \end{table}
-
-  Considering an extended form of the human example, the optimization function could be:
-  1 $\cdot$ (funny(x)~$\leftarrow$~human(x)) + (-2) $\cdot$ (lazy(x)~$\leftarrow$~human(x)), for example.
-  Clauses with a weight of zero are ignored.
 \end{example}
 
-The \svm{} software was chosen to find the scoring function since it is a robust and efficient system.
-Contrary to many other ranking systems, it uses a linear model.
+The \svm{} software was chosen to find the scoring function since it uses a linear model and offers an efficient implementation.
 The weights that it assigns to the features can be used directly as weights for the optimization criteria.
 The input format that is used by \svm{} is also supported by many other learn-to-rank implementations.
-Therefore, other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be plugged into the implementation if so desired.
+Therefore, other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.
 
@@
 This optimization task can be solved in IDP, using inductive definitions, aggregates and minimization.
-The only limitation is that the weights must be integer values.
+The only limitation is that the current version only supports integer values.
 Therefore, the weights of the clauses are divided by the smallest absolute weight, multiplied by a constant and rounded to the closest integer if necessary.
-Removing this restriction for IDP is currently being researched and future releases might be able to deal with rational numbers directly.
 
@@
 For every clause a predicate $t(i)$ is added to capture the truth value of the clause.
-A function $\mathit{cost}(i)$ specifies the cost of not satisfying the clause, which is equal to the integer weight of the clause.
+A function $\mathit{cost}(i)$ specifies the cost of not satisfying the clause, which is equal to the weight of the clause.
 \begin{eqnarray*}
@@
 
-Using $t$ and $\mathit{cost}$, a function $\mathit{actual}(i)$ is then defined in IDP as 0 if $t(i)$ and $\mathit{cost}(i)$ if $\lnot t(i)$.
+Using $t$ and $\mathit{cost}$, a function $\mathit{actual}(i)$ is then defined in IDP as $\mathbf{1}_{\lnot t(i)} \cdot \mathit{cost}(i)$.
 This function is used in the term $\sum_i actual(i)$ to be minimized, which will allow IDP to search for an optimal solution.
+
+% -------------------------------------------------
+% Evaluation
+% -------------------------------------------------
 
@@
 Several experiments aim to measure the accuracy and efficiency of the learning systems for constraints and optimization criteria.
+In order to account for non-determinism, experiments measuring execution times or optimization scores are usually performed eight times and the results are averaged.
 
@@
 Four problems have been used to evaluate constraint learning.
-Two are existing problems: map coloring and sudoku.
-Map coloring is a problem where countries are assigned colors and neighboring countries may not have the same color.
-The examples to learn from each consist of a set of three countries that are correctly colored.
-For sudoku a $4 \times 4$ solved sudoku is given.
-Furthermore, there is the elevator problem with a soft constraint that holds on 2 of the 3 given examples.
-The last problem is the co-housing problem with 4 hard constraints and 5 given examples.
+The first problem is map coloring, where countries are assigned colors and neighboring countries may not have the same color.
+Two examples containing each three correctly colored countries are given.
+The second problem is sudoku and a single, solved $4 \times 4$ sudoku is given.
+For the third problem (elevator) three examples have been generated, two of which respect the underlying soft constraint.
+The last problem (co-housing) contains four hard constraints and five examples have been generated that respect the constraints.
 
-\paragraph{Accuracy}
-The essential constraints were found for all problems.
-Additionally, constraints were found that describe some structure in the problems, such as that countries are never their own neighbor.
+\begin{question}[Accuracy]
+  Are the essential constraints discovered and what influence do different parameters have on the accuracy?
+\end{question}
+For all problems the essential constraints are found.
+Often additional constraints were found that describe some structure in the problems.
+For example, for the map coloring problem a learned constraint states that countries are never their own neighbor.
 These kind of constraints may help a constraint solver to work more efficiently.
 
-If the limits on the number of variables or literals per clause are too large, constraints are found that over-fit the training data.
-These constraints are too specific and will exclude valid solutions.
+The learning process is parameterized by the maximal amount of variables and literals allowed per clause.
+If these limits are too large, constraints are found that over-fit the training data.
+These constraints are too specific and will exclude valid solutions that are not in the training set.
 On the other hand, if the chosen limits are too small, the necessary constraints will not be found.
 
+Over-fitting can be addressed by removing constraints that exclude valid solutions (manually or automatically) as well as providing more (or larger) training examples.
+Negative examples can be used to detect under-fitting, indicating that the parameter values are too small.
+
 \begin{table}
-  \caption{Execution times overview (average over 8 runs)}
+  \caption{Execution times overview}
   \begin{tabularx}{\linewidth}{rl|ll}
@@
 
-\paragraph{Efficiency}
-Table~\ref{tbl:uitvoering} shows the execution time for several experiments, demonstrating the effect of the syntactical restrictions.
-Moreover, all efficiency measures (i.e. symmetric predicates, range restriction and connected clauses) are able to improve the execution time.
+\begin{question}[Efficiency]
+  How fast is the clause learning system and what is the effect of various design decisions and input on the execution time?
+\end{question}
+
+Table~\ref{tbl:uitvoering} shows the execution times for several experiments.
+It shows that smaller problems can be solved efficiently and shows that removing the syntactical restrictions exponentially increase the search space.
+
+All efficiency measures (i.e. symmetric predicates, subset test and representative test) have been able to improve the execution time, sometimes by more than 50\%.
+The overhead they introduce is more than compensated by the efficiency gains they cause.
 
@@
 Therefore it would be useful to adapt these parameters dynamically.
+
 Adding additional examples only increases the execution time by a constant factor.
@@
 
-\paragraph{Compared to humans}
+\begin{question}[Compared to humans]
+  How do learned constraints compare to human programmed constraints?
+\end{question}
 Human programmed theories for map coloring and sudoku are available on the website of the IDP system.
@@
 This time is measured for the learned theories as well as for hand made theories.
-The learned theories are slightly adapted to be able to solve the same problems and corrected if they contains an unfair advantage.
-  \begin{table}
+The learned theories are slightly adapted to be able to solve the same problems and corrected if they contain an unfair advantage.
+
+  \begin{table}[!htp]
     \caption{CPU times human vs. learned theory}
     \begin{tabularx}{\linewidth}{lr|X}
-      \textbf{Problem} & \textbf{Type} & \textbf{Mean CPU time (s)} \\
+      \textbf{Problem} & \textbf{Type} & \textbf{Average CPU time (s)} \\
       \toprule
@@
 The efficiency of the learning system for optimization criteria depends mainly on the efficiency of learning soft constraints.
-Therefore, the experiments in this section are focused on the accuracy of the optimization criteria and the influence of different factors on the accuracy.
+Therefore, the experiments in this section are focused on the accuracy of the optimization criteria and the influence of different factors.
 
-The previously mentioned problem of selecting a place to live, a workplace and a school is used as an example.
-In the chosen scenario, there are 18 possible options which form the available examples.
+The moving problem with input according to example~\ref{ex:moving-input} is used to evaluate the learning of optimization criteria.
+18 possible configurations are used as available examples.
+Four weighted soft constraints are used to represent the underlying model for the users preferences.
 Two approaches are used for evaluation.
@@
 The second approach uses all examples as train as well as test set.
-An underlying model of four weighted soft constraints is used to calculate a score for each example.
-All possible pairs of examples are generated and translated to pairwise inequalities using the calculated scores.
-A fraction of these rankings are given as input for learning.
-Experiments use different sizes of training sets and different fractions of rankings.
-Noise is simulated by flipping the inequalities for a fraction of the selected rankings.
 
-The experiments are run 8 times, and the examples and rankings to be used as well as the rankings to be inverted are selected randomly.
-Optimization criteria are searched based on the training set and the selected inequalities.
-These criteria are used to predict the better example out of every pair of examples.
-Experiments are scored by comparing the amount of correctly predicted pairs (1.0 indicating full agreement).
+For every experiment a fraction of the training examples is selected randomly.
+Pairwise rankings are generated by picking two of the selected examples and predicting the better example using the underlying model.
+The selected examples and a random subset of the possible rankings are then used as input for learning.
+Noise is simulated by flipping a subset of the input rankings.
+
+To evaluate the learned optimization criteria, all possible pairs of examples in the test set are generated.
+The learned model is used to predict the better example for every pair.
+The score is then calculated as the fraction of correctly predicted pairs. 
+Pairs for which the underlying model ranks both example the same are omitted.
+
+\begin{question}[Accuracy]
+  How accurately can learned optimization criteria approximate underlying models?
+\end{question}
 
 \begin{figure}
-  \begin{minipage}{0.5\textwidth}
-
   \centering
@@
   \label{fig:fractie}
+\end{figure}
 
-  \end{minipage}
-  \begin{minipage}{0.5\textwidth}
+Figure~\ref{fig:fractie} shows how the scores improve as the amount of examples and the fraction of included preferences increases.
+In all cases, more than half the pairs of examples are correctly predicted and high scores can be obtained even for small datasets.
+Additional experiments have shown that even for lower scores the learned optimization criteria are often capable of identifying the correct optimal solution.
 
+The standard underlying model can be directly expressed using the soft constraints that can be learned.
+Even though clausal theories can be very expressive, important restrictions have been placed on the learned clauses in this paper.
+In order to test the accuracy for models which cannot be directly expressed, a model consisting of two disconnected clauses has been tested as well.
+While there are limits to the expressibility, the learned optimization criteria were able to obtain similar scores and seem to be robust with respect to the exact formulation.
+
+\begin{question}[Noise]
+  What is the influence of noise on the accuracy?
+\end{question}
+
+The influence of noise is shown in figure~\ref{fig:ruis}.
+Hereby, the second approach of testing is applied, using potentially overlapping train- and test sets.
+High scores are obtained, even despite significant levels of noise.
+The figure also shows that providing more rankings improves the robustness of the algorithm, even if the relative amount of noise remains unchanged.
+
+\begin{figure}
   \centering
@@
   \label{fig:ruis}
-
-  \end{minipage}
 \end{figure}
 
-Figure~\ref{fig:fractie} shows how the scores improve as the amount of examples and the fraction of included inequalities increases.
-In all cases, more than half the pairs of examples are correctly predicted and high scores can be obtained even for small datasets.
-Other experiments have shown that even for lower scores the optimization criteria are often capable of finding an optimal solution.
-
-The underlying model can be directly expressed using the soft constraints that can be learned.
-However, similar scores are obtained for models which cannot be directly modeled (e.g. the use of weighted soft constraints which are disconnected).
-While there are limits to the expressibility, the learned criteria are shown to be robust with respect to the exact formulation.
-
-The influence of noise is shown in figure~\ref{fig:ruis}.
-Hereby, the second approach of testing is used, where all the examples are used as test set.
-The algorithm obtains high scores despite significant levels of noise.
-Figure~\ref{fig:ruis} also shows that providing more rankings improves the robustness of the algorithm, even if the relative amount of noise remains unchanged.
-Using the first approach with a training set of 40\%, scores over 0.7 can be obtained for noise levels of 20\%.
-
+\begin{question}[Threshold]
+  What is the effect of the soft constraint threshold on the accuracy?
+\end{question}
   \begin{table}
@@
 
-Finally, table~\ref{tbl:limiet} shows that increasing the threshold used for finding soft constraints does not improve the score.
-This experiment used a 40\% training set and 40\% of the available rankings.
-However, if the size of the problem and examples is increased, higher threshold will likely be appropriate.
+Table~\ref{tbl:limiet} demonstrates that increasing the threshold used for finding soft constraints does not improve the score.
+This experiment used 40\% of the examples as training set and 40\% of the available rankings.
+However, if the size of the problem and examples is increased, a higher threshold will likely be appropriate.
+
+% -------------------------------------------------
+% Conclusion
+% -------------------------------------------------
 
 \section{Conclusion}
-The research in this thesis has focused on solving two problems: 1) Learning constraints; and 2) Learning optimization criteria.
+The research in this thesis has focused on automatically acquiring constraints and optimization criteria from examples and rankings.
 Implementation for both clause learning and clausal optimization have been provided that accomplish these tasks using first order logic clauses.
 
-The constraint learning system has been able to learn the relevant hard and soft constraints in all experiments.
-As such, the first goal of this research has been accomplished.
-All problems have been solved in limited time (less than a minute) and often even within a few seconds.
+The constraint learning implementation has been able to learn the relevant hard and soft constraints in all experiments.
 For each problem, only a small number of examples was given to learn from.
-The system requires only a minimal amount of information from the user.
-However, the system also allows for the use of expressive background knowledge.
+The system requires only a minimal amount of information from the user, however, it also allows for the use of expressive background knowledge.
 The learned constraints are domain independent, which facilitates the construction of positive examples.
 
-The second goal of this research, learning optimization criteria, has been accomplished as well.
-Optimization criteria are discovered that enable an optimal solution to be found.
-Even for small datasets and noise on the rankings, constraints are found that enable most examples to be ranked correctly.
+Using the constraint learning implementation, the goal of learning optimization criteria has been accomplished.
+Optimization criteria can be learned that enable optimal solutions to be found.
+Even for small datasets and noisy rankings, constraints are found that enable most examples to be ranked correctly.
 
@@
 Rankings expressing that examples are equally ranked are currently ignored.
-However, it could be interesting to incorporate this information into the algorithm, whenever it is provided explicitly.
+Whenever provided explicitly, however, it could be interesting to incorporate this information into the algorithm.
 
 As mentioned earlier, the domain independence of the clauses has several advantages.
-In some cases, however, specific objects are inherently present in the problem.
+In some cases, however, specific objects are inherently present in any problem instantiation.
 The constraint learning system could be enhanced to include such global constants.

### 2015-10-22 Paper.tex
Improved paper, before example rewrite and clean up

@@
 \begin{abstract}
-This paper presents a system that is able to automatically learn both constraints from examples and optimization criteria from rankings.
-Both hard and soft constraints can be learned and constraints are represented as first order logical clauses.
-Choosing such clauses as representations enables the use of ILP techniques such as clausal discovery.
-In this research, optimization criteria are represented as weighted clauses.
-For all examples, the clause learning system could find the essential constraints efficiently.
-The learning system for optimization criteria is shown to achieve accurate results and can learn criteria that identify the correct optimal solution even for few examples and noisy rankings.
-Finally, a procedure is outlined that allows weighted first order logic clauses to be used for optimization in practice.
+% This paper presents a system that is able to automatically learn both constraints from examples and optimization criteria from rankings.
+% Both hard and soft constraints can be learned and constraints are represented as first-order logical clauses.
+% Choosing such clauses as representations enables the use of ILP techniques such as clausal discovery.
+% In this research, optimization criteria are represented as weighted clauses.
+% For all examples, the clause learning system could find the essential constraints efficiently.
+% The learning system for optimization criteria is shown to achieve accurate results and can learn criteria that identify the correct optimal solution even for few examples and noisy rankings.
+% Finally, a procedure is outlined that allows weighted first-order logic clauses to be used for optimization in practice.
+
+While there exist several approaches in the constraint programming community to learn a constraint theory, few of them have considered the learning of constraint optimization problems.
+To alleviate this situation, we introduce an initial approach to learning first-order weighted MAX-SAT theories. 
+It employs inductive logic programming techniques to learn a set of first-order clauses and then uses preference learning techniques to learn the weights of the clauses.
+In order to learn these weighted clauses, the clausal optimization system uses examples of possible worlds and a set of preferences that state what examples are preferred over others.
+% "Change" The examples take the form of rankings that state that one possible world is preferred over another one.
+The technique is also empirically evaluated on a number of examples.
+
 \end{abstract}
@@
 \section{Introduction}
-This paper introduces an approach for learning constraints and optimization criteria.
-Using constraints and optimization criteria such as the ones proposed in this paper allows problems to be modeled \emph{declaratively}.
-This means that the representation is decoupled from the process of finding a (optimal) solution for the problem.
-However, non-experts often experience difficulties when modeling a problem using constraints \cite{Wallace:PrinciplesCP}.
-The first goal of this research is to automate the step of modeling a problem, in order to, amongst others, make constraint solving more accessible.
-In this approach a few examples of what a user considers valid solutions are used by a computer system to find constraints.
-These constraints can then be used by an existing constraint solver.
+% This paper introduces an approach for learning constraints and optimization criteria.
+% Using constraints and optimization criteria such as the ones proposed in this paper allows problems to be modeled \emph{declaratively}.
+% This means that the representation is decoupled from the process of finding a (optimal) solution for the problem.
+% However, non-experts often experience difficulties when modeling a problem using constraints \cite{Wallace:PrinciplesCP}.
+% The first goal of this research is to automate the step of modeling a problem, in order to, amongst others, make constraint solving more accessible.
+% In this approach a few examples of what a user considers valid solutions are used by a computer system to find constraints.
+% These constraints can then be used by an existing constraint solver.
 
-Constraint learning attempts to find constraints automatically given examples.
-By using first-order logical clauses as representation for constraints, techniques from Inductive Logic Programming (ILP) can be used for learning.
-Hereby, this research follows the same approach as \cite{Lallouet:LearningCP}, exploiting parallels between ILP and constraint learning.
-In this case, the clausal discovery algorithm \cite{DeRaedt:ClausalDiscovery} is used to find constraints.
+% Constraint learning attempts to find constraints automatically given examples.
+% By using first-order logical clauses as representation for constraints, techniques from Inductive Logic Programming (ILP) can be used for learning.
+% Hereby, this research follows the same approach as \cite{Lallouet:LearningCP}, exploiting parallels between ILP and constraint learning.
+% In this case, the clausal discovery algorithm \cite{DeRaedt:ClausalDiscovery} is used to find constraints.
 
-There has been increasing interest in formulating declarative models for constraint satisfaction as well as optimization problems.
-By identifying soft constraints that are only satisfied by some examples and assigning weights to them based on their importance, an optimal solution can be sought.
-The use of weighted soft constraints to model user preferences is inspired, amongst others, by the research of \cite{campigotto2011active}.
-Their work examines attempts to obtain similar results in an interactive context for propositional constraints.
-In this research, first-order constraints are used and optimization criteria are modeled using a first-order version of weighted MAX-SAT.
-This paper also outlines how these constraints can be used in an existing solver to obtain an optimal solution in practice.
+% There has been increasing interest in formulating declarative models for constraint satisfaction as well as optimization problems.
+% By identifying soft constraints that are only satisfied by some examples and assigning weights to them based on their importance, an optimal solution can be sought.
+% The use of weighted soft constraints to model user preferences is inspired, amongst others, by the research of \cite{campigotto2011active}.
+% Their work examines attempts to obtain similar results in an interactive context for propositional constraints.
+% In this research, first-order constraints are used and optimization criteria are modeled using a first-order version of weighted MAX-SAT.
+% This paper also outlines how these constraints can be used in an existing solver to obtain an optimal solution in practice.
 
-Since the underlying goal of this research is to facilitate the generation of a formal representation, the input format aims to be easy to use.
-It has been a key focus to make reasonable assumptions concerning the knowledge a user can provide.
-One important difference compared to similar research (including \cite{campigotto2011active}) is that users provide partial rankings over examples instead of absolute scores.
+% Since the underlying goal of this research is to facilitate the generation of a formal representation, the input format aims to be easy to use.
+% It has been a key focus to make reasonable assumptions concerning the knowledge a user can provide.
+% One important difference compared to similar research (including \cite{campigotto2011active}) is that users provide partial rankings over examples instead of absolute scores.
 
-This paper is based on research that has been conducted during the course of my masters thesis~\cite{kolb2015thesis}.
-The key contribution presented in this paper is a novel approach for learning weighted first order clauses from rankings to capture user preferences.
-In this context an implementation of clausal discovery has been realized that supports soft constraints.
-Moreover, an approach to obtaining optimal solutions for the weighted first order clauses is presented.
+% This paper is based on research that has been conducted during the course of my masters thesis~\cite{kolb2015thesis}.
+% The key contribution presented in this paper is a novel approach for learning weighted first-order clauses from rankings to capture user preferences.
+% In this context an implementation of clausal discovery has been realized that supports soft constraints.
+% Moreover, an approach to obtaining optimal solutions for the weighted first-order clauses is presented.
+
+An important class of declarative models is concerned with (discrete) constrained optimization problems.
+These combine constraint satisfaction problems with an optimization function that specifies which solutions are optimal.
+While these models are powerful and elegant, they are often also hard to obtain.
+\cite{Wallace:PrinciplesCP} shows that the task of modeling constraint satisfaction problems is difficult, certainly for non-experts.
+This has motivated several researchers to investigate whether such models can be learned.
+
+Learning constraints has been addressed in constraint programming (e.g.
+\cite{Beldiceanu:ModelSeeker}, \cite{bessiere2013constraint}) and inductive logic programming (e.g.
+\cite{DeRaedt:ClausalDiscovery}).
+Recent efforts \cite{Lallouet:LearningCP} have attempted to combine these two fields.
+Furthermore, \cite{campigotto2011active}) have learned propositional MAX-SAT theories from examples together with their target value.
+%"Expand something here yet?" ...%Each example ....
+Also, work in statistical relational learning (SRL) on learning the structure of Markov Logic networks (MLN) or other SRL models is relevant, as MAP inference can be used to query for the most likely state.
+
+We use weighted first-order logical theories to represent constraint optimization problems.
+The weights will then capture the optimization function, and (first-order) weighted MAX-SAT can be used to determine the \emph{best} solutions.
+Furthermore, we shall learn such models by combining inductive logic programming (ILP) principles (such as clausal discovery \cite{DeRaedt:ClausalDiscovery}) with preference learning.
+ILP will be used to learn the clausal theories, and preference learning to determine the weights.
+Therefore, users will provide examples of possible worlds and a set of preferences that specify which examples are preferred over other ones.
+% "Unify usage of examples / preferences" Therefore, the examples will take the form of rankings, where each ranking specifies that one particular possible world is preferred over another one.
+This setting extends traditional constraint learning and ILP approaches with the ability to learn soft constraint and to integrate constraint satisfaction and optimization.
+It extends the approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than from examples with target values.
+Moreover, it differs the SRL and MLN approaches in that one learns from preferences rather than from examples.
+ % and rankings are used rather than possible worlds + the target value of the examples, and differs from the SRL and MLN approaches in that one learns from rankings rather than from examples.
+
+This paper is organized as follows.
+Section~2 will explain some background knowledge and related work.
+In Section~3 we shall introduce the formalism and the setting and in Section~4 the algorithm and the clausal optimization system to solve this problem.
+In Section~5 we shall report on some experiments, and finally, in Section~6, we conclude.
+% "Rewrite" In Section 2, we shall introduce the formalism and the setting,
+% "Rewrite" in Section 3, the algorithm and system X to solve this problem, in Section 4, 
+% "Rewrite" we shall report on some experiments, and Finally, in Section 5 ,we conclude
 
@@
 \paragraph{Clauses}
-First-order logical clauses consist of a disjunction of literals.
-Literals are defined as being either an atom (positive literal) or a negated atom (negative literal).
-Clauses are often expressed using a body and a head: $\mathit{head} \leftarrow \mathit{body}$.
+In order to reason about clauses, the standard notions of (function free) first-order logic are described.
+An atom $p(t_1, t_2, ..., t_n)$ consists of a predicate symbol $p$ that is applied to $n$ terms $t_i$.
+Terms are either constants or variables.
+In this paper constants will be written in uppercase ($\mathit{John}$) and variables in lowercase ($\mathit{x}$).
+A literal can be an atom $a$ (positive literal) or a negated atom $\lnot a$ (negative literal).
+Clauses are disjunctions of literals $l_1 \lor l_2 \lor ... \lor l_k$ and are assumed to be universally quantified.
+A clause can be expressed using a body and a head $\mathit{head} \leftarrow \mathit{body}$.
 Hereby, positive literals are grouped in the head and negative literals are grouped in the body.
-%For example, $\lnot a_1 \lor \lnot a_2 \lor a_3 \lor a_4$ corresponds to the clause $a_3, a_4 \leftarrow a_1, a_2$.
+Sometimes clauses are also written as a set of literals $\{l_1, l_2, ..., l_k\}$.
 
-Terms are used to reason about objects and consist of either constants denoting specific objects or variables that can represent any object.
-In this paper constants will be written in uppercase ($\mathit{John}$) and variables in lowercase ($\mathit{x}$).
-Relations between logical objects are represented by predicates, consisting of a predicate symbol and a set of terms (e.g. $\mathit{married(John, x)}$).
-Predicates are atoms and can be assigned a truth value.
+Atoms are grounded if they only contain constants.
+A Herbrand interpretation is a set of ground atoms.
+All atoms in the interpretation are assumed to be true and all other possible ground atoms are assumed to be false.
+$\sym{I} \models c$ is used to denote that interpretation \sym{I} satisfies the clause $c$.
+
+\paragraph{Optimization}
+The Boolean Satisfiability Problem (SAT) attempts to determine whether a propositional formula can be satisfied.
+Formulas can be rewritten into a special form called the Conjunctive Normal Form (CNF), which consists of a conjunction of clauses (disjunctions).
+An extension of this problem is the Maximum Satisfiability Problem (MAX-SAT), which is an optimization problem and attempts to maximize the amount of satisfied clauses.
+By assigning positive integer weights to clauses and maximizing the sum of the weights of satisfied clauses, one obtains the weighted MAX-SAT problem.
+
+Related work \cite{campigotto2011active} has creates a weighted MAX-SAT problem by learning propositional clauses and weights using examples and absolute scores.
+Inspired by their results, our research tries to learn first-order logic clauses and weights using examples and rankings.
+The resulting optimization problem can be seen as an extension of weighted MAX-SAT in which the clauses are in first-order logic.
 
@@
 \paragraph{Clausal Discovery}
-The clausal discovery algorithm proposed by \cite{DeRaedt:ClausalDiscovery} attempts to learn first order logic clauses from positive examples.
+The clausal discovery algorithm \cite{DeRaedt:ClausalDiscovery} attempts to learn first-order logic clauses from positive examples.
 Using a refinement operator, the search space is traversed by incrementally generalizing clauses until they cover all examples, starting from the empty clause~($\square$).
@@
 
-\paragraph{Logic Solver}
-The logic system IDP \cite{de2013prototype,wittocx2008idp} is used to calculate whether examples satisfy clauses and is used as a theorem prover.
-It supports an extension of first order logic ($FO(\cdot)$) and can also be used to generate new solutions.
-Background knowledge is given as an IDP theory.
-
-\paragraph{Optimization}
-The Boolean Satisfiability Problem (SAT) attempts to determine whether a propositional formula can be satisfied.
-Formulas can be rewritten into a special form called the conjunctive normal form (CNF), which consists of a conjunction of clauses (disjunctions).
-An extension of this problem is the Maximum Satisfiability Problem (MAX-SAT).
-In MAX-SAT the amount of satisfied clauses is maximized.
-By assigning positive integer weights to clauses and maximizing the sum of the weights of satisfied clauses, one obtains the weighted MAX-SAT problem.
-The work by \cite{campigotto2011active} attempts to learn propositional formulas and weights automatically using examples and absolute scores.
-Inspired by their results, this research tries to learn first-order logic clauses and weights using examples and rankings.
-The resulting optimization problem can be seen as an extension of weighted MAX-SAT in which the clauses are in first-order logic.
+% \paragraph{Logic Solver} " v ------ v   Dissolve"
+% The logic system IDP \cite{de2013prototype,wittocx2008idp} is used to calculate whether examples satisfy clauses and is used as a theorem prover.
+% It supports an extension of first-order logic ($FO(\cdot)$) and can also be used to generate new solutions.
+% Background knowledge is given as an IDP theory.
 
@@
 \\\\
-Examples consist of a domain and an exhaustive description of all relations that hold on the domain objects.
+Examples are Herbrand interpretations.
+They consist of set of constants, the domain, and all ground atoms over these constants that are true.
 The learned clauses are domain-independent and contain only variables.
@@
 \subsection{Learning optimization criteria}
-\emph{Given a set of examples and a set of partial rankings over these examples, the goal is to identify soft constraints and weights learn such that the order described by these optimization criteria maximally corresponds with the given rankings.}
+\emph{Given a set of examples \sym{E} and a set of preferences \sym{P} over these examples, the goal is to learn soft constraints and weights such that the order described by these optimization criteria maximally corresponds with~\sym{P}.}
 \\\\
-The soft constraints consist of clauses that are satisfied by some but not all of the examples.
+Every preference $p \in \sym{P}$ describes a relative order $e_1 \succ e_2 \succ ... \succ e_n$ over a subset of examples $\{e_1, e_2, ..., e_n\} \subseteq \sym{E}$.
+A preference $e_1 \succ e_2$ indicates that out of examples $e_1$ and $e_2$ the user prefers $e_1$.
+
+The soft constraints consist of clauses \sym{C} that are satisfied by some but not all of the examples.
 They can be identified using a system that implements the constraint learning task.
-Partial rankings reason over sets of examples and describe an order over them (e.g. the ranking $e_i > e_j$ expresses the users preference of example $e_i$ over example $e_j$).
-The weighted clauses (i.e. tuples $(weight, clause)$) can be used to calculate a score for every example:
+The weighted clauses \sym{W} are tuples $\mathbb{R} \times \sym{C}$ that can be used to calculate a score for any example $e$:
 \begin{eqnarray*}
-  score(e) = \sum\limits_{(\mathit{w}, \mathit{c})} \mathit{w} \cdot \mathbf{1}_{c \text{ covers } e}
+  score(e) = \sum\limits_{(\mathit{w}, \mathit{c}) \in \sym{W}} \mathit{w} \cdot \mathbf{1}_{e \models c}
 \end{eqnarray*}
 
-\begin{example}[Moving]
+\begin{example}[Moving] "Rewrite"
   \label{ex:moving}
   Consider the scenario moving to a new city.
+  Predicates $\mathit{live\_in}, \mathit{work\_in}$ and $\mathit{school\_in}$ indicate the choice of housing, job and school.
+  There might be different areas in the city which
+
   The user has to choose a job, a place to live and a school.
@@
 In this research, both constraint learning and learning optimization criteria were implemented.
-Clauses are learned using examples.
-The clause learning implementation uses the IDP system to test if clauses satisfy the examples and identify redundant clauses.
-When learning weighted clauses, soft constraints are first identified in the examples using the clause learning implementation.
-These constraints are assigned weights based on the given rankings.
+A two-step approach is used to learn optimization criteria.
+First, clauses learning is used to find soft constraints in the given examples.
+Then, user-provided preference information is used to learn weights for these constraints.
+
+% "Dissolve" The clause learning implementation uses the IDP system to test if clauses satisfy the examples and identify redundant clauses.
 
 \subsection{Input}
-The input consists of global definitions and a set of examples.
-Definitions consist of either typing information or predicate definitions.
-Typing information relates directly to the problem domain, improves the accuracy and efficiency of the learned clauses and is usually easy to provide for the user.
-Predicate definitions define predicate symbols and the types of their terms.
-Two special types of predicates are allowed: calculated predicates and symmetric predicates.
-Calculated predicates are not included in examples, instead background knowledge is used to generate them.
-This background knowledge is passed directly to the IDP system.
-Symmetric predicates are predicates for which the order of the terms does not matter.
-Every example contains a domain consisting of a set of constants and an exhaustive list of the relations (predicates) that hold on these constants.
-If optimization criteria are to be learned, a set of pairwise over the given examples is also provided.
-Rankings are used instead of absolute scores since rankings only concern few examples at a time.
-Because rankings only ask for relative positioning of a few examples, they are easier to provide.
+The input consists of global definitions that specify the types and predicates that are used to describe examples and a set of examples \sym{E}.
+Typing information relates directly to the problem domain and is usually easy to provide.
+Including types also improves the accuracy and efficiency of the learned clauses.
 
-\begin{example}
+Examples are interpretations, they specify the constants in their domain and then all predicates that are true.
+Predicates can also be generated by background knowledge, in which case they are not included in the example.
+
+When learning optimization criteria, preference information is also provided by the user.
+Each preference is of the form $e_1 \succ e_2 \succ ... \succ e_n$ with $e_i \in \sym{E}$.
+Preference look at the relative position of a few examples at a time and are typically easier to provide than absolute scores for all examples.
+
+\begin{example} "Rewrite"
   \label{ex:moving-input}
@@
   A model could, for example, use a type \textit{Area} and predicate definitions low\_crime(\textit{Area}), cheap(\textit{Area}), school\_in(\textit{Area}), work\_in(\textit{Area}) and live\_in(\textit{Area}).
-  Examples would that specify various areas $A_i$ and describe the conditions (e.g.
+  Examples would then specify various areas $A_i$ and describe the conditions (e.g.
 cheap($A_1$), low\_crime($A_2$)) as well as the users choices (e.g.
@@
   \begin{algorithmic}
-  \State $Q \gets \{\square\}$, $\sym{T} \gets \{\}$
-  \While{$\lnot isempty(Q)$}
-    \State $c \gets next(Q)$
-    \If{$covers(c, \textit{Examples})$}
-      \If{$\lnot entails(\sym{T}, c)$}
+  \State Given: Examples \sym{E} and threshold $t$
+  \State $\sym{Q} \gets \{\square\}$, $\sym{T} \gets \{\}$
+  \While{$\#\sym{Q} > 0$}
+    \State $c \gets next(\sym{Q})$
+    \If{$\#\{e \in \sym{E} | e \models c\} \geq t$}
+      \If{$\lnot (\sym{T} \models c)$}
         \State $\sym{T} = \sym{T} \cup c$
@@
     \Else
-      \State $Q \gets Q \cup \rho(c)$
+      \State $\sym{Q} \gets \sym{Q} \cup \rho(c)$
     \EndIf
@@
 The clause learning system is based on the clausal discovery algorithm (alg.~\ref{alg:cd}).
-The logical system IDP is used for two essential functions: $\mathit{covers}$ and $\mathit{entails}$.
-The $\mathit{covers}$ function calculates which examples are covered by a clause.
-The $\mathit{entails}$ function calculates whether an accepted clause is logically entailed by background knowledge or clauses that are already in the result set.
+Note the double use of the $\models$ operator.
+On the one hand, it is used in $e \models c$ to denote that clause~$c$ is true in the example (interpretation)~$e$ ($c$ covers $e$).
+This test is used to compute if a clause covers enough examples.
+On the other hand, in $\sym{T} \models c$ it signifies that \sym{T} entails $c$ (i.e. for every interpretation it holds that if \sym{T} is true, $c$ is also true).
+The algorithm uses this test to remove redundant clauses.
+For both cases the IDP knowledge base system \cite{de2013prototype,wittocx2008idp} is used to compute $\models$.
+The algorithm can find soft constraints ($t < \# \sym{E}$) or hard constraints ($t = \# \sym{E}$).
 
-Starting with the empty clause, clauses are refined until they cover enough examples.
-If a clause covers enough examples it is added to result set, provided it is not entailed by background knowledge or other clauses in the result set.
-To avoid extensive calls to $\mathit{covers}$ and $\mathit{entails}$, two additional tests are used.
+Starting with the empty clause ($\square$), clauses are refined until they cover enough examples.
+If a clause covers enough examples it is added to result set, provided it is not entailed by clauses in the result set.
+In the refinement step, a clause $c = \{l_1, l_2, ..., l_k\}$ is extended by adding a new literal $l$.
+The new clause, $c' = c \cup \{l\}$ is more than general than $c$ ($c' \models c$) and potentially covers more examples.
+By adding just one literal the refinement operator $\rho$ generates a set of clauses that are minimally more general.
+In order to do this efficiently, a list of atoms is pre-calculated in advance, given a maximal amount of variables that can occur within a clause.
+Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.
+The implementation natively supports symmetric predicates for which the order of the terms do not matter.
+Object Identity is used to specify that variables with different names must denote different objects.
+Additionally there are two syntactic restrictions.
+Clauses must be connected (i.e. new atoms must always contain a variable that has already occurred) and range-restricted (i.e. no new variables may be introduced in the head of the clause).
+
+To avoid extensive coverage ($e \models c$) and entailment ($\sym{T} \models c$) calculations, two additional tests are used.
 The subset-test rejects clauses that are supersets of a clause that covers the same or all examples and has been accepted earlier.
@@
 
-In the refinement step, clauses are extended to be more general and potentially cover more examples.
-The refinement operator uses a list of atoms that may be added to a clause.
-This list is calculated in advance, given a maximal amount of variables.
-Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.
-Object Identity is used to specify that variables with different names must denote different objects.
-Additionally there are two syntactic restrictions.
-Clauses must be connected (i.e. new atoms must always contain a variable that has already occurred) and range-restricted (i.e. no new variables may be introduced in the head of the clause).
+% In the refinement step, clauses are extended to be more general and potentially cover more examples.
+% The refinement operator uses a list of atoms that may be added to a clause.
+% This list is calculated in advance, given a maximal amount of variables.
 
@@
 
-\begin{example}
+\begin{example} "Rewrite"
   Consider examples~$e_1, e_2, e_3$, rankings $e_1 > e_2, e_2 > e_3$ and clauses $c_1, c_2, c_3$.
@@
 
-The \svm{} software was chosen to find the scoring function since it uses a linear model and offers an efficient implementation.
+\svm{} \cite{joachims2006training} was chosen to find the scoring function since it uses a linear model and offers an efficient implementation.
 The weights that it assigns to the features can be used directly as weights for the optimization criteria.
@@
 
-\begin{question}[Threshold]
-  What is the effect of the soft constraint threshold on the accuracy?
-\end{question}
-  \begin{table}
+  \begin{table}[!htp]
     \caption{Scores for different thresholds ($t$)}
@@
 
+\begin{question}[Threshold]
+  What is the effect of the soft constraint threshold on the accuracy?
+\end{question}
+
 Table~\ref{tbl:limiet} demonstrates that increasing the threshold used for finding soft constraints does not improve the score.
@@
 The research in this thesis has focused on automatically acquiring constraints and optimization criteria from examples and rankings.
-Implementation for both clause learning and clausal optimization have been provided that accomplish these tasks using first order logic clauses.
+Implementation for both clause learning and clausal optimization have been provided that accomplish these tasks using first-order logic clauses.
 
@@
 The author thanks his promoters Dr. Luc De Raedt en Dr. ir. Anton Dries.
-Furthermore, he expresses his gratitude for the help provided by Bart Bogaerts, Dr. Jesse Davis, Dr. Marc Denecker and Vladimir Dzyuba.
+Samuel Kolb is supported by the Research Foundation - Flanders (FWO).
 

### 2015-10-23 Paper.tex
Overhauled example

@@
 In order to reason about clauses, the standard notions of (function free) first-order logic are described.
-An atom $p(t_1, t_2, ..., t_n)$ consists of a predicate symbol $p$ that is applied to $n$ terms $t_i$.
+An atom $p(t_1, t_2, ..., t_n)$ consists of a predicate symbol~$p$ that is applied to $n$ terms~$t_i$.
 Terms are either constants or variables.
@@
 \paragraph{Learning constraints}
-Constraint learning is a hard problem.
+Constraint learning attempts to automatically identify constraints in one or multiple examples.
 The search space of constraints is typically very large and, unlike in typical machine learning problems, there are usually only few examples to learn from.
 
-Different systems attempt to learn constraints using different methods.
+There are different approaches to accomplishing this task.
 The problem of having little examples can be alleviated by interactively generating data.
@@
 ModelSeeker \cite{Beldiceanu:ModelSeeker} attempts to structure the problem variables in different ways.
-The system then tests for different global constraints whether they hold within this structure.
+The system then uses a large catalog of global constraints to to find the ones that hold within this structure. 
 This approach allows for constraint learning from a few examples.
@@
 The weighted clauses \sym{W} are tuples $\mathbb{R} \times \sym{C}$ that can be used to calculate a score for any example $e$:
-\begin{eqnarray*}
-  score(e) = \sum\limits_{(\mathit{w}, \mathit{c}) \in \sym{W}} \mathit{w} \cdot \mathbf{1}_{e \models c}
-\end{eqnarray*}
+\begin{eqnarray}
+  \label{eqn:score}
+  score(e, \sym{W}) = \sum\limits_{(\mathit{w}, \mathit{c}) \in \sym{W}} \mathit{w} \cdot \mathbf{1}_{e \models c}
+\end{eqnarray}
 
-\begin{example}[Moving] "Rewrite"
+\begin{example}[Moving]
   \label{ex:moving}
-  Consider the scenario moving to a new city.
-  Predicates $\mathit{live\_in}, \mathit{work\_in}$ and $\mathit{school\_in}$ indicate the choice of housing, job and school.
-  There might be different areas in the city which
+  Consider the scenario of moving to a new city and choosing areas for housing, work and school.
+  The choices are captured by the predicates $\mathit{live\_in}$, $\mathit{work\_in}$ and $\mathit{school\_in}$.
+  Additionally, the predicates $\mathit{cheap}$ and $\mathit{low\_crime}$ are used to designate cheap and low crime areas, respectively.
 
-  The user has to choose a job, a place to live and a school.
-  There are multiple job offers, houses and schools available across different areas of the city.
-  User provided examples include a choice of where to work, where to live and what school is chosen.
-  By comparing and ranking some of the examples, the user can express his preferences.
-  Weighted soft-constraints can now be learned and used to identify the best choice.
-  In this case, a possible soft constraint could be that the school and work place are in the same area.
-  The weights that have been learned indicate how desirable or how undesirable a constraint is.
-  Since the constraints are domain independent they can be used for any (unseen) city.
+  Assume the first-order weighted MAX-SAT model \sym{M} describes the optimization problem to be solved:
+  \begin{flalign*}
+    (w_1, c_1) &= (0.50, \mathit{low\_crime}(a) \leftarrow \mathit{live\_in}(a)) \\
+    (w_2, c_2) &= (0.25, \mathit{school\_in}(a) \leftarrow \mathit{work\_in}(a)) \\
+    (w_3, c_3) &= (1.00, \mathit{low\_crime}(a) \leftarrow \mathit{school\_in}(a)) \\
+    (w_4, c_4) &= (-1.00, \mathit{false} \leftarrow \mathit{live\_in}(a) \land \mathit{cheap}(a))
+  \end{flalign*}
+  Let examples $e_1$, $e_2$ and~$e_3$ be interpretations that reason over the same city.
+  Specifically, let all examples share constants (areas) $A_1$, $A_2$ and $A_3$ and ground atoms $\mathit{cheap}(A_1)$, $\mathit{cheap}(A_3)$, $\mathit{low\_crime}(A_1)$ and $\mathit{low\_crime}(A_2)$.
+Additionally, they specify:
+  \begin{flalign*}
+    &e_1: \mathit{live\_in}(A_1), \mathit{work\_in}(A_2), \mathit{school\_in(A_2)} \\
+    &e_2: \mathit{live\_in}(A_2), \mathit{work\_in}(A_2), \mathit{school\_in(A_3)} \\
+    &e_3: \mathit{live\_in}(A_2), \mathit{work\_in}(A_2), \mathit{school\_in(A_1)}
+  \end{flalign*}
+
+  The score of an example for the model~\sym{M} can be calculated using (\ref{eqn:score}).
+  Example~$e_1$ fulfills clauses $c_1$, $c_2$ and~$c_3$, therefore, $score(e_1, \sym{M}) = 0.5 + 0.25 + 1 + 0 = 1.75$, $score(e_2, \sym{M}) = -1$ and $score(e_3, \sym{M}) = 0$.
+  Therefore, the model \sym{M} would prefer $e_1$ over~$e_2$ ($e_1 \succ e_2$) and $e_2$ over~$e_3$ ($e_2 \succ e_3$).
+  The aim of our research is to learn a model like \sym{M} given such examples and preferences.
 \end{example}
@@
 Note the double use of the $\models$ operator.
-On the one hand, it is used in $e \models c$ to denote that clause~$c$ is true in the example (interpretation)~$e$ ($c$ covers $e$).
+On the one hand, it is used in $e \models c$ to denote that clause~$c$ is true in the example (interpretation)~$e$ and $c$ is said to cover $e$.
 This test is used to compute if a clause covers enough examples.
@@
 \begin{example} "Rewrite"
-  Consider examples~$e_1, e_2, e_3$, rankings $e_1 > e_2, e_2 > e_3$ and clauses $c_1, c_2, c_3$.
+  Consider examples~$e_1, e_2, e_3$, rankings $e_1 \succ e_2, e_2 \succ e_3$ and clauses $c_1, c_2, c_3$.
   If the clauses cover examples according to table~\ref{tbl:cover_examples}, then, for example, the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.

### 2015-10-23 Paper.tex
Import Anton introduction

@@
 \begin{abstract}
-% This paper presents a system that is able to automatically learn both constraints from examples and optimization criteria from rankings.
-% Both hard and soft constraints can be learned and constraints are represented as first-order logical clauses.
-% Choosing such clauses as representations enables the use of ILP techniques such as clausal discovery.
-% In this research, optimization criteria are represented as weighted clauses.
-% For all examples, the clause learning system could find the essential constraints efficiently.
-% The learning system for optimization criteria is shown to achieve accurate results and can learn criteria that identify the correct optimal solution even for few examples and noisy rankings.
-% Finally, a procedure is outlined that allows weighted first-order logic clauses to be used for optimization in practice.
-
 While there exist several approaches in the constraint programming community to learn a constraint theory, few of them have considered the learning of constraint optimization problems.
@@
 In order to learn these weighted clauses, the clausal optimization system uses examples of possible worlds and a set of preferences that state what examples are preferred over others.
-% "Change" The examples take the form of rankings that state that one possible world is preferred over another one.
 The technique is also empirically evaluated on a number of examples.
@@
 \section{Introduction}
-% This paper introduces an approach for learning constraints and optimization criteria.
-% Using constraints and optimization criteria such as the ones proposed in this paper allows problems to be modeled \emph{declaratively}.
-% This means that the representation is decoupled from the process of finding a (optimal) solution for the problem.
-% However, non-experts often experience difficulties when modeling a problem using constraints \cite{Wallace:PrinciplesCP}.
-% The first goal of this research is to automate the step of modeling a problem, in order to, amongst others, make constraint solving more accessible.
-% In this approach a few examples of what a user considers valid solutions are used by a computer system to find constraints.
-% These constraints can then be used by an existing constraint solver.
-
-% Constraint learning attempts to find constraints automatically given examples.
-% By using first-order logical clauses as representation for constraints, techniques from Inductive Logic Programming (ILP) can be used for learning.
-% Hereby, this research follows the same approach as \cite{Lallouet:LearningCP}, exploiting parallels between ILP and constraint learning.
-% In this case, the clausal discovery algorithm \cite{DeRaedt:ClausalDiscovery} is used to find constraints.
-
-% There has been increasing interest in formulating declarative models for constraint satisfaction as well as optimization problems.
-% By identifying soft constraints that are only satisfied by some examples and assigning weights to them based on their importance, an optimal solution can be sought.
-% The use of weighted soft constraints to model user preferences is inspired, amongst others, by the research of \cite{campigotto2011active}.
-% Their work examines attempts to obtain similar results in an interactive context for propositional constraints.
-% In this research, first-order constraints are used and optimization criteria are modeled using a first-order version of weighted MAX-SAT.
-% This paper also outlines how these constraints can be used in an existing solver to obtain an optimal solution in practice.
-
-% Since the underlying goal of this research is to facilitate the generation of a formal representation, the input format aims to be easy to use.
-% It has been a key focus to make reasonable assumptions concerning the knowledge a user can provide.
-% One important difference compared to similar research (including \cite{campigotto2011active}) is that users provide partial rankings over examples instead of absolute scores.
-
-% This paper is based on research that has been conducted during the course of my masters thesis~\cite{kolb2015thesis}.
-% The key contribution presented in this paper is a novel approach for learning weighted first-order clauses from rankings to capture user preferences.
-% In this context an implementation of clausal discovery has been realized that supports soft constraints.
-% Moreover, an approach to obtaining optimal solutions for the weighted first-order clauses is presented.
-
 An important class of declarative models is concerned with (discrete) constrained optimization problems.
@@
 Furthermore, \cite{campigotto2011active}) have learned propositional MAX-SAT theories from examples together with their target value.
-%"Expand something here yet?" ...%Each example ....
-Also, work in statistical relational learning (SRL) on learning the structure of Markov Logic networks (MLN) or other SRL models is relevant, as MAP inference can be used to query for the most likely state.
+Also, work in statistical relational learning (SRL) on learning the structure of Markov Logic networks (MLN) or other SRL models~\anton{[reference]} is relevant, as MAP inference can be used to query for the most likely state.
 
 We use weighted first-order logical theories to represent constraint optimization problems.
-The weights will then capture the optimization function, and (first-order) weighted MAX-SAT can be used to determine the \emph{best} solutions.
-Furthermore, we shall learn such models by combining inductive logic programming (ILP) principles (such as clausal discovery \cite{DeRaedt:ClausalDiscovery}) with preference learning.
-ILP will be used to learn the clausal theories, and preference learning to determine the weights.
-Therefore, users will provide examples of possible worlds and a set of preferences that specify which examples are preferred over other ones.
-% "Unify usage of examples / preferences" Therefore, the examples will take the form of rankings, where each ranking specifies that one particular possible world is preferred over another one.
+The weights capture the optimization function, and (first-order) weighted MAX-SAT can be used to determine the \emph{best} solutions.
+Furthermore, we learn such models by combining inductive logic programming (ILP) principles (such as clausal discovery \cite{DeRaedt:ClausalDiscovery}) with preference learning.
+We use ILP to learn the clausal theories, and preference learning to determine the weights.
+Therefore, users provide examples of possible worlds and a set of preferences that specify which examples are preferred over other ones.
 This setting extends traditional constraint learning and ILP approaches with the ability to learn soft constraint and to integrate constraint satisfaction and optimization.
-It extends the approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than from examples with target values.
-Moreover, it differs the SRL and MLN approaches in that one learns from preferences rather than from examples.
- % and rankings are used rather than possible worlds + the target value of the examples, and differs from the SRL and MLN approaches in that one learns from rankings rather than from examples.
+It extends the approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than from examples with target values.\anton{can't parse}
+Moreover, it differs from the SRL and MLN approaches in that one learns from preferences rather than from examples.
 
 This paper is organized as follows.
-Section~2 will explain some background knowledge and related work.
-In Section~3 we shall introduce the formalism and the setting and in Section~4 the algorithm and the clausal optimization system to solve this problem.
-In Section~5 we shall report on some experiments, and finally, in Section~6, we conclude.
-% "Rewrite" In Section 2, we shall introduce the formalism and the setting,
-% "Rewrite" in Section 3, the algorithm and system X to solve this problem, in Section 4, 
-% "Rewrite" we shall report on some experiments, and Finally, in Section 5 ,we conclude
+Section~2 explain some background knowledge and related work.
+In Section~3 we introduce the formalism and the setting, in Section~4 the algorithm and the clausal optimization system to solve this problem, and in Section~5 we report on some experiments.
+Finally, Section~6 concludes this paper.
 

### 2015-10-23 Paper.tex
Import Anton Background

@@
 \section{Background}
-In this section a brief overview of the important concepts and relevant literature will be provided.
+In this section we provide a brief overview of the important concepts and relevant literature.
 
@@
 In order to reason about clauses, the standard notions of (function free) first-order logic are described.
-An atom $p(t_1, t_2, ..., t_n)$ consists of a predicate symbol~$p$ that is applied to $n$ terms~$t_i$.
+An atom $p(t_1, t_2, ..., t_n)$ consists of a predicate symbol $p$ that is applied to $n$ terms $t_i$.
 Terms are either constants or variables.
@@
 
-Related work \cite{campigotto2011active} has creates a weighted MAX-SAT problem by learning propositional clauses and weights using examples and absolute scores.
+In related work \cite{campigotto2011active} a weighted MAX-SAT problem is created by learning propositional clauses and weights using examples and absolute scores.
 Inspired by their results, our research tries to learn first-order logic clauses and weights using examples and rankings.
@@
 ModelSeeker \cite{Beldiceanu:ModelSeeker} attempts to structure the problem variables in different ways.
-The system then uses a large catalog of global constraints to to find the ones that hold within this structure. 
+The system then uses a large catalog of global constraints to find the ones that hold within this structure. 
 This approach allows for constraint learning from a few examples.
 
-As mentioned earlier, the research of \cite{Lallouet:LearningCP} already tried to use ILP techniques to learn constraints.
+As mentioned earlier, the research of \cite{Lallouet:LearningCP} already tried to use ILP techniques to learn constraints.\anton{"tried to" sounds a bit negative}
 In their approach, the search space of constraints is explored in a bidirectional manner, primarily using negative examples.
@@
 The result set consists of the most specific clauses covering all examples.
-
-% \paragraph{Logic Solver} " v ------ v   Dissolve"
-% The logic system IDP \cite{de2013prototype,wittocx2008idp} is used to calculate whether examples satisfy clauses and is used as a theorem prover.
-% It supports an extension of first-order logic ($FO(\cdot)$) and can also be used to generate new solutions.
-% Background knowledge is given as an IDP theory.
 

### 2015-10-23 Paper.tex
Import own background change
Import Anton changes approach

@@
 \section{Approach}
-In this research, both constraint learning and learning optimization criteria were implemented.
-A two-step approach is used to learn optimization criteria.
-First, clauses learning is used to find soft constraints in the given examples.
+% In this research, both constraint learning and learning optimization criteria were implemented.
+In this work, we use a two-step approach to learn optimization criteria.
+First, clause learning is used to find soft constraints in the given examples.
 Then, user-provided preference information is used to learn weights for these constraints.
@@
 The input consists of global definitions that specify the types and predicates that are used to describe examples and a set of examples \sym{E}.
-Typing information relates directly to the problem domain and is usually easy to provide.
+\anton{Explain: what is type information?}
+Type information relates directly to the problem domain and is usually easy to provide.
 Including types also improves the accuracy and efficiency of the learned clauses.
 
-Examples are interpretations, they specify the constants in their domain and then all predicates that are true.
+Examples are interpretations, they specify the constants in their domain and all predicates that are true.
 Predicates can also be generated by background knowledge, in which case they are not included in the example.
@@
 Each preference is of the form $e_1 \succ e_2 \succ ... \succ e_n$ with $e_i \in \sym{E}$.
-Preference look at the relative position of a few examples at a time and are typically easier to provide than absolute scores for all examples.
+Preferences describe the relative position of a few examples at a time and are typically easier to provide than absolute scores for all examples.
 
@@
 The clause learning system is based on the clausal discovery algorithm (alg.~\ref{alg:cd}).
-Note the double use of the $\models$ operator.
+Note the double use of the $\models$ operator. \anton{This paragraph seems strange.}
 On the one hand, it is used in $e \models c$ to denote that clause~$c$ is true in the example (interpretation)~$e$ and $c$ is said to cover $e$.
@@
 Starting with the empty clause ($\square$), clauses are refined until they cover enough examples.
-If a clause covers enough examples it is added to result set, provided it is not entailed by clauses in the result set.
+If a clause covers enough examples it is added to the result set, provided it is not entailed by other clauses already in the result set.
 In the refinement step, a clause $c = \{l_1, l_2, ..., l_k\}$ is extended by adding a new literal $l$.
-The new clause, $c' = c \cup \{l\}$ is more than general than $c$ ($c' \models c$) and potentially covers more examples.
-By adding just one literal the refinement operator $\rho$ generates a set of clauses that are minimally more general.
-In order to do this efficiently, a list of atoms is pre-calculated in advance, given a maximal amount of variables that can occur within a clause.
+The new clause, $c' = c \cup \{l\}$ is more general than $c$ ($c' \models c$) and potentially covers more examples.
+By adding just one literal the refinement operator $\rho$ generates the set of clauses that are minimally more general.
+In order to do this efficiently, a list of atoms is pre-calculated, given a maximal amount of variables that can occur within a clause.
 Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.
@@
 Clauses must be connected (i.e. new atoms must always contain a variable that has already occurred) and range-restricted (i.e. no new variables may be introduced in the head of the clause).
+\anton{This is a vague list of unconnected statements.}
 
@@
 Because clauses can be formulated in multiple ways, a representative-test is used to remove clauses that are not in canonical form.
+\anton{Not clear, needs more explanation, or should be removed.}
 
@@
 Therefore, every example~$e$ can be translated to a vector of boolean features by introducing a feature~$f_i$ for every clause~$c_i$, with $f_i = \mathbf{1}_{c_i \text{ covers } e}$.
+\anton{What does this mean?}
 Existing software can be used to learn a linear scoring function $\sum_i w_i \cdot f_i$ over these features based on the given rankings.
@@
 \begin{example} "Rewrite"
-  Consider examples~$e_1, e_2, e_3$, rankings $e_1 \succ e_2, e_2 \succ e_3$ and clauses $c_1, c_2, c_3$.
+  Consider examples~$e_1, e_2, e_3$, rankings $e_1 > e_2, e_2 > e_3$ and clauses $c_1, c_2, c_3$.
   If the clauses cover examples according to table~\ref{tbl:cover_examples}, then, for example, the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.
@@
 Using $t$ and $\mathit{cost}$, a function $\mathit{actual}(i)$ is then defined in IDP as $\mathbf{1}_{\lnot t(i)} \cdot \mathit{cost}(i)$.
-This function is used in the term $\sum_i actual(i)$ to be minimized, which will allow IDP to search for an optimal solution.
+This function is used in the optimization criterion $\sum_i actual(i)$ which should be minimized. % , which will allow IDP to search for an optimal solution.
 

### 2015-10-23 Paper.tex
Import own changes approach
Multiple fixes

@@
 
+\newcommand{\anton}[1]{{\color{green}(#1)}}
+\newcommand{\sam}[1]{{\color{red}(#1)}}
+
 \setlength{\tabcolsep}{6pt}
@@
 It employs inductive logic programming techniques to learn a set of first-order clauses and then uses preference learning techniques to learn the weights of the clauses.
-In order to learn these weighted clauses, the clausal optimization system uses examples of possible worlds and a set of preferences that state what examples are preferred over others.
+In order to learn these weighted clauses, the clausal optimization system uses examples of possible worlds and a set of preferences that state which examples are preferred over other ones.
 The technique is also empirically evaluated on a number of examples.
@@
 These combine constraint satisfaction problems with an optimization function that specifies which solutions are optimal.
-While these models are powerful and elegant, they are often also hard to obtain.
-\cite{Wallace:PrinciplesCP} shows that the task of modeling constraint satisfaction problems is difficult, certainly for non-experts.
+While these models are powerful and elegant, they are often also hard to obtain, certainly for non-experts \cite{Wallace:PrinciplesCP}.
 This has motivated several researchers to investigate whether such models can be learned.
@@
 Recent efforts \cite{Lallouet:LearningCP} have attempted to combine these two fields.
-Furthermore, \cite{campigotto2011active}) have learned propositional MAX-SAT theories from examples together with their target value.
-Also, work in statistical relational learning (SRL) on learning the structure of Markov Logic networks (MLN) or other SRL models~\anton{[reference]} is relevant, as MAP inference can be used to query for the most likely state.
+Furthermore, \cite{campigotto2011active}) have learned propositional weighted MAX-SAT theories from examples.
+Also, work in statistical relational learning (SRL) on learning the structure of Markov Logic networks (MLN) or other SRL models \cite{kok2005learning} is relevant, as MAP inference can be used to query for the most likely state.
 
@@
 Therefore, users provide examples of possible worlds and a set of preferences that specify which examples are preferred over other ones.
-This setting extends traditional constraint learning and ILP approaches with the ability to learn soft constraint and to integrate constraint satisfaction and optimization.
-It extends the approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than from examples with target values.\anton{can't parse}
+This setting extends traditional constraint learning and ILP approaches with the ability to learn soft constraints and integrates constraint satisfaction and optimization.
+It also extends the approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than examples with target values.
 Moreover, it differs from the SRL and MLN approaches in that one learns from preferences rather than from examples.
@@
 This paper is organized as follows.
-Section~2 explain some background knowledge and related work.
+Section~2 reviews background and related work.
 In Section~3 we introduce the formalism and the setting, in Section~4 the algorithm and the clausal optimization system to solve this problem, and in Section~5 we report on some experiments.
@@
 \paragraph{Clauses}
-In order to reason about clauses, the standard notions of (function free) first-order logic are described.
+We use (function free) first-order clauses; which we now introduce.
 An atom $p(t_1, t_2, ..., t_n)$ consists of a predicate symbol~$p$ that is applied to $n$ terms~$t_i$.
@@
 Hereby, positive literals are grouped in the head and negative literals are grouped in the body.
-Sometimes clauses are also written as a set of literals $\{l_1, l_2, ..., l_k\}$.
+Sometimes clauses are also written as sets of literals $\{l_1, l_2, ..., l_k\}$.
 
-Atoms are grounded if they only contain constants.
+Atoms are grounded if they contain no variables.
 A Herbrand interpretation is a set of ground atoms.
@@
 The Boolean Satisfiability Problem (SAT) attempts to determine whether a propositional formula can be satisfied.
-Formulas can be rewritten into a special form called the Conjunctive Normal Form (CNF), which consists of a conjunction of clauses (disjunctions).
-An extension of this problem is the Maximum Satisfiability Problem (MAX-SAT), which is an optimization problem and attempts to maximize the amount of satisfied clauses.
+Formulas can be rewritten into Conjunctive Normal Form (CNF), which consists of a conjunction of clauses (disjunctions).
+An extension of this problem is the Maximum Satisfiability Problem (MAX-SAT), which is an optimization problem and attempts to maximize the number of satisfied clauses.
 By assigning positive integer weights to clauses and maximizing the sum of the weights of satisfied clauses, one obtains the weighted MAX-SAT problem.
 
-In related work \cite{campigotto2011active} a weighted MAX-SAT problem is created by learning propositional clauses and weights using examples and absolute scores.
+In related work \cite{campigotto2011active} weighted MAX-SAT problems are induced by learning propositional clauses and weights using examples and absolute scores.
+\sam{Change this section}
 Inspired by their results, our research tries to learn first-order logic clauses and weights using examples and rankings.
@@
 Constraint learning attempts to automatically identify constraints in one or multiple examples.
-The search space of constraints is typically very large and, unlike in typical machine learning problems, there are usually only few examples to learn from.
+This task is hard because the search space of constraints is typically very large and, unlike in typical machine learning problems, there are usually only few examples to learn from.
 
@@
 
-ModelSeeker \cite{Beldiceanu:ModelSeeker} attempts to structure the problem variables in different ways.
+ModelSeeker \cite{Beldiceanu:ModelSeeker} structures the problem variables in different ways.
 The system then uses a large catalog of global constraints to find the ones that hold within this structure. 
@@
 
-As mentioned earlier, the research of \cite{Lallouet:LearningCP} already tried to use ILP techniques to learn constraints.\anton{"tried to" sounds a bit negative}
+As mentioned earlier, the research of \cite{Lallouet:LearningCP} used ILP techniques to learn constraints.
 In their approach, the search space of constraints is explored in a bidirectional manner, primarily using negative examples.
 
-\paragraph{Clausal Discovery}
+\paragraph{Clausal Discovery} \sam{Dissolve}
 The clausal discovery algorithm \cite{DeRaedt:ClausalDiscovery} attempts to learn first-order logic clauses from positive examples.
@@
 \section{Problem Statement}
-The research presented in this paper aims to learn constraints and optimization criteria based on positive examples and user preferences.
-Since it aims to simplify the process of generating formal representations it should be easy to use and require input that a user can reasonably be expected to provide.
+\label{sec:problem}
+We want to learn constraints and optimization criteria based on positive examples and user preferences.
+The goal of this task is to simplify the process of generating formal representations, therefore, any input should be reasonably easy for a user to provide.
 
@@
 The learned clauses are domain-independent and contain only variables.
-
-% \begin{example} " TODO - necessary?"
-%   Consider sudoku as an example of a problem where hard constraints are to be found.
-%   Given one or more solved sudokus, the learning system has to identify the rules of the game.
-%   Since the learned clauses only contain universally quantified variables, for example $4 \times 4$ sudokus could be used to find constraints.
-%   The learned constraints are independent of the size of the sudoku and can also be used to solve $9 \times 9$ sudokus, for example.
-
-%   In some problems, one is interested in regularities that  often but not always occur.
-%   An example would be regularities in certain types of buildings.
-%   While certain regularities would occur in many buildings, there are often exceptions and different styles.
-%   In this case, one could attempt to find constraints on the structure of the floor plans that are satisfied by a given amount of examples.
-% \end{example}
 
@@
 
-  Assume the first-order weighted MAX-SAT model \sym{M} describes the optimization problem to be solved:
+  Assume the following first-order weighted MAX-SAT model \sym{M} describes the optimization problem to solve:
   \begin{flalign*}
@@
   The score of an example for the model~\sym{M} can be calculated using (\ref{eqn:score}).
-  Example~$e_1$ fulfills clauses $c_1$, $c_2$ and~$c_3$, therefore, $score(e_1, \sym{M}) = 0.5 + 0.25 + 1 + 0 = 1.75$, $score(e_2, \sym{M}) = -1$ and $score(e_3, \sym{M}) = 0$.
-  Therefore, the model \sym{M} would prefer $e_1$ over~$e_2$ ($e_1 \succ e_2$) and $e_2$ over~$e_3$ ($e_2 \succ e_3$).
-  The aim of our research is to learn a model like \sym{M} given such examples and preferences.
+  Example~$e_1$ is covered by clauses $c_1$, $c_2$ and~$c_3$, therefore, $score(e_1, \sym{M}) = 0.5 + 0.25 + 1 + 0 = 1.75$.
+  Similarly, $score(e_2, \sym{M}) = -1$ and $score(e_3, \sym{M}) = 0$ can be computer.
+  Based on these scores, the model \sym{M} would prefer $e_1$ over~$e_2$ ($e_1 \succ e_2$), $e_1$ over~$e_3$ ($e_1 \succ e_3$) and $e_2$ over~$e_3$ ($e_2 \succ e_3$).
+  The aim of our research is to learn a model like \sym{M}, given such examples and preferences.
 \end{example}
 
-Aside from learning optimization criteria, this paper presents an approach to use these optimization criteria to find an optimal solution in practice.
+Apart from learning optimization criteria, this paper presents an approach to use these optimization criteria to find an optimal solution in practice.
 
@@
 \section{Approach}
-% In this research, both constraint learning and learning optimization criteria were implemented.
-In this work, we use a two-step approach to learn optimization criteria.
+In this work, we use a two-step approach to learn first-order weighted MAX-SAT.
 First, clause learning is used to find soft constraints in the given examples.
@@
 
-% "Dissolve" The clause learning implementation uses the IDP system to test if clauses satisfy the examples and identify redundant clauses.
-
-\subsection{Input}
+\subsection{Input} \sam{Dissolve, partially or entirely}
 The input consists of global definitions that specify the types and predicates that are used to describe examples and a set of examples \sym{E}.
@@
 
-\begin{example} "Rewrite"
+\begin{example} \sam{Rewrite}
   \label{ex:moving-input}
@@
 \subsection{Clausal Discovery}
-The clause learning system is based on the clausal discovery algorithm (alg.~\ref{alg:cd}).
-Note the double use of the $\models$ operator. \anton{This paragraph seems strange.}
+The clause learning system is based on clausal discovery \cite{DeRaedt:ClausalDiscovery}.
+Algorithm~\ref{alg:cd} illustrates how clausal discovery works on a high level.
+Note the double use of the $\models$ operator.\sam{Review this once yet}
 On the one hand, it is used in $e \models c$ to denote that clause~$c$ is true in the example (interpretation)~$e$ and $c$ is said to cover $e$.
 This test is used to compute if a clause covers enough examples.
-On the other hand, in $\sym{T} \models c$ it signifies that \sym{T} entails $c$ (i.e. for every interpretation it holds that if \sym{T} is true, $c$ is also true).
+On the other hand, in $\sym{T} \models c$ it signifies that \sym{T} logically entails $c$.
 The algorithm uses this test to remove redundant clauses.
 For both cases the IDP knowledge base system \cite{de2013prototype,wittocx2008idp} is used to compute $\models$.
-The algorithm can find soft constraints ($t < \# \sym{E}$) or hard constraints ($t = \# \sym{E}$).
+The algorithm can find both soft constraints ($t < \# \sym{E}$) or hard constraints ($t = \# \sym{E}$).
 
@@
 If a clause covers enough examples it is added to the result set, provided it is not entailed by other clauses already in the result set.
-In the refinement step, a clause $c = \{l_1, l_2, ..., l_k\}$ is extended by adding a new literal $l$.
+In the refinement step, a clause $c = \{l_1, l_2, ..., l_k\}$ is extended (refined) by adding a new literal $l$.
 The new clause, $c' = c \cup \{l\}$ is more general than $c$ ($c' \models c$) and potentially covers more examples.
 By adding just one literal the refinement operator $\rho$ generates the set of clauses that are minimally more general.
-In order to do this efficiently, a list of atoms is pre-calculated, given a maximal amount of variables that can occur within a clause.
+In order to do this efficiently, a list of atoms is pre-calculated, given a maximal number of variables that can occur within a clause.
 Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.
-The implementation natively supports symmetric predicates for which the order of the terms do not matter.
-Object Identity is used to specify that variables with different names must denote different objects.
-Additionally there are two syntactic restrictions.
+
+A few additional measures are used to enable the implementation to be more efficient.
+Most importantly, there are two syntactic restrictions on the form of the clauses.
 Clauses must be connected (i.e. new atoms must always contain a variable that has already occurred) and range-restricted (i.e. no new variables may be introduced in the head of the clause).
-\anton{This is a vague list of unconnected statements.}
+Moreover, Object Identity is used to specify that variables with different names must denote different objects.
+Therefore, a clause~$c$ with variables \sym{V} is extended to the clause~$c_{OI}$ by adding literals $x_i \neq x_j$, for $x_i, x_j \in \sym{V}, i \neq j$.
+Finally, the implementation natively supports symmetric predicates for which the order of the terms do not matter
 
@@
 \subsection{Optimization}
-The first step in finding weighted soft constraints is to identify soft constraints by using the constraint learning system with a (low) threshold.
-Examples are characterized by the soft constraints (clauses) which they satisfy.
-Therefore, every example~$e$ can be translated to a vector of boolean features by introducing a feature~$f_i$ for every clause~$c_i$, with $f_i = \mathbf{1}_{c_i \text{ covers } e}$.
-\anton{What does this mean?}
-Existing software can be used to learn a linear scoring function $\sum_i w_i \cdot f_i$ over these features based on the given rankings.
-For an unseen example, the feature vector is computed by calculating what clauses cover the new example.
-The scoring function can then be used to calculate a score for that example.
+In order to find weighted soft constraints, soft constraints~\sym{C} for the given examples are learned by using the constraint learning system with a (low) threshold.
+The example are then translated to Boolean feature vectors of length $\# \sym{C}$ by introducing a feature for every clause~$c_i \in \sym{C}$.
+For an example~$e$, the $i$-th value of its feature vector $\mathbf{f}_e$ is calculated as $\mathbf{f}_e(i) = \mathbf{1}_{e \models c_i}$.
+Next, existing software is used to learn the linear function $\sum_i w_i \cdot c_i$ over these features, based on user provided preferences.
+This function corresponds with the scoring function (\ref{eqn:score})) introduced in Section~\ref{sec:problem}.
+It can be used to calculate scores for (unseen) examples and by maximizing the score function an optimal solution can be found.
 
-\begin{example} "Rewrite"
-  Consider examples~$e_1, e_2, e_3$, rankings $e_1 > e_2, e_2 > e_3$ and clauses $c_1, c_2, c_3$.
+\begin{example} \sam{Rewrite}
+  \label{ex:features}
+  Consider examples~$e_1, e_2, e_3$, rankings $e_1 \succ e_2, e_2 \succ e_3$ and clauses $c_1, c_2, c_3$.
   If the clauses cover examples according to table~\ref{tbl:cover_examples}, then, for example, the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.
@@
     \toprule
-    $e_1$             & $c_1$               & (1, 0, 0) \\
-    $e_2$             & $c_2$               & (0, 1, 0) \\
-    $e_3$             & $c_1$, $c_2$, $c_3$ & (1, 1, 1) \\
+    $e_1$             & $c_1$               & $\mathbf{f}_{e_1} = (1, 0, 0)$ \\
+    $e_2$             & $c_2$               & $\mathbf{f}_{e_2} = (0, 1, 0)$ \\
+    $e_3$             & $c_1$, $c_2$, $c_3$ & $\mathbf{f}_{e_3} = (1, 1, 1)$ \\
   \end{tabularx}
@@
 
-\svm{} \cite{joachims2006training} was chosen to find the scoring function since it uses a linear model and offers an efficient implementation.
+\svm{} \cite{joachims2006training} was chosen to find the scoring function since it uses a linear model and offers a robust and efficient implementation.
 The weights that it assigns to the features can be used directly as weights for the optimization criteria.
-The input format that is used by \svm{} is also supported by many other learn-to-rank implementations.
-Therefore, other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.
+While this paper uses \svm{} other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.
 
 \subsection{Optimal Solution}
-Solvers, such as IDP, can use clauses directly as hard constraints to generate a solution.
+Solvers, such as IDP \sam{more on IDP, can be used as ASP solver?}, can use clauses directly as hard constraints to generate a solution.
 There is no solver that can use the chosen optimization criteria directly.
@@
 A function $\mathit{cost}(i)$ specifies the cost of not satisfying the clause, which is equal to the weight of the clause.
-\begin{eqnarray*}
-  t(i) \Leftrightarrow \forall v_1, ..., v_n : c_i. \\
-  cost(i) = w_i.
-\end{eqnarray*}
+\begin{flalign*}
+  &t(i) \Leftrightarrow \forall v_1, ..., v_n : c_i. \\
+  &cost(i) = w_i.
+\end{flalign*}
 
-Using $t$ and $\mathit{cost}$, a function $\mathit{actual}(i)$ is then defined in IDP as $\mathbf{1}_{\lnot t(i)} \cdot \mathit{cost}(i)$.
+Using $t$ and $\mathit{cost}$, a function $\mathit{actual}(i) = \mathbf{1}_{\lnot t(i)} \cdot \mathit{cost}(i)$ is then defined inductively.
 This function is used in the optimization criterion $\sum_i actual(i)$ which should be minimized. % , which will allow IDP to search for an optimal solution.
@@
 \section{Evaluation}
-Several experiments aim to measure the accuracy and efficiency of the learning systems for constraints and optimization criteria.
-In order to account for non-determinism, experiments measuring execution times or optimization scores are usually performed eight times and the results are averaged.
-
+In our experiments we aim to measure the accuracy and efficiency of the learning systems for constraints and optimization criteria.
+In order to account for non-determinism, experiments measuring execution times or optimization scores were repeated eight times and their result were averaged.
+\sam{First step evaluation constraint learning, then optimization}
 \subsection{Constraints}
-Four problems have been used to evaluate constraint learning.
-The first problem is map coloring, where countries are assigned colors and neighboring countries may not have the same color.
-Two examples containing each three correctly colored countries are given.
-The second problem is sudoku and a single, solved $4 \times 4$ sudoku is given.
-For the third problem (elevator) three examples have been generated, two of which respect the underlying soft constraint.
-The last problem (co-housing) contains four hard constraints and five examples have been generated that respect the constraints.
+We used four problems to evaluate constraint learning: \sam{clarify hard constraints}
+\begin{description}
+\item[Map Coloring] Countries are assigned colors and neighboring countries may not have the same color.
+Two examples, each containing three correctly colored countries, are given.
+\item [Sudoku] A single, solved $4 \times 4$ sudoku is given.
+\item[Elevator] Three examples have been generated, two of which respect the underlying soft constraint.
+\item[Co-housing] Contains four hard constraints and five examples have been generated that respect the constraints.
+\end{description}
+\anton{more elaborate descriptions of these?}
+\sam{link to thesis}
 
@@
 Often additional constraints were found that describe some structure in the problems.
-For example, for the map coloring problem a learned constraint states that countries are never their own neighbor.
+For example, for the map coloring problem one of the learned constraint states that countries are never their own neighbor.
 These kind of constraints may help a constraint solver to work more efficiently.
@@
 \midrule
-Range             & Map coloring        & 4.629   ($\pm$ 0.199)     \\ % & 2.928                     
+Range \sam{Remove}             & Map coloring        & 4.629   ($\pm$ 0.199)     \\ % & 2.928                     
 restriction       & Sudoku              & 16.118  ($\pm$ 0.154)     \\ % & 3.367                     
@@
 
-\begin{question}[Efficiency]
+\begin{question}[Efficiency] \sam{review}
   How fast is the clause learning system and what is the effect of various design decisions and input on the execution time?
@@
 \end{question}
-Human programmed theories for map coloring and sudoku are available on the website of the IDP system.
+Human programmed theories for map coloring and sudoku are available on the website of the IDP system \footnote{https://dtai.cs.kuleuven.be/software/idp}.
 These theories usually focus on being compact and contain only the essential constraints.
@@
 This time is measured for the learned theories as well as for hand made theories.
-The learned theories are slightly adapted to be able to solve the same problems and corrected if they contain an unfair advantage.
+The learned theories are slightly adapted to be able to solve the same problems and remove unfair advantages such as preprocessed input data.
 
@@
 
-\subsection{Optimization}
+\subsection{Optimization} \sam{Preferences, not rankings}
 The efficiency of the learning system for optimization criteria depends mainly on the efficiency of learning soft constraints.
@@
 The moving problem with input according to example~\ref{ex:moving-input} is used to evaluate the learning of optimization criteria.
-18 possible configurations are used as available examples.
-Four weighted soft constraints are used to represent the underlying model for the users preferences.
-Two approaches are used for evaluation.
-In the first approach, the examples are split into disjoint train and test sets.
-The second approach uses all examples as train as well as test set.
+18 possible configurations are used as available examples~\sym{E}.
+Four weighted soft constraints are used to represent the underlying model~\sym{M} for the users preferences.
+Two approaches are used for evaluation and differ in the way they construct the training ($\sym{E}_{train}$) and test sets ($\sym{E}_{test}$).
+In the first approach, given a parameter~$p_{train}$, the examples~\sym{E} are randomly partitioned into disjoint train and test sets ($\sym{E}_{test}$) such that $\#\sym{E}_{train} = p_{train} \cdot \#\sym{E}$.
+The second approach uses $\sym{E}_{train} = \sym{E}_{test} = \sym{E}$.
 
-For every experiment a fraction of the training examples is selected randomly.
-Pairwise rankings are generated by picking two of the selected examples and predicting the better example using the underlying model.
-The selected examples and a random subset of the possible rankings are then used as input for learning.
-Noise is simulated by flipping a subset of the input rankings.
+Preferences are generated by picking two examples $e_1$,~$e_2 \in \sym{E}_{train}$ and ordering them according to their scores $score(e_i, \sym{M})$.
+Consider all possible pairwise preferences over $\sym{E}_{train}$: $\mathit{pref}(\sym{E}_{train})$.
+Given the fraction $p_{pref}$ of preferences to use, then a random subset~$\sym{P}_{train} \subset \mathit{pref}(\sym{E}_{train})$ of size $p_{pref} \cdot \# \sym{P}$ is used for learning.
+Errors in the input preferences are simulated by flipping preferences (e.g. $e_1 \succ e_2$ becomes $e_1 \prec e_2$).
+Given the fraction $p_{error}$ of errors to introduce, a random subset $\sym{P}_{error} \cup \sym{P}_{train}$ of the preferences are flipped, where $\#\sym{P}_{error} = p_{error} \cdot \sym{P}_{train}$.
 
 To evaluate the learned optimization criteria, all possible pairs of examples in the test set are generated.
-The learned model is used to predict the better example for every pair.
-The score is then calculated as the fraction of correctly predicted pairs. 
-Pairs for which the underlying model ranks both example the same are omitted.
+Both the learned model and the underlying model~$\sym{M}$ are used to predict the better example for every pair.
+The accuracy of the learned model is then calculated as the fraction of correctly predicted pairs. 
+Pairs for which~\sym{M} ranks both examples the same are omitted.
 
@@
 In all cases, more than half the pairs of examples are correctly predicted and high scores can be obtained even for small datasets.
-Additional experiments have shown that even for lower scores the learned optimization criteria are often capable of identifying the correct optimal solution.
+Additional experiments have shown that even for lower scores the learned optimization criteria are often capable of identifying the correct optimal solution. \sam{Refer to thesis}
 
@@
 \section{Conclusion}
-The research in this thesis has focused on automatically acquiring constraints and optimization criteria from examples and rankings.
+We presented an approach to automatically acquire constraints and optimization criteria from examples and preferences.
 Implementation for both clause learning and clausal optimization have been provided that accomplish these tasks using first-order logic clauses.
@@
 The system requires only a minimal amount of information from the user, however, it also allows for the use of expressive background knowledge.
-The learned constraints are domain independent, which facilitates the construction of positive examples.
+The learned constraints are domain independent \sam{review domain}, which facilitates the construction of positive examples.
 
@@
 Optimization criteria can be learned that enable optimal solutions to be found.
-Even for small datasets and noisy rankings, constraints are found that enable most examples to be ranked correctly.
+Even for small datasets and noisy rankings, constraints are found that rank most examples correctly.
 
-Aside from learning formal representations automatically from examples, this research shows how these representations can be used in practice.
+Aside from learning formal representations automatically from examples, we have shown how these representations can be used in practice.
 This also forms an important step to enable the learning of optimization criteria in an interactive setting.
 
-\paragraph{Future work}
-This research offers multiple opportunities for future work.
+\paragraph{Future work} \sam{shorten}
+There are several opportunities for future work.
 It would be interesting to adapt the number of variables and literals in a clause dynamically.
@@
 Learned clauses must be specific enough to not cover any negative examples. 
+\anton{These restrictions are not really described in the paper, so maybe not that interesting to mention here? Or at least add a sentence to introduce them.}
 
@@
 
-As mentioned earlier, the domain independence of the clauses has several advantages.
+As mentioned earlier, the domain independence \sam{review domain} of the clauses has several advantages.
 In some cases, however, specific objects are inherently present in any problem instantiation.
@@
 
-Finally, it would desirable to improve the implementation in order to tackle larger sized problems.
+Finally, it would be desirable to improve the implementation in order to tackle larger sized problems.
 All the experiments were conducted with problems of limited size and the computation time increases rapidly if there are more predicates, variables and literals to be used in clauses.
@@
 \section*{Acknowledgments}
-The author thanks his promoters Dr. Luc De Raedt en Dr. ir. Anton Dries.
+The author thanks his promoters Luc De Raedt en Anton Dries.
 Samuel Kolb is supported by the Research Foundation - Flanders (FWO).

### 2015-10-23 Paper.tex
Trim and changes in thesis

@@
 
-To avoid extensive coverage ($e \models c$) and entailment ($\sym{T} \models c$) calculations, two additional tests are used.
-The subset-test rejects clauses that are supersets of a clause that covers the same or all examples and has been accepted earlier.
-Because clauses can be formulated in multiple ways, a representative-test is used to remove clauses that are not in canonical form.
-\anton{Not clear, needs more explanation, or should be removed.}
-
-% In the refinement step, clauses are extended to be more general and potentially cover more examples.
-% The refinement operator uses a list of atoms that may be added to a clause.
-% This list is calculated in advance, given a maximal amount of variables.
-
 \subsection{Optimization}
@@
 
-The learning process is parameterized by the maximal amount of variables and literals allowed per clause.
+The learning process is parameterized by the maximal number of variables and literals allowed per clause.
 If these limits are too large, constraints are found that over-fit the training data.
@@
 The moving problem with input according to example~\ref{ex:moving-input} is used to evaluate the learning of optimization criteria.
-18 possible configurations are used as available examples~\sym{E}.
-Four weighted soft constraints are used to represent the underlying model~\sym{M} for the users preferences.
+18 possible configurations are used as available examples~(\sym{E}).
+Four weighted soft constraints are used to represent the underlying model~(\sym{M}) for the users preferences.
 Two approaches are used for evaluation and differ in the way they construct the training ($\sym{E}_{train}$) and test sets ($\sym{E}_{test}$).
-In the first approach, given a parameter~$p_{train}$, the examples~\sym{E} are randomly partitioned into disjoint train and test sets ($\sym{E}_{test}$) such that $\#\sym{E}_{train} = p_{train} \cdot \#\sym{E}$.
+In the first approach, given a parameter~$p_{train}$, the examples are randomly partitioned into disjoint train and test sets ($\sym{E}_{test}$), such that $\#\sym{E}_{train} = p_{train} \cdot \#\sym{E}$.
 The second approach uses $\sym{E}_{train} = \sym{E}_{test} = \sym{E}$.
 
-Preferences are generated by picking two examples $e_1$,~$e_2 \in \sym{E}_{train}$ and ordering them according to their scores $score(e_i, \sym{M})$.
+Preferences are generated by picking two examples $e_1$,~$e_2 \in \sym{E}_{train}$ and ordering them according to their score: $score(e_i, \sym{M})$.
 Consider all possible pairwise preferences over $\sym{E}_{train}$: $\mathit{pref}(\sym{E}_{train})$.
@@
 
-Figure~\ref{fig:fractie} shows how the scores improve as the amount of examples and the fraction of included preferences increases.
+Figure~\ref{fig:fractie} shows how the scores improve as the number of examples and the fraction of included preferences increases.
 In all cases, more than half the pairs of examples are correctly predicted and high scores can be obtained even for small datasets.

### 2015-10-23 Paper.tex
Title caps & more formal optimization experiments

@@
 
-\paragraph{Learning constraints}
-Constraint learning attempts to automatically identify constraints in one or multiple examples.
-This task is hard because the search space of constraints is typically very large and, unlike in typical machine learning problems, there are usually only few examples to learn from.
+\paragraph{Learning Constraints}
+The goal of constraint learning is to automatically identify constraints in one or multiple examples.
+This task is difficult because the search space of constraints is typically very large and, unlike in typical machine learning problems, there are usually only few examples to learn from.
 
@@
 
-\subsection{Constraint learning}
+\subsection{Constraint Learning}
 \emph{Given a set of examples and a limit $t$, the goal is to find maximally specific clauses that are satisfied by at least $t$ of the given examples.}
@@
 
-\subsection{Learning optimization criteria}
+\subsection{Learning Optimization Criteria}
 \emph{Given a set of examples \sym{E} and a set of preferences \sym{P} over these examples, the goal is to learn soft constraints and weights such that the order described by these optimization criteria maximally corresponds with~\sym{P}.}
@@
 \end{question}
-Human programmed theories for map coloring and sudoku are available on the website of the IDP system \footnote{https://dtai.cs.kuleuven.be/software/idp}.
+Human programmed theories for map coloring and sudoku are available on the website of the IDP solver\footnote{https://dtai.cs.kuleuven.be/software/idp}.
 These theories usually focus on being compact and contain only the essential constraints.
@@
 Consider all possible pairwise preferences over $\sym{E}_{train}$: $\mathit{pref}(\sym{E}_{train})$.
-Given the fraction $p_{pref}$ of preferences to use, then a random subset~$\sym{P}_{train} \subset \mathit{pref}(\sym{E}_{train})$ of size $p_{pref} \cdot \# \sym{P}$ is used for learning.
+If the fraction of preferences to be used is~$p_{pref}$, then a random subset~$\sym{P}_{train} \subset \mathit{pref}(\sym{E}_{train})$ of size $p_{pref} \cdot \# \sym{P}$ is used for learning.
 Errors in the input preferences are simulated by flipping preferences (e.g. $e_1 \succ e_2$ becomes $e_1 \prec e_2$).
@@
 
-Figure~\ref{fig:fractie} shows how the scores improve as the number of examples and the fraction of included preferences increases.
+\sam{change figures}
+Figure~\ref{fig:fractie} shows how the scores improve as the fractions of examples ($p_{train}$) and preferences ($p_{pref}$) that are used for learning increase.
 In all cases, more than half the pairs of examples are correctly predicted and high scores can be obtained even for small datasets.
@@
 
-\begin{question}[Noise]
-  What is the influence of noise on the accuracy?
+\begin{question}[Errors]
+  What is the influence of errors in preferences on the accuracy?
 \end{question}
 
-The influence of noise is shown in figure~\ref{fig:ruis}.
-Hereby, the second approach of testing is applied, using potentially overlapping train- and test sets.
+The influence of errors is shown in figure~\ref{fig:ruis}.
+Hereby, the second approach of testing is applied using overlapping train- and test sets.
 High scores are obtained, even despite significant levels of noise.
-The figure also shows that providing more rankings improves the robustness of the algorithm, even if the relative amount of noise remains unchanged.
+The figure also shows that increasing the number of preferences used for learning ($p_{pref}$) improves the robustness of the algorithm, even if the relative amount of errors ($p_{error}$) remains unchanged.
 
@@
   \begin{table}[!htp]
-    \caption{Scores for different thresholds ($t$)}
+    \caption{Accuracy for different thresholds ($t$)}
     \begin{tabularx}{\linewidth}{XXXX}
@@
 
-Table~\ref{tbl:limiet} demonstrates that increasing the threshold used for finding soft constraints does not improve the score.
-This experiment used 40\% of the examples as training set and 40\% of the available rankings.
-However, if the size of the problem and examples is increased, a higher threshold will likely be appropriate.
+Table~\ref{tbl:limiet} demonstrates that increasing the threshold used for finding soft constraints does not improve the accuracy ($p_{train} = p_{pref} = 0.4$).
+However, if the number of examples is increased, a higher threshold will likely be appropriate.
 
@@
 The system requires only a minimal amount of information from the user, however, it also allows for the use of expressive background knowledge.
-The learned constraints are domain independent \sam{review domain}, which facilitates the construction of positive examples.
+Constraints that are learned from examples do not depend on the domains used in those examples.
+This 
+The learned constraints are independent of the domains used in the examples, which facilitates the construction of positive examples.
 

### 2015-10-23 Paper.tex
Shorten table and shorten efficiency

@@
   \caption{Execution times overview}
-  \begin{tabularx}{\linewidth}{rl|ll}
+  \begin{tabularx}{\linewidth}{r|X}
 
-\textbf{Omitted}  & \textbf{Problem}    & \textbf{Average time (s)} \\ % & \textbf{Time (baseline)}  
+\textbf{Problem}    & \textbf{Average time (s)} \\ % & \textbf{Time (baseline)}  
 \toprule
-Nothing           & Map coloring        & 1.581   ($\pm$ 0.117)     \\ % & 1.000                     
-   (baseline)     & Sudoku              & 4.787   ($\pm$ 0.062)     \\ % & 1.000                     
-                  & Elevator            & 3.182   ($\pm$ 0.073)     \\ % & 1.000                     
-                  & Co-housing          & 25.903  ($\pm$ 0.446)     \\ % & 1.000                     
-\midrule
-Range \sam{Remove}             & Map coloring        & 4.629   ($\pm$ 0.199)     \\ % & 2.928                     
-restriction       & Sudoku              & 16.118  ($\pm$ 0.154)     \\ % & 3.367                     
-                  & Elevator            & 40.453  ($\pm$ 0.319)     \\ % & 12.713                    
-                  & Co-housing          & 207.768 ($\pm$ 0.330)     \\ % & 8.021                     
-\midrule
-Connected         & Map coloring        & 1.589   ($\pm$ 0.110)     \\ % & 1.005                     
-    clauses       & Sudoku              & 7.068   ($\pm$ 0.150)     \\ % & 1.476                     
-                  & Elevator            & 6.157   ($\pm$ 0.114)     \\ % & 1.935                     
-                  & Co-housing          & 103.633 ($\pm$ 0.131)     \\ % & 4.001                     
+Map Coloring        & 1.581   ($\pm$ 0.117)     \\ % & 1.000                     
+Sudoku              & 4.787   ($\pm$ 0.062)     \\ % & 1.000                     
+Elevator            & 3.182   ($\pm$ 0.073)     \\ % & 1.000                     
+Co-housing          & 25.903  ($\pm$ 0.446)     \\ % & 1.000                     
+% \midrule
+% Range \sam{Remove}             & Map coloring        & 4.629   ($\pm$ 0.199)     \\ % & 2.928                     
+% restriction       & Sudoku              & 16.118  ($\pm$ 0.154)     \\ % & 3.367                     
+%                   & Elevator            & 40.453  ($\pm$ 0.319)     \\ % & 12.713                    
+%                   & Co-housing          & 207.768 ($\pm$ 0.330)     \\ % & 8.021                     
+% \midrule
+% Connected         & Map coloring        & 1.589   ($\pm$ 0.110)     \\ % & 1.005                     
+%     clauses       & Sudoku              & 7.068   ($\pm$ 0.150)     \\ % & 1.476                     
+%                   & Elevator            & 6.157   ($\pm$ 0.114)     \\ % & 1.935                     
+%                   & Co-housing          & 103.633 ($\pm$ 0.131)     \\ % & 4.001                     
   \end{tabularx}
@@
 
-\begin{question}[Efficiency] \sam{review}
+\begin{question}[Efficiency]
   How fast is the clause learning system and what is the effect of various design decisions and input on the execution time?
@@
 
-Table~\ref{tbl:uitvoering} shows the execution times for several experiments.
-It shows that smaller problems can be solved efficiently and shows that removing the syntactical restrictions exponentially increase the search space.
+Table~\ref{tbl:uitvoering} shows the execution times for the different problems.
+It shows that smaller problems can be solved efficiently.
 
-All efficiency measures (i.e. symmetric predicates, subset test and representative test) have been able to improve the execution time, sometimes by more than 50\%.
-The overhead they introduce is more than compensated by the efficiency gains they cause.
+For a discussion of the experiments evaluating the influence design decisions and input we refer to \cite{kolb2015thesis} and only state the most important conclusions.
+Removing syntactical restrictions is shown to exponentially increase the execution time.
+Moreover, the other efficiency measures (e.g. symmetric predicates) are shown to be effective, individually achieving speedups of up to 50\%.
 
-Further experiments show that increasing the number of variables or literals per clause impacts the efficiency.
-Especially the combination of more variables \emph{and} literals can steeply increase the execution time.
-Therefore it would be useful to adapt these parameters dynamically.
-
-Adding additional examples only increases the execution time by a constant factor.
+Increasing the number of variables or literals per clause negatively impacts the efficiency.
+Especially allowing both more variables \emph{and} more literals can steeply increase the execution time.
+Adding additional examples, however, only increases the execution time by a constant factor.
 Since adding more examples can help improve the accuracy, this trade-off is often worthwhile.
@@
 \subsection{Optimization} \sam{Preferences, not rankings}
+\sam{Score, not accuracy}
+\sam{Change figure}
 The efficiency of the learning system for optimization criteria depends mainly on the efficiency of learning soft constraints.

### 2015-10-24 Packaged/Paper.tex
Work on final version (prior to reviews)

--- 2015-10-23 Paper.tex
@@
 
-\graphicspath{{Graphics/}}
-
 \newcommand{\sym}[1]{\ensuremath{\mathcal{#1}}}
@@
 
-\newtheorem{definition}{Definition}[section]
 \newtheorem{question}{Question}
-\AfterEndEnvironment{definition}{\noindent\ignorespaces}
-
 \theoremstyle{definition}
 \newtheorem{example}{Example}
-\newtheorem{experiment}{Experiment}[question]
-\newtheorem{observation}{Observation}[question]
-
-\newcommand{\anton}[1]{{\color{green}(#1)}}
-\newcommand{\sam}[1]{{\color{red}(#1)}}
 
@@
 
-Learning constraints has been addressed in constraint programming (e.g.
-\cite{Beldiceanu:ModelSeeker}, \cite{bessiere2013constraint}) and inductive logic programming (e.g.
-\cite{DeRaedt:ClausalDiscovery}).
+Learning constraints has been addressed in constraint programming (e.g. \cite{Beldiceanu:ModelSeeker,bessiere2013constraint}) and inductive logic programming (e.g. \cite{DeRaedt:ClausalDiscovery}).
 Recent efforts \cite{Lallouet:LearningCP} have attempted to combine these two fields.
-Furthermore, \cite{campigotto2011active}) have learned propositional weighted MAX-SAT theories from examples.
+Furthermore, \cite{campigotto2011active} have learned propositional weighted MAX-SAT theories from examples.
 Also, work in statistical relational learning (SRL) on learning the structure of Markov Logic networks (MLN) or other SRL models \cite{kok2005learning} is relevant, as MAP inference can be used to query for the most likely state.
@@
 Section~2 reviews background and related work.
-In Section~3 we introduce the formalism and the setting, in Section~4 the algorithm and the clausal optimization system to solve this problem, and in Section~5 we report on some experiments.
+We introduce the formalism and the setting in Section~3, the algorithm and the clausal optimization system in Section~4, and in Section~5 we report on some experiments.
 Finally, Section~6 concludes this paper.
@@
 In this paper constants will be written in uppercase ($\mathit{John}$) and variables in lowercase ($\mathit{x}$).
-A literal can be an atom $a$ (positive literal) or a negated atom $\lnot a$ (negative literal).
+A literal can be an atom~$a$ (positive literal) or a negated atom~$\lnot a$ (negative literal).
 Clauses are disjunctions of literals $l_1 \lor l_2 \lor ... \lor l_k$ and are assumed to be universally quantified.
@@
 \paragraph{Optimization}
-The Boolean Satisfiability Problem (SAT) attempts to determine whether a propositional formula can be satisfied.
+The Boolean Satisfiability problem (SAT) is the problem of determining whether a propositional formula can be satisfied.
 Formulas can be rewritten into Conjunctive Normal Form (CNF), which consists of a conjunction of clauses (disjunctions).
@@
 In related work \cite{campigotto2011active} weighted MAX-SAT problems are induced by learning propositional clauses and weights using examples and absolute scores.
-\sam{Change this section}
-Inspired by their results, our research tries to learn first-order logic clauses and weights using examples and rankings.
-The resulting optimization problem can be seen as an extension of weighted MAX-SAT in which the clauses are in first-order logic.
+The representation we use is an extension of weighted MAX-SAT to first-order logic.
+We also learn this model using preferences over sets of examples (possible worlds), rather than assigning target values to examples.
 
@@
 ModelSeeker \cite{Beldiceanu:ModelSeeker} structures the problem variables in different ways.
-The system then uses a large catalog of global constraints to find the ones that hold within this structure. 
+A large catalog of global constraints is used to find constraints that hold within this structure.
 This approach allows for constraint learning from a few examples.
 
-As mentioned earlier, the research of \cite{Lallouet:LearningCP} used ILP techniques to learn constraints.
-In their approach, the search space of constraints is explored in a bidirectional manner, primarily using negative examples.
-
-\paragraph{Clausal Discovery} \sam{Dissolve}
-The clausal discovery algorithm \cite{DeRaedt:ClausalDiscovery} attempts to learn first-order logic clauses from positive examples.
-Using a refinement operator, the search space is traversed by incrementally generalizing clauses until they cover all examples, starting from the empty clause~($\square$).
-The result set consists of the most specific clauses covering all examples.
+As mentioned earlier, the research of \cite{Lallouet:LearningCP} uses ILP techniques to learn constraints.
+They explore the search space of constraints in a bidirectional manner, primarily using negative examples.
 
@@
 Examples are Herbrand interpretations.
-They consist of set of constants, the domain, and all ground atoms over these constants that are true.
+They consist of a set of constants, the domain, and all ground atoms over these constants that are true.
 The learned clauses are domain-independent and contain only variables.
@@
 A preference $e_1 \succ e_2$ indicates that out of examples $e_1$ and $e_2$ the user prefers $e_1$.
+Such preferences only consider a few examples at a time and are usually easier to provide than absolute scores.
 
@@
   Example~$e_1$ is covered by clauses $c_1$, $c_2$ and~$c_3$, therefore, $score(e_1, \sym{M}) = 0.5 + 0.25 + 1 + 0 = 1.75$.
-  Similarly, $score(e_2, \sym{M}) = -1$ and $score(e_3, \sym{M}) = 0$ can be computer.
+  For examples $e_2$ and~$e_3$ we can compute $score(e_2, \sym{M}) = -1$ and $score(e_3, \sym{M}) = 0$, analogously.
   Based on these scores, the model \sym{M} would prefer $e_1$ over~$e_2$ ($e_1 \succ e_2$), $e_1$ over~$e_3$ ($e_1 \succ e_3$) and $e_2$ over~$e_3$ ($e_2 \succ e_3$).
@@
 \section{Approach}
-In this work, we use a two-step approach to learn first-order weighted MAX-SAT.
+We use a two-step approach to learn first-order weighted MAX-SAT.
 First, clause learning is used to find soft constraints in the given examples.
@@
 
-\subsection{Input} \sam{Dissolve, partially or entirely}
-The input consists of global definitions that specify the types and predicates that are used to describe examples and a set of examples \sym{E}.
-\anton{Explain: what is type information?}
-Type information relates directly to the problem domain and is usually easy to provide.
-Including types also improves the accuracy and efficiency of the learned clauses.
-
-Examples are interpretations, they specify the constants in their domain and all predicates that are true.
-Predicates can also be generated by background knowledge, in which case they are not included in the example.
-
-When learning optimization criteria, preference information is also provided by the user.
-Each preference is of the form $e_1 \succ e_2 \succ ... \succ e_n$ with $e_i \in \sym{E}$.
-Preferences describe the relative position of a few examples at a time and are typically easier to provide than absolute scores for all examples.
-
-\begin{example} \sam{Rewrite}
-  \label{ex:moving-input}
-  Consider the moving problem (example~\ref{ex:moving-input}).
-  A model could, for example, use a type \textit{Area} and predicate definitions low\_crime(\textit{Area}), cheap(\textit{Area}), school\_in(\textit{Area}), work\_in(\textit{Area}) and live\_in(\textit{Area}).
-  Examples would then specify various areas $A_i$ and describe the conditions (e.g.
+% \begin{example}
+%   \label{ex:moving-input}
+%   Consider the moving problem (example~\ref{ex:moving-input}).
+%   A model could, for example, use a type \textit{Area} and predicate definitions low\_crime(\textit{Area}), cheap(\textit{Area}), school\_in(\textit{Area}), work\_in(\textit{Area}) and live\_in(\textit{Area}).
+%   Examples would then specify various areas $A_i$ and describe the conditions (e.g.
 cheap($A_1$), low\_crime($A_2$)) as well as the users choices (e.g.
 school\_in($A_2$), work\_in($A_2$), live\_in($A_1$)).
-  % Imagine a problem that describes humans with a type \textit{human} and predicates human(\textit{human}), male(\textit{human}) and female(\textit{human}).
-  % An example could contain the constants S and A, both of type \textit{human}, and the relations: human(S), human(A), male(S), female(A).
-  % Clauses to be learned would be: human(x) $\leftarrow$ male(x), human(x) $\leftarrow$ female(x),  female(x) $\lor$ male(x) $\leftarrow$ human(x) and \textit{false} $\leftarrow$ female(x) $\land$ male(x).
-\end{example}
+%   % Imagine a problem that describes humans with a type \textit{human} and predicates human(\textit{human}), male(\textit{human}) and female(\textit{human}).
+%   % An example could contain the constants S and A, both of type \textit{human}, and the relations: human(S), human(A), male(S), female(A).
+%   % Clauses to be learned would be: human(x) $\leftarrow$ male(x), human(x) $\leftarrow$ female(x),  female(x) $\lor$ male(x) $\leftarrow$ human(x) and \textit{false} $\leftarrow$ female(x) $\land$ male(x).
+% \end{example}
 
@@
     \If{$\#\{e \in \sym{E} | e \models c\} \geq t$}
-      \If{$\lnot (\sym{T} \models c)$}
+      \If{$\sym{T} \not\models c$}
         \State $\sym{T} = \sym{T} \cup c$
@@
 The clause learning system is based on clausal discovery \cite{DeRaedt:ClausalDiscovery}.
-Algorithm~\ref{alg:cd} illustrates how clausal discovery works on a high level.
-Note the double use of the $\models$ operator.\sam{Review this once yet}
-On the one hand, it is used in $e \models c$ to denote that clause~$c$ is true in the example (interpretation)~$e$ and $c$ is said to cover $e$.
-This test is used to compute if a clause covers enough examples.
-On the other hand, in $\sym{T} \models c$ it signifies that \sym{T} logically entails $c$.
-The algorithm uses this test to remove redundant clauses.
-For both cases the IDP knowledge base system \cite{de2013prototype,wittocx2008idp} is used to compute $\models$.
+Algorithm~\ref{alg:cd} illustrates how clausal discovery works on a high level\footnote{Note the double usage of $\models$.
+Its meaning depends on whether it is prefixed by an interpretation or a logical theory.}.
+To compute if a clause~$c$ covers enough examples, the number of examples~$e$ is counted for which $e \models c$ (i.e. $c$ is true in the example~$e$ or $c$ covers $e$).
+In order to remove redundant clauses a new clause~$c$ is only added to the result set $\sym{T}$ if $\sym{T} \not\models c$ (i.e. $c$ is not logically entailed by \sym{T}).
+Both $e \models c$ and $\sym{T} \not\models c$ are computed by the IDP knowledge base system \cite{de2013prototype,wittocx2008idp}.
 The algorithm can find both soft constraints ($t < \# \sym{E}$) or hard constraints ($t = \# \sym{E}$).
@@
 By adding just one literal the refinement operator $\rho$ generates the set of clauses that are minimally more general.
-In order to do this efficiently, a list of atoms is pre-calculated, given a maximal number of variables that can occur within a clause.
+In order to do this efficiently, our implementation uses a pre-calculated list of atoms, given a maximal number of variables that can occur within a clause.
 Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.
@@
 Therefore, a clause~$c$ with variables \sym{V} is extended to the clause~$c_{OI}$ by adding literals $x_i \neq x_j$, for $x_i, x_j \in \sym{V}, i \neq j$.
-Finally, the implementation natively supports symmetric predicates for which the order of the terms do not matter
+Finally, the implementation natively supports typed constants and predicates, symmetric predicates (i.e.
+predicates where the order of terms does not matter) and predicates that are generated using background knowledge.
 
@@
 In order to find weighted soft constraints, soft constraints~\sym{C} for the given examples are learned by using the constraint learning system with a (low) threshold.
-The example are then translated to Boolean feature vectors of length $\# \sym{C}$ by introducing a feature for every clause~$c_i \in \sym{C}$.
+The examples are then translated to Boolean feature vectors of length $\# \sym{C}$ by introducing a feature for every clause~$c_i \in \sym{C}$.
 For an example~$e$, the $i$-th value of its feature vector $\mathbf{f}_e$ is calculated as $\mathbf{f}_e(i) = \mathbf{1}_{e \models c_i}$.
 Next, existing software is used to learn the linear function $\sum_i w_i \cdot c_i$ over these features, based on user provided preferences.
-This function corresponds with the scoring function (\ref{eqn:score})) introduced in Section~\ref{sec:problem}.
+This function corresponds to the scoring function (\ref{eqn:score}) introduced in Section~3.
 It can be used to calculate scores for (unseen) examples and by maximizing the score function an optimal solution can be found.
 
-\begin{example} \sam{Rewrite}
+\begin{example}
   \label{ex:features}
-  Consider examples~$e_1, e_2, e_3$, rankings $e_1 \succ e_2, e_2 \succ e_3$ and clauses $c_1, c_2, c_3$.
-  If the clauses cover examples according to table~\ref{tbl:cover_examples}, then, for example, the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.
+  Consider examples~$e_1, e_2, e_3$, preferences $e_1 \succ e_2, e_2 \succ e_3$ and clauses $c_1, c_2, c_3$.
+  If the clauses cover examples according to table~\ref{tbl:cover_examples}, then, for example, the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given preferences.
   A new example that is covered by~$c_1$ and~$c_2$, would be assigned a score of $1$, according to this function.
   This score has no value in the absolute sense, it can only be used to compare it to other examples ranked by the same function.
-  In this example clause~$c_1$ represents a desirable property, clause~$c_2$ is ignored because it does not influence the ranking and clause~$c_3$ represents an undesirable property.
+  In this example clause~$c_1$ represents a desirable property and clause~$c_3$ represents an undesirable property.
 
@@
 The weights that it assigns to the features can be used directly as weights for the optimization criteria.
-While this paper uses \svm{} other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.
+While the current implementation uses \svm{}, other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.
 
 \subsection{Optimal Solution}
-Solvers, such as IDP \sam{more on IDP, can be used as ASP solver?}, can use clauses directly as hard constraints to generate a solution.
-There is no solver that can use the chosen optimization criteria directly.
-Weighted MAX-SAT solvers use propositional clauses and only allow for positive weights.
-
-This optimization task can be solved in IDP, using inductive definitions, aggregates and minimization.
-The only limitation is that the current version only supports integer values.
-Therefore, the weights of the clauses are divided by the smallest absolute weight, multiplied by a constant and rounded to the closest integer if necessary.
+Typical MAX-SAT solvers use propositional clauses and only allow for positive weights.
+However, the first-order logic solver IDP can solve our optimization task as it supports inductive definitions, aggregates and minimization.
+The only limitation is that the current version is restricted to integer values and does not support rational numbers.
+Therefore, the weights of the clauses are divided by the smallest absolute weight, multiplied by a constant and rounded to the closest integer.
 
@@
 In our experiments we aim to measure the accuracy and efficiency of the learning systems for constraints and optimization criteria.
-In order to account for non-determinism, experiments measuring execution times or optimization scores were repeated eight times and their result were averaged.
-\sam{First step evaluation constraint learning, then optimization}
+Non-deterministic experiments (e.g. measuring execution times or accuracy of clausal optimization) were repeated eight times and their result were averaged.
+Some problems and experiments are only described briefly or summarized, for more details we refer to \cite{kolb2015thesis}\footnote{Available online at https://github.com/samuelkolb/thesis}.
+
+First, experiments concerning the accuracy and efficiency of the constraint learning implementation are discussed.
+Thereafter, we discuss the evaluation of the accuracy of the clausal optimization system\footnote{The system (v1.0) used to perform all experiments is available online at https://github.com/samuelkolb/clausal-discovery/releases}.
+
 \subsection{Constraints}
-We used four problems to evaluate constraint learning: \sam{clarify hard constraints}
+We used four problems to evaluate constraint learning:
 \begin{description}
@@
 Two examples, each containing three correctly colored countries, are given.
-\item [Sudoku] A single, solved $4 \times 4$ sudoku is given.
-\item[Elevator] Three examples have been generated, two of which respect the underlying soft constraint.
-\item[Co-housing] Contains four hard constraints and five examples have been generated that respect the constraints.
+\item[Sudoku] A single, solved $4 \times 4$ sudoku is given as an example for this problem.
+\item[Elevator] This problem uses three predicates $\mathit{inside}$, $\mathit{crowded}$ and $\mathit{panic}$.
+The predicate $\mathit{crowded}$ is generated by background knowledge and is true for all elevators with three or more people inside.
+Out of the three examples provided for this problem, two satisfy the target soft constraint:
+\begin{flalign*}
+  &\mathit{panic}(p)~\leftarrow~\mathit{crowded}(e)~\land~\mathit{inside(p, e)}.
+\end{flalign*}
+\item[Co-housing] The co-housing problem uses background knowledge to specify one of its predicates ($\mathit{cheap})$) and to impose some additional constraints.
+Five examples are provided, which all adhere to the following four constraints on the housing choices of two friends:
+\begin{enumerate}
+  \item If the two friends do not live in the same area, they work in the same area.
+  \item If a person does not work and live in the same area, that person needs a car.
+  \item A person who lives in a cheap area has a car.
+  \item If the two friends live together, they live in an expensive area.
+\end{enumerate}
 \end{description}
-\anton{more elaborate descriptions of these?}
-\sam{link to thesis}
+The clausal discovery implementation is used to learn soft constraints for the elevator problem ($t = 1.5$) and hard constraints for the other problems.
 
@@
 Often additional constraints were found that describe some structure in the problems.
-For example, for the map coloring problem one of the learned constraint states that countries are never their own neighbor.
+For example, for the map coloring problem one of the learned constraints states that countries are never their own neighbor.
 These kind of constraints may help a constraint solver to work more efficiently.
@@
 \begin{table}
-  \caption{Execution times overview}
+  \caption{Execution times for learning constraints}
   \begin{tabularx}{\linewidth}{r|X}
@@
 Elevator            & 3.182   ($\pm$ 0.073)     \\ % & 1.000                     
-Co-housing          & 25.903  ($\pm$ 0.446)     \\ % & 1.000                     
+Co-housing          & 25.903  ($\pm$ 0.446)      % & 1.000                     
 % \midrule
-% Range \sam{Remove}             & Map coloring        & 4.629   ($\pm$ 0.199)     \\ % & 2.928                     
+% Range             & Map coloring        & 4.629   ($\pm$ 0.199)     \\ % & 2.928                     
 % restriction       & Sudoku              & 16.118  ($\pm$ 0.154)     \\ % & 3.367                     
@@
 
-Table~\ref{tbl:uitvoering} shows the execution times for the different problems.
+Table~\ref{tbl:uitvoering} shows the execution times for the different problems\footnote{All experiments were run on a MacBook Pro with an i7 quad-core processor}.
 It shows that smaller problems can be solved efficiently.
@@
 Especially allowing both more variables \emph{and} more literals can steeply increase the execution time.
-Adding additional examples, however, only increases the execution time by a constant factor.
+Including additional examples, however, only increases the execution time by a constant factor.
 Since adding more examples can help improve the accuracy, this trade-off is often worthwhile.
@@
 
-\subsection{Optimization} \sam{Preferences, not rankings}
-\sam{Score, not accuracy}
-\sam{Change figure}
+\subsection{Optimization}
 The efficiency of the learning system for optimization criteria depends mainly on the efficiency of learning soft constraints.
@@
 
-The moving problem with input according to example~\ref{ex:moving-input} is used to evaluate the learning of optimization criteria.
-18 possible configurations are used as available examples~(\sym{E}).
-Four weighted soft constraints are used to represent the underlying model~(\sym{M}) for the users preferences.
-Two approaches are used for evaluation and differ in the way they construct the training ($\sym{E}_{train}$) and test sets ($\sym{E}_{test}$).
-In the first approach, given a parameter~$p_{train}$, the examples are randomly partitioned into disjoint train and test sets ($\sym{E}_{test}$), such that $\#\sym{E}_{train} = p_{train} \cdot \#\sym{E}$.
+Example~\ref{ex:moving} introduced the moving city scenario which is used to evaluate the learning of optimization criteria.
+18 possible configurations are used as available examples~(\sym{E}), differing in their values for $\mathit{live\_in}$, $\mathit{work\_in}$ and $\mathit{school\_in}$.
+The four weighted soft constraints are used to represent the underlying model~(\sym{M}) for the users preferences.
+Two approaches are used for evaluation and vary in the way they construct the training ($\sym{E}_{train}$) and test sets ($\sym{E}_{test}$).
+In the first approach, given a parameter~$p_{train}$, the examples are randomly partitioned into disjoint training and test sets ($\sym{E}_{test}$), such that $\#\sym{E}_{train} = p_{train} \cdot \#\sym{E}$.
 The second approach uses $\sym{E}_{train} = \sym{E}_{test} = \sym{E}$.
@@
   \centering
-    \includegraphics[width=1.1\linewidth]{rankings}
-  \caption{Influence fraction of inequalities}
+    \includegraphics[width=1\linewidth]{accuracy}
+  \caption{Influence of input size ($p_{train}$ and $p_{pref}$)}
   \label{fig:fractie}
@@
 
-\sam{change figures}
-Figure~\ref{fig:fractie} shows how the scores improve as the fractions of examples ($p_{train}$) and preferences ($p_{pref}$) that are used for learning increase.
-In all cases, more than half the pairs of examples are correctly predicted and high scores can be obtained even for small datasets.
-Additional experiments have shown that even for lower scores the learned optimization criteria are often capable of identifying the correct optimal solution. \sam{Refer to thesis}
+Figure~\ref{fig:fractie} shows how the accuracy improve as the fractions of examples ($p_{train}$) and preferences ($p_{pref}$) that are used for learning increase.
+In all cases, more than half the pairs of examples are correctly predicted and the high accuracy values can be obtained even for small datasets.
+Additional experiments have shown that even for lower accuracy values the learned optimization criteria are often capable of identifying the correct optimal solution \cite{kolb2015thesis}.
 
@@
 In order to test the accuracy for models which cannot be directly expressed, a model consisting of two disconnected clauses has been tested as well.
-While there are limits to the expressibility, the learned optimization criteria were able to obtain similar scores and seem to be robust with respect to the exact formulation.
+While there are limits to the expressibility, the learned optimization criteria were able to obtain similar levels of accuracy and seem to be robust with respect to the exact formulation.
 
@@
 The influence of errors is shown in figure~\ref{fig:ruis}.
-Hereby, the second approach of testing is applied using overlapping train- and test sets.
-High scores are obtained, even despite significant levels of noise.
+For this experiment, the second approach of testing has been applied, using overlapping training and test sets ($\sym{E}_{train} = \sym{E}_{test} = \sym{E}$).
+High accuracy values are obtained, despite significant numbers of errors.
 The figure also shows that increasing the number of preferences used for learning ($p_{pref}$) improves the robustness of the algorithm, even if the relative amount of errors ($p_{error}$) remains unchanged.
@@
   \centering
-    \includegraphics[width=1.1\linewidth]{noise}
-  \caption{Influence of noise}
+    \includegraphics[width=1\linewidth]{errors}
+  \caption{Influence of preference errors ($p_{error}$)}
   \label{fig:ruis}
@@
 We presented an approach to automatically acquire constraints and optimization criteria from examples and preferences.
-Implementation for both clause learning and clausal optimization have been provided that accomplish these tasks using first-order logic clauses.
+Implementations for both clause discovery and clausal optimization have been provided that accomplish these tasks using first-order logic clauses.
 
 The constraint learning implementation has been able to learn the relevant hard and soft constraints in all experiments.
-For each problem, only a small number of examples was given to learn from.
-The system requires only a minimal amount of information from the user, however, it also allows for the use of expressive background knowledge.
-Constraints that are learned from examples do not depend on the domains used in those examples.
-This 
-The learned constraints are independent of the domains used in the examples, which facilitates the construction of positive examples.
+For each problem, the relevant constraints could be learned from only a few examples.
+Although the system requires only examples and preferences, users can provide more information through background knowledge.
+The constraints learned by clausal discovery are independent of the domains of specific problem instantiations (examples), which facilitates the construction of positive examples.
 
-Using the constraint learning implementation, the goal of learning optimization criteria has been accomplished.
-Optimization criteria can be learned that enable optimal solutions to be found.
-Even for small datasets and noisy rankings, constraints are found that rank most examples correctly.
+Using the constraint learning implementation, the goal of learning a weighted first-order MAX-SAT model has been accomplished.
+The learned models allow optimal solutions to be found.
+Even for small datasets and erroneous preferences, optimization criteria are found that rank most examples correctly.
 
@@
 
-\paragraph{Future work} \sam{shorten}
+\paragraph{Future work}
 There are several opportunities for future work.
-It would be interesting to adapt the number of variables and literals in a clause dynamically.
-This could be accomplished, for example, by using negative examples.
-Learned clauses must be specific enough to not cover any negative examples. 
-\anton{These restrictions are not really described in the paper, so maybe not that interesting to mention here? Or at least add a sentence to introduce them.}
-
-Additionally, it would be interesting to add interactivity to the learning system for the generation of examples or rankings.
-Rankings expressing that examples are equally ranked are currently ignored.
-Whenever provided explicitly, however, it could be interesting to incorporate this information into the algorithm.
-
-As mentioned earlier, the domain independence \sam{review domain} of the clauses has several advantages.
-In some cases, however, specific objects are inherently present in any problem instantiation.
-The constraint learning system could be enhanced to include such global constants.
+It would be interesting to adapt the number of variables and literals per clause dynamically, as they have a strong influence on the efficiency and accuracy of the clause learning system.
+Additionally, interactivity could be added to the learning system by actively generating examples or preferences and querying the user about their validity.
 
@@
 
-\newpage
-
 %

### 2015-10-30 Paper.tex
Implementing Angelika's suggestions and rewriting, part 1

@@
 
-\graphicspath{{Graphics/}}
+\graphicspath{{Graphics/}{}}
 
@@
 
-\newtheorem{definition}{Definition}[section]
 \newtheorem{question}{Question}
-\AfterEndEnvironment{definition}{\noindent\ignorespaces}
-
 \theoremstyle{definition}
 \newtheorem{example}{Example}
-\newtheorem{experiment}{Experiment}[question]
-\newtheorem{observation}{Observation}[question]
 
-\newcommand{\anton}[1]{{\color{green}(#1)}}
 \newcommand{\sam}[1]{{\color{red}(#1)}}
+\newcommand{\abs}[1]{\left|#1\right|}
 
@@
 The technique is also empirically evaluated on a number of examples.
-
+These experiments show that the system is capable of learning clauses and weights that accurately capture underlying models.
 \end{abstract}
@@
 \section{Introduction}
+\sam{it might be good to state somewhere early that you are in the "learning from interpretations" setting, but additionally have preferences}
+\sam{Structure}
 An important class of declarative models is concerned with (discrete) constrained optimization problems.
@@
 We use (function free) first-order clauses; which we now introduce.
-An atom $p(t_1, t_2, ..., t_n)$ consists of a predicate symbol~$p$ that is applied to $n$ terms~$t_i$.
+An atom $p(t_1, t_2, ..., t_n)$ consists of a predicate symbol~$p$ that is applied to $n$ terms~$t_i$. \sam{Casing}
 Terms are either constants or variables.
 In this paper constants will be written in uppercase ($\mathit{John}$) and variables in lowercase ($\mathit{x}$).
-A literal can be an atom $a$ (positive literal) or a negated atom $\lnot a$ (negative literal).
+A literal can be an atom~$a$ (positive literal) or a negated atom~$\lnot a$ (negative literal).
 Clauses are disjunctions of literals $l_1 \lor l_2 \lor ... \lor l_k$ and are assumed to be universally quantified.
@@
 
+\sam{versus MLN, PSL}
 In related work \cite{campigotto2011active} weighted MAX-SAT problems are induced by learning propositional clauses and weights using examples and absolute scores.
-\sam{Change this section}
-Inspired by their results, our research tries to learn first-order logic clauses and weights using examples and rankings.
-The resulting optimization problem can be seen as an extension of weighted MAX-SAT in which the clauses are in first-order logic.
+The representation we use is an extension of weighted MAX-SAT to first-order logic.
+We also learn this model using preferences over sets of examples (possible worlds), rather than assigning target values to examples.
 
-\paragraph{Learning Constraints}
+\paragraph{Constraint Learning}
+\sam{Define constraints}
 The goal of constraint learning is to automatically identify constraints in one or multiple examples.
@@
 We want to learn constraints and optimization criteria based on positive examples and user preferences.
-The goal of this task is to simplify the process of generating formal representations, therefore, any input should be reasonably easy for a user to provide.
+Since our aim is to simplify the process of generating formal representations, any input should be reasonably easy for a user to provide.
 
-\subsection{Constraint Learning}
-\emph{Given a set of examples and a limit $t$, the goal is to find maximally specific clauses that are satisfied by at least $t$ of the given examples.}
+In order to apply ILP techniques for constraint learning, we will use first-order logical clauses as constraints.
+
+\paragraph*{Clause Learning}
+\emph{Given a set of examples and a threshold $t$, find maximally specific clauses that are satisfied by at least $t$ of the given examples.}
 \\\\
 Examples are Herbrand interpretations.
-They consist of set of constants, the domain, and all ground atoms over these constants that are true.
-The learned clauses are domain-independent and contain only variables.
-
-\subsection{Learning Optimization Criteria}
-\emph{Given a set of examples \sym{E} and a set of preferences \sym{P} over these examples, the goal is to learn soft constraints and weights such that the order described by these optimization criteria maximally corresponds with~\sym{P}.}
+They consist of a set of constants, the domain, and all ground atoms over these constants that are true.
+The learned clauses are domain-independent hard or soft constraints and contain only variables.
 \\\\
-Every preference $p \in \sym{P}$ describes a relative order $e_1 \succ e_2 \succ ... \succ e_n$ over a subset of examples $\{e_1, e_2, ..., e_n\} \subseteq \sym{E}$.
-A preference $e_1 \succ e_2$ indicates that out of examples $e_1$ and $e_2$ the user prefers $e_1$.
-
-The soft constraints consist of clauses \sym{C} that are satisfied by some but not all of the examples.
+Our goal is to learn optimization criteria, weighted soft constraints, based on examples (interpretations) and user preferences.
+A (pairwise) preference $e_1 \succ e_2$ describes a relative order over examples $e_1$ and~$e_2$, stating that $e_1$ is preferred over~$e_2$.
+Preferences only consider two examples at a time and are usually easier to provide than absolute scores.
+The soft constraints consist of clauses~(\sym{C}) that are satisfied by some but not all of the examples.
 They can be identified using a system that implements the constraint learning task.
-The weighted clauses \sym{W} are tuples $\mathbb{R} \times \sym{C}$ that can be used to calculate a score for any example $e$:
+Weighted clauses~($\sym{W}_{\sym{C}}$) are tuples $\mathbb{R} \times \sym{C}$ that can be used to calculate a score for any example $e$:
 \begin{eqnarray}
   \label{eqn:score}
-  score(e, \sym{W}) = \sum\limits_{(\mathit{w}, \mathit{c}) \in \sym{W}} \mathit{w} \cdot \mathbf{1}_{e \models c}
+  score_{\sym{W}_{\sym{C}}}(e) = \sum\limits_{(\mathit{w}, \mathit{c}) \in \sym{W}_{\sym{C}}} \mathit{w} \cdot \mathbf{1}_{e \models c}
 \end{eqnarray}
+An interpretation~$e_{opt}$ is an optimal solution with respect to~$\sym{W}_{\sym{C}}$ if $score_{\sym{W}_{\sym{C}}}(e_{opt})$ is maximal.
 
@@
 
-  Assume the following first-order weighted MAX-SAT model \sym{M} describes the optimization problem to solve:
+  Assume the following set of weighted clauses~\sym{M} describes the optimization problem to solve:
   \begin{flalign*}
@@
   \end{flalign*}
-  Let examples $e_1$, $e_2$ and~$e_3$ be interpretations that reason over the same city.
-  Specifically, let all examples share constants (areas) $A_1$, $A_2$ and $A_3$ and ground atoms $\mathit{cheap}(A_1)$, $\mathit{cheap}(A_3)$, $\mathit{low\_crime}(A_1)$ and $\mathit{low\_crime}(A_2)$.
-Additionally, they specify:
+  Let examples $e_1$, $e_2$ and~$e_3$ be Herbrand interpretations that reason over the same city, using the same constants (areas) $A_1$, $A_2$ and $A_3$.
+  Consider the set~\sym{S} of shared ground atoms $\{\mathit{cheap}(A_1)$, $\mathit{cheap}(A_3)$, $\mathit{low\_crime}(A_1)$, $\mathit{low\_crime}(A_2)\}$ and define the examples as:
   \begin{flalign*}
-    &e_1: \mathit{live\_in}(A_1), \mathit{work\_in}(A_2), \mathit{school\_in(A_2)} \\
-    &e_2: \mathit{live\_in}(A_2), \mathit{work\_in}(A_2), \mathit{school\_in(A_3)} \\
-    &e_3: \mathit{live\_in}(A_2), \mathit{work\_in}(A_2), \mathit{school\_in(A_1)}
+    &e_1 = \{\mathit{live\_in}(A_1), \mathit{work\_in}(A_2), \mathit{school\_in(A_2)}\} \cup \sym{S} \\
+    &e_2 = \{\mathit{live\_in}(A_2), \mathit{work\_in}(A_2), \mathit{school\_in(A_3)}\} \cup \sym{S}\\
+    &e_3 = \{\mathit{live\_in}(A_2), \mathit{work\_in}(A_2), \mathit{school\_in(A_1)}\} \cup \sym{S}.
   \end{flalign*}
 
-  The score of an example for the model~\sym{M} can be calculated using (\ref{eqn:score}).
-  Example~$e_1$ is covered by clauses $c_1$, $c_2$ and~$c_3$, therefore, $score(e_1, \sym{M}) = 0.5 + 0.25 + 1 + 0 = 1.75$.
-  Similarly, $score(e_2, \sym{M}) = -1$ and $score(e_3, \sym{M}) = 0$ can be computer.
-  Based on these scores, the model \sym{M} would prefer $e_1$ over~$e_2$ ($e_1 \succ e_2$), $e_1$ over~$e_3$ ($e_1 \succ e_3$) and $e_2$ over~$e_3$ ($e_2 \succ e_3$).
-  The aim of our research is to learn a model like \sym{M}, given such examples and preferences.
+  The score of an example can be calculated using~(\ref{eqn:score}).
+  Example~$e_1$ is covered by clauses $c_1$, $c_2$ and~$c_3$, therefore, $score_{\sym{M}}(e_1) = 0.5 + 0.25 + 1 + 0 = 1.75$.
+  Analogously we can compute $score_{\sym{M}}(e_2) = -1$ and $score_{\sym{M}}(e_3) = 0$.
+  Using $score_{\sym{M}}$, the following preferences can be generated: $e_1 \succ e_2$, $e_1 \succ e_3$ and $e_2 \succ e_3$.
 \end{example}
 
-Apart from learning optimization criteria, this paper presents an approach to use these optimization criteria to find an optimal solution in practice.
+The aim of our research is to learn weighted clauses such that the order imposed by these optimization criteria maximally corresponds with the given preference information.
+
+\paragraph*{Clausal Optimization}
+\emph{Given examples~\sym{E}, preferences~\sym{P} and threshold~$t$, learn clauses~\sym{C} that cover at least $t$ examples and find weights~$\sym{W}_{\sym{C}}$ that maximize:
+\begin{flalign*}\abs{\{e_1 \succ e_2 \in \sym{P} | score_{\sym{W}_{\sym{C}}}(e_1) > score_{\sym{W}_{\sym{C}}(e_2, })\}}.\end{flalign*}}
 
@@
 \section{Approach}
-In this work, we use a two-step approach to learn first-order weighted MAX-SAT.
+We use a two-step approach to learn weighted first-order clauses.
 First, clause learning is used to find soft constraints in the given examples.
@@
   \State $\sym{Q} \gets \{\square\}$, $\sym{T} \gets \{\}$
-  \While{$\#\sym{Q} > 0$}
+  \While{$\abs{\sym{Q}} > 0$}
     \State $c \gets next(\sym{Q})$
-    \If{$\#\{e \in \sym{E} | e \models c\} \geq t$}
-      \If{$\lnot (\sym{T} \models c)$}
+    \If{$\abs{\{e \in \sym{E} | e \models c\}} \geq t$}
+      \If{$\sym{T} \not\models c$}
         \State $\sym{T} = \sym{T} \cup c$
@@
 
-\subsection{Clausal Discovery}
+\subsection{Clause Learning}
 The clause learning system is based on clausal discovery \cite{DeRaedt:ClausalDiscovery}.
 Algorithm~\ref{alg:cd} illustrates how clausal discovery works on a high level.
-Note the double use of the $\models$ operator.\sam{Review this once yet}
-On the one hand, it is used in $e \models c$ to denote that clause~$c$ is true in the example (interpretation)~$e$ and $c$ is said to cover $e$.
-This test is used to compute if a clause covers enough examples.
-On the other hand, in $\sym{T} \models c$ it signifies that \sym{T} logically entails $c$.
-The algorithm uses this test to remove redundant clauses.
-For both cases the IDP knowledge base system \cite{de2013prototype,wittocx2008idp} is used to compute $\models$.
-The algorithm can find both soft constraints ($t < \# \sym{E}$) or hard constraints ($t = \# \sym{E}$).
+To compute if a clause~$c$ covers enough examples, the number of examples~$e$ is counted for which $e \models c$ (i.e. $c$ is true in the example~$e$ or $c$ covers $e$).
+In order to remove redundant clauses a new clause~$c$ is only added to the result set $\sym{T}$ if $\sym{T} \not\models c$ (i.e. $c$ is not logically entailed by \sym{T}).
+Both $e \models c$ and $\sym{T} \not\models c$ are computed by the IDP knowledge base system \cite{de2013prototype,wittocx2008idp}.
+The algorithm can find both soft constraints ($t < \abs{\sym{E}}$) or hard constraints ($t = \abs{\sym{E}}$).
 
@@
 By adding just one literal the refinement operator $\rho$ generates the set of clauses that are minimally more general.
-In order to do this efficiently, a list of atoms is pre-calculated, given a maximal number of variables that can occur within a clause.
+In order to do this efficiently, our implementation uses a pre-calculated list of atoms, given a maximal number of variables that can occur within a clause.
 Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.
 
-A few additional measures are used to enable the implementation to be more efficient.
+A few additional measures are used to make the implementation more efficient.
 Most importantly, there are two syntactic restrictions on the form of the clauses.
@@
 Therefore, a clause~$c$ with variables \sym{V} is extended to the clause~$c_{OI}$ by adding literals $x_i \neq x_j$, for $x_i, x_j \in \sym{V}, i \neq j$.
-Finally, the implementation natively supports symmetric predicates for which the order of the terms do not matter
+Finally, the implementation natively supports typed constants and predicates, symmetric predicates (i.e. predicates where the order of terms does not matter) and background knowledge.
+A predicate can be marked as \emph{calculated} in which case it is defined in the background knowledge, rather than being provided in the examples.
 
-\subsection{Optimization}
-In order to find weighted soft constraints, soft constraints~\sym{C} for the given examples are learned by using the constraint learning system with a (low) threshold.
-The example are then translated to Boolean feature vectors of length $\# \sym{C}$ by introducing a feature for every clause~$c_i \in \sym{C}$.
-For an example~$e$, the $i$-th value of its feature vector $\mathbf{f}_e$ is calculated as $\mathbf{f}_e(i) = \mathbf{1}_{e \models c_i}$.
-Next, existing software is used to learn the linear function $\sum_i w_i \cdot c_i$ over these features, based on user provided preferences.
-This function corresponds with the scoring function (\ref{eqn:score})) introduced in Section~\ref{sec:problem}.
-It can be used to calculate scores for (unseen) examples and by maximizing the score function an optimal solution can be found.
+\subsection{Clausal Optimization}
+In order to find weighted soft constraints, the clause learning system is used with a low threshold to learn (soft) clauses~\sym{C}.
+The examples are then translated to Boolean feature vectors of length $\abs{\sym{C}}$ by introducing a feature for every clause~$c_i \in \sym{C}$.
+For an example~$e$, the value of its feature vector $\mathbf{v}_e$ for clause~$c_i$ is calculated as $\mathbf{v}_e(i) = \mathbf{1}_{e \models c_i}$.
+Next, a linear function $f(\mathbf{v})=\sum_i w_i \cdot \mathbf{v}(i)$ is learned that assigns weights~$w_i$ to clauses~$c_i$, based on the user-provided preferences.
+The scoring function~$f$ agrees with a preference $e_1 \succ e_2$ if $f(\mathbf{v}_{e_1}) > f(\mathbf{v}_{e_2})$ and corresponds to the scoring function (\ref{eqn:score}).
 
-\begin{example} \sam{Rewrite}
+\begin{example}
   \label{ex:features}
-  Consider examples~$e_1, e_2, e_3$, rankings $e_1 \succ e_2, e_2 \succ e_3$ and clauses $c_1, c_2, c_3$.
-  If the clauses cover examples according to table~\ref{tbl:cover_examples}, then, for example, the function $(1 \cdot c_1) + (0\cdot c_2) + (-2\cdot c_3)$ perfectly models the given rankings.
-  A new example that is covered by~$c_1$ and~$c_2$, would be assigned a score of $1$, according to this function.
-  This score has no value in the absolute sense, it can only be used to compare it to other examples ranked by the same function.
-  In this example clause~$c_1$ represents a desirable property, clause~$c_2$ is ignored because it does not influence the ranking and clause~$c_3$ represents an undesirable property.
+  Consider examples~$e_1, e_2, e_3$, preferences $e_1 \succ e_2, e_2 \succ e_3$ and clauses $c_1, c_2, c_3$, whereby clauses cover examples according to Table~\ref{tbl:cover_examples}.
+  The function $f(\mathbf{v}) = (1, 0, -2) \cdot \mathbf{v}$ assigns scores $f(\mathbf{v_{e_1}}) = 1$, $f(\mathbf{v_{e_2}}) = 0$ and $f(\mathbf{v_{e_3}}) = -1$ and perfectly models the given preferences since $f(\mathbf{v_{e_1}}) > f(\mathbf{v_{e_2}})$ and $f(\mathbf{v_{e_2}}) > f(\mathbf{v_{e_3}})$.
+  This score has no value in the absolute sense, it can only be used to compare between examples.
+  In this example clause~$c_1$ represents a desirable property and clause~$c_3$ represents an undesirable property.
 
@@
     \toprule
-    $e_1$             & $c_1$               & $\mathbf{f}_{e_1} = (1, 0, 0)$ \\
-    $e_2$             & $c_2$               & $\mathbf{f}_{e_2} = (0, 1, 0)$ \\
-    $e_3$             & $c_1$, $c_2$, $c_3$ & $\mathbf{f}_{e_3} = (1, 1, 1)$ \\
+    $e_1$             & $c_1$               & $\mathbf{v}_{e_1} = (1, 0, 0)$ \\
+    $e_2$             & $c_2$               & $\mathbf{v}_{e_2} = (0, 1, 0)$ \\
+    $e_3$             & $c_1$, $c_2$, $c_3$ & $\mathbf{v}_{e_3} = (1, 1, 1)$ \\
   \end{tabularx}
@@
 \svm{} \cite{joachims2006training} was chosen to find the scoring function since it uses a linear model and offers a robust and efficient implementation.
-The weights that it assigns to the features can be used directly as weights for the optimization criteria.
-While this paper uses \svm{} other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.
+The weights that it assigns to the features can be used directly as weights for the clauses.
+While the current implementation uses \svm{}, other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.
 
 \subsection{Optimal Solution}
-Solvers, such as IDP \sam{more on IDP, can be used as ASP solver?}, can use clauses directly as hard constraints to generate a solution.
-There is no solver that can use the chosen optimization criteria directly.
-Weighted MAX-SAT solvers use propositional clauses and only allow for positive weights.
-
-This optimization task can be solved in IDP, using inductive definitions, aggregates and minimization.
-The only limitation is that the current version only supports integer values.
-Therefore, the weights of the clauses are divided by the smallest absolute weight, multiplied by a constant and rounded to the closest integer if necessary.
+Typical MAX-SAT solvers use propositional clauses and only allow for positive weights.
+However, the first-order logic solver IDP can solve our optimization task as it supports inductive definitions, aggregates and minimization.
+The only limitation is that the current version is restricted to integer values and does not support rational numbers.
+Therefore, the weights of the clauses are divided by the smallest absolute weight, multiplied by a constant and rounded to the closest integer.
 
@@
 In our experiments we aim to measure the accuracy and efficiency of the learning systems for constraints and optimization criteria.
-In order to account for non-determinism, experiments measuring execution times or optimization scores were repeated eight times and their result were averaged.
-\sam{First step evaluation constraint learning, then optimization}
-\subsection{Constraints}
-We used four problems to evaluate constraint learning: \sam{clarify hard constraints}
+Non-deterministic experiments (e.g. measuring execution times or accuracy of clausal optimization) were repeated eight times and their result were averaged.
+Some problems and experiments are only described briefly or summarized, for more details we refer to \cite{kolb2015thesis}\footnote{Available online at https://github.com/samuelkolb/thesis}.
+
+First, experiments concerning the accuracy and efficiency of the constraint learning implementation are discussed.
+Thereafter, we discuss the evaluation of the accuracy of the clausal optimization system\footnote{The system (v1.0) used to perform all experiments is available online at https://github.com/samuelkolb/clausal-discovery/releases}.
+
+\subsection{Clause Learning}
+We used four problems to evaluate constraint learning:
 \begin{description}
@@
 Two examples, each containing three correctly colored countries, are given.
-\item [Sudoku] A single, solved $4 \times 4$ sudoku is given.
-\item[Elevator] Three examples have been generated, two of which respect the underlying soft constraint.
-\item[Co-housing] Contains four hard constraints and five examples have been generated that respect the constraints.
+\item[Sudoku] A single, solved $4 \times 4$ sudoku is given as an example for this problem.
+\item[Elevator] This problem uses three predicates $\mathit{inside}$, $\mathit{crowded}$ and $\mathit{panic}$, with $\mathit{crowded}$ being a calculated predicate.
+It is defined by background knowledge to be true for all elevators with three or more people inside.
+Out of the three examples provided for this problem, two satisfy the target soft constraint:
+\begin{flalign*}
+  &\mathit{panic}(p)~\leftarrow~\mathit{crowded}(e)~\land~\mathit{inside(p, e)}.
+\end{flalign*}
+\item[Co-housing] The co-housing problem also uses background knowledge to specify one of its predicates ($\mathit{cheap}$) and to impose some additional constraints.
+Five examples are provided, which all adhere to the following four constraints on the housing choices of two friends:
+\begin{enumerate}
+  \item If the two friends do not live in the same area, they work in the same area.
+  \item If a person does not work and live in the same area, that person needs a car.
+  \item A person who lives in a cheap area has a car.
+  \item If the two friends live together, they live in an expensive area.
+\end{enumerate}
 \end{description}
-\anton{more elaborate descriptions of these?}
-\sam{link to thesis}
+The clausal discovery implementation is used to learn soft constraints for the elevator problem ($t = 1$) and hard constraints for the other problems.
 
@@
 
-Table~\ref{tbl:uitvoering} shows the execution times for the different problems.
-It shows that smaller problems can be solved efficiently.
+Table~\ref{tbl:uitvoering} shows the execution times for the different problems\footnote{All experiments were run on a MacBook Pro with an i7 quad-core processor}.
+It shows that the evaluated problems can be solved efficiently.
 
@@
 These theories usually focus on being compact and contain only the essential constraints.
-Table~\ref{tbl:mens} shows the results of two experiments that measure the time to find a solution for a new problem.
+Table~\ref{tbl:mens} shows the results of two experiments that measure the time to find a solution (model) for a problem.
 This time is measured for the learned theories as well as for hand made theories.
@@
 
-\subsection{Optimization} \sam{Preferences, not rankings}
-\sam{Score, not accuracy}
-\sam{Change figure}
+\subsection{Clausal Optimization}
 The efficiency of the learning system for optimization criteria depends mainly on the efficiency of learning soft constraints.
@@
 
-The moving problem with input according to example~\ref{ex:moving-input} is used to evaluate the learning of optimization criteria.
-18 possible configurations are used as available examples~(\sym{E}).
-Four weighted soft constraints are used to represent the underlying model~(\sym{M}) for the users preferences.
-Two approaches are used for evaluation and differ in the way they construct the training ($\sym{E}_{train}$) and test sets ($\sym{E}_{test}$).
-In the first approach, given a parameter~$p_{train}$, the examples are randomly partitioned into disjoint train and test sets ($\sym{E}_{test}$), such that $\#\sym{E}_{train} = p_{train} \cdot \#\sym{E}$.
+Example~\ref{ex:moving} introduced the moving city scenario which is used to evaluate the learning of optimization criteria.
+18 possible configurations are used as available examples~(\sym{E}), differing in their values for $\mathit{live\_in}$, $\mathit{work\_in}$ and $\mathit{school\_in}$.
+The four weighted soft constraints are used to represent the underlying model~(\sym{M}) for the users preferences.
+Two approaches are used for evaluation and vary in the way they construct the training ($\sym{E}_{train}$) and test sets ($\sym{E}_{test}$).
+In the first approach, given a parameter~$p_{train}$, the examples are randomly partitioned into disjoint training and test sets ($\sym{E}_{test}$), such that $\abs{\sym{E}_{train}} = p_{train} \cdot \abs{\sym{E}}$.
 The second approach uses $\sym{E}_{train} = \sym{E}_{test} = \sym{E}$.
 
-Preferences are generated by picking two examples $e_1$,~$e_2 \in \sym{E}_{train}$ and ordering them according to their score: $score(e_i, \sym{M})$.
-Consider all possible pairwise preferences over $\sym{E}_{train}$: $\mathit{pref}(\sym{E}_{train})$.
-If the fraction of preferences to be used is~$p_{pref}$, then a random subset~$\sym{P}_{train} \subset \mathit{pref}(\sym{E}_{train})$ of size $p_{pref} \cdot \# \sym{P}$ is used for learning.
+The available input preferences are obtained by generating all preferences~$\sym{P}$ over $\sym{E}_{train}$ using $score_{\sym{M}}$.
+If the fraction of preferences to be used is~$p_{pref}$, then a random subset~$\sym{P}_{train} \subset \sym{P}$ of size $p_{pref} \cdot \abs{\sym{P}}$ is used for learning.
 Errors in the input preferences are simulated by flipping preferences (e.g. $e_1 \succ e_2$ becomes $e_1 \prec e_2$).
-Given the fraction $p_{error}$ of errors to introduce, a random subset $\sym{P}_{error} \cup \sym{P}_{train}$ of the preferences are flipped, where $\#\sym{P}_{error} = p_{error} \cdot \sym{P}_{train}$.
+Given the fraction $p_{error}$ of errors to introduce, a random subset $\sym{P}_{error} \subset \sym{P}_{train}$ of the preferences are flipped, where $\abs{\sym{P}_{error}} = p_{error} \cdot \abs{\sym{P}_{train}}$.
 
-To evaluate the learned optimization criteria, all possible pairs of examples in the test set are generated.
-Both the learned model and the underlying model~$\sym{M}$ are used to predict the better example for every pair.
-The accuracy of the learned model is then calculated as the fraction of correctly predicted pairs. 
-Pairs for which~\sym{M} ranks both examples the same are omitted.
+To evaluate a learned model~$\sym{L}$, all possible preferences over $\sym{E}_{test}$ are generated using both $score_{\sym{M}}$ and $score_{\sym{L}}$, yielding $\sym{P}_\sym{M}$ and $\sym{P}_\sym{L}$, respectively.
+The accuracy of the learned model is then calculated as the fraction of correctly predicted preferences $(\frac{\abs{\sym{P}_\sym{M} \cap \sym{P}_\sym{L}}}{\abs{\sym{P}_\sym{M}}})$.
 
@@
 
-The influence of errors is shown in figure~\ref{fig:ruis}.
-Hereby, the second approach of testing is applied using overlapping train- and test sets.
-High scores are obtained, even despite significant levels of noise.
+The influence of errors is shown in Figure~\ref{fig:ruis}.
+For this experiment, the second approach of testing has been applied, using overlapping training and test sets ($\sym{E}_{train} = \sym{E}_{test} = \sym{E}$).
+High accuracy values are obtained, despite significant numbers of errors.
 The figure also shows that increasing the number of preferences used for learning ($p_{pref}$) improves the robustness of the algorithm, even if the relative amount of errors ($p_{error}$) remains unchanged.
@@
 We presented an approach to automatically acquire constraints and optimization criteria from examples and preferences.
-Implementation for both clause learning and clausal optimization have been provided that accomplish these tasks using first-order logic clauses.
+Implementations for both clause discovery and clausal optimization have been provided that accomplish these tasks using first-order logic clauses.
 
 The constraint learning implementation has been able to learn the relevant hard and soft constraints in all experiments.
-For each problem, only a small number of examples was given to learn from.
-The system requires only a minimal amount of information from the user, however, it also allows for the use of expressive background knowledge.
-Constraints that are learned from examples do not depend on the domains used in those examples.
-This 
-The learned constraints are independent of the domains used in the examples, which facilitates the construction of positive examples.
+For each problem, the relevant constraints could be learned from only a few examples.
+Although the system requires only examples and preferences, users can provide more information through background knowledge.
+The constraints learned by clausal discovery are independent of the constants of specific problem instantiations (examples), which facilitates the construction of positive examples.
 
-Using the constraint learning implementation, the goal of learning optimization criteria has been accomplished.
-Optimization criteria can be learned that enable optimal solutions to be found.
-Even for small datasets and noisy rankings, constraints are found that rank most examples correctly.
+Using the constraint learning implementation, the goal of learning a weighted first-order MAX-SAT model has been accomplished.
+The learned models allow optimal solutions to be found.
+Even for small datasets and erroneous preferences, optimization criteria are found that rank most examples correctly.
 
@@
 \section*{Acknowledgments}
-The author thanks his promoters Luc De Raedt en Anton Dries.
+The author thanks his promoters Luc De Raedt and Anton Dries.
 Samuel Kolb is supported by the Research Foundation - Flanders (FWO).
-
-\newpage
 

### 2015-10-31 Paper.tex
Predicate casing

@@
 We use (function free) first-order clauses; which we now introduce.
-An atom $p(t_1, t_2, ..., t_n)$ consists of a predicate symbol~$p$ that is applied to $n$ terms~$t_i$. \sam{Casing}
+An atom $P(t_1, t_2, ..., t_n)$ consists of a predicate symbol~$P$ that is applied to $n$ terms~$t_i$.
 Terms are either constants or variables.
-In this paper constants will be written in uppercase ($\mathit{John}$) and variables in lowercase ($\mathit{x}$).
+In this paper predicates and constants will be written in uppercase ($\mathit{John}$) and variables in lowercase ($\mathit{x}$).
 A literal can be an atom~$a$ (positive literal) or a negated atom~$\lnot a$ (negative literal).
@@
   Consider the scenario of moving to a new city and choosing areas for housing, work and school.
-  The choices are captured by the predicates $\mathit{live\_in}$, $\mathit{work\_in}$ and $\mathit{school\_in}$.
-  Additionally, the predicates $\mathit{cheap}$ and $\mathit{low\_crime}$ are used to designate cheap and low crime areas, respectively.
+  The choices are captured by the predicates $\mathit{LiveIn}$, $\mathit{WorkIn}$ and $\mathit{SchoolIn}$.
+  Additionally, the predicates $\mathit{Cheap}$ and $\mathit{LowCrime}$ are used to designate cheap and low crime areas, respectively.
 
@@
   \begin{flalign*}
-    (w_1, c_1) &= (0.50, \mathit{low\_crime}(a) \leftarrow \mathit{live\_in}(a)) \\
-    (w_2, c_2) &= (0.25, \mathit{school\_in}(a) \leftarrow \mathit{work\_in}(a)) \\
-    (w_3, c_3) &= (1.00, \mathit{low\_crime}(a) \leftarrow \mathit{school\_in}(a)) \\
-    (w_4, c_4) &= (-1.00, \mathit{false} \leftarrow \mathit{live\_in}(a) \land \mathit{cheap}(a))
+    (w_1, c_1) &= (0.50, \mathit{LowCrime}(a) \leftarrow \mathit{LiveIn}(a)) \\
+    (w_2, c_2) &= (0.25, \mathit{SchoolIn}(a) \leftarrow \mathit{WorkIn}(a)) \\
+    (w_3, c_3) &= (1.00, \mathit{LowCrime}(a) \leftarrow \mathit{SchoolIn}(a)) \\
+    (w_4, c_4) &= (-1.00, \mathit{false} \leftarrow \mathit{LiveIn}(a) \land \mathit{Cheap}(a))
   \end{flalign*}
   Let examples $e_1$, $e_2$ and~$e_3$ be Herbrand interpretations that reason over the same city, using the same constants (areas) $A_1$, $A_2$ and $A_3$.
-  Consider the set~\sym{S} of shared ground atoms $\{\mathit{cheap}(A_1)$, $\mathit{cheap}(A_3)$, $\mathit{low\_crime}(A_1)$, $\mathit{low\_crime}(A_2)\}$ and define the examples as:
+  Consider the set~\sym{S} of shared ground atoms $\{\mathit{Cheap}(A_1)$, $\mathit{Cheap}(A_3)$, $\mathit{LowCrime}(A_1)$, $\mathit{LowCrime}(A_2)\}$ and define the examples as:
   \begin{flalign*}
-    &e_1 = \{\mathit{live\_in}(A_1), \mathit{work\_in}(A_2), \mathit{school\_in(A_2)}\} \cup \sym{S} \\
-    &e_2 = \{\mathit{live\_in}(A_2), \mathit{work\_in}(A_2), \mathit{school\_in(A_3)}\} \cup \sym{S}\\
-    &e_3 = \{\mathit{live\_in}(A_2), \mathit{work\_in}(A_2), \mathit{school\_in(A_1)}\} \cup \sym{S}.
+    &e_1 = \{\mathit{LiveIn}(A_1), \mathit{WorkIn}(A_2), \mathit{SchoolIn(A_2)}\} \cup \sym{S} \\
+    &e_2 = \{\mathit{LiveIn}(A_2), \mathit{WorkIn}(A_2), \mathit{SchoolIn(A_3)}\} \cup \sym{S}\\
+    &e_3 = \{\mathit{LiveIn}(A_2), \mathit{WorkIn}(A_2), \mathit{SchoolIn(A_1)}\} \cup \sym{S}.
   \end{flalign*}
@@
 In order to model the optimization problem in IDP, every clause $c_i$ with variables $v_1, ..., v_n$ is represented by a number $i$.
-For every clause a predicate $t(i)$ is added to capture the truth value of the clause.
-A function $\mathit{cost}(i)$ specifies the cost of not satisfying the clause, which is equal to the weight of the clause.
+For every clause a predicate $T(i)$ is added to capture the truth value of the clause.
+A function $\mathit{Cost}(i)$ specifies the cost of not satisfying the clause, which is equal to the weight of the clause.
 \begin{flalign*}
-  &t(i) \Leftrightarrow \forall v_1, ..., v_n : c_i. \\
-  &cost(i) = w_i.
+  &T(i) \Leftrightarrow \forall v_1, ..., v_n : c_i. \\
+  &Cost(i) = w_i.
 \end{flalign*}
 
-Using $t$ and $\mathit{cost}$, a function $\mathit{actual}(i) = \mathbf{1}_{\lnot t(i)} \cdot \mathit{cost}(i)$ is then defined inductively.
-This function is used in the optimization criterion $\sum_i actual(i)$ which should be minimized. % , which will allow IDP to search for an optimal solution.
+Using $T$ and $\mathit{Tost}$, a function $\mathit{Actual}(i) = \mathbf{1}_{\lnot T(i)} \cdot \mathit{Cost}(i)$ is then defined inductively.
+This function is used in the optimization criterion $\sum_i Actual(i)$ which should be minimized. % , which will allow IDP to search for an optimal solution.
 
@@
 \item[Sudoku] A single, solved $4 \times 4$ sudoku is given as an example for this problem.
-\item[Elevator] This problem uses three predicates $\mathit{inside}$, $\mathit{crowded}$ and $\mathit{panic}$, with $\mathit{crowded}$ being a calculated predicate.
+\item[Elevator] This problem uses three predicates $\mathit{Inside}$, $\mathit{Crowded}$ and $\mathit{Panic}$, with $\mathit{Crowded}$ being a calculated predicate.
 It is defined by background knowledge to be true for all elevators with three or more people inside.
@@
 \begin{flalign*}
-  &\mathit{panic}(p)~\leftarrow~\mathit{crowded}(e)~\land~\mathit{inside(p, e)}.
+  &\mathit{Panic}(p)~\leftarrow~\mathit{Crowded}(e)~\land~\mathit{Inside(p, e)}.
 \end{flalign*}
-\item[Co-housing] The co-housing problem also uses background knowledge to specify one of its predicates ($\mathit{cheap}$) and to impose some additional constraints.
+\item[Co-housing] The co-housing problem also uses background knowledge to specify one of its predicates ($\mathit{Cheap}$) and to impose some additional constraints.
 Five examples are provided, which all adhere to the following four constraints on the housing choices of two friends:
@@
 Example~\ref{ex:moving} introduced the moving city scenario which is used to evaluate the learning of optimization criteria.
-18 possible configurations are used as available examples~(\sym{E}), differing in their values for $\mathit{live\_in}$, $\mathit{work\_in}$ and $\mathit{school\_in}$.
+18 possible configurations are used as available examples~(\sym{E}), differing in their values for $\mathit{LiveIn}$, $\mathit{WorkIn}$ and $\mathit{SchoolIn}$.
 The four weighted soft constraints are used to represent the underlying model~(\sym{M}) for the users preferences.
@@
      0.823 & 0.740 & 0.788 & 0.735 \\
-     ($\pm$ 0.073)&
-($\pm$ 0.078)&
-($\pm$ 0.074)&
-($\pm$ 0.063)
+     ($\pm$ 0.073) & ($\pm$ 0.078) & ($\pm$ 0.074) & ($\pm$ 0.063)
     \end{tabularx}

### 2015-10-31 Kolb/Paper.tex
Final version

--- 2015-10-31 Paper.tex
@@
 % -------------------------------------------------
-
 \section{Introduction}
-\sam{it might be good to state somewhere early that you are in the "learning from interpretations" setting, but additionally have preferences}
-\sam{Structure}
 An important class of declarative models is concerned with (discrete) constrained optimization problems.
@@
 Furthermore, \cite{campigotto2011active} have learned propositional weighted MAX-SAT theories from examples.
-Also, work in statistical relational learning (SRL) on learning the structure of Markov Logic networks (MLN) or other SRL models \cite{kok2005learning} is relevant, as MAP inference can be used to query for the most likely state.
 
@@
 Furthermore, we learn such models by combining inductive logic programming (ILP) principles (such as clausal discovery \cite{DeRaedt:ClausalDiscovery}) with preference learning.
-We use ILP to learn the clausal theories, and preference learning to determine the weights.
+We use ILP to learn the clausal theories from interpretations, and preference learning to determine the weights.
 Therefore, users provide examples of possible worlds and a set of preferences that specify which examples are preferred over other ones.
@@
 It also extends the approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than examples with target values.
-Moreover, it differs from the SRL and MLN approaches in that one learns from preferences rather than from examples.
+
+Weighted first-order MAX-SAT problems relate to Markov Logic networks (MLN), as MAP inference in MLNs can be used to query for the most likely state.
+Therefore, work in statistical relational learning (SRL) on learning the structure of MLNs or other SRL models \cite{kok2005learning} is also relevant.
+However, our work differs from the SRL and MLN approaches in that one learns from preferences rather than from examples.
 
@@
 
-\sam{versus MLN, PSL}
 In related work \cite{campigotto2011active} weighted MAX-SAT problems are induced by learning propositional clauses and weights using examples and absolute scores.
@@
 \paragraph{Constraint Learning}
-\sam{Define constraints}
 The goal of constraint learning is to automatically identify constraints in one or multiple examples.
@@
 
-\begin{example}[Moving]
+\begin{example}[Moving city]
   \label{ex:moving}
@@
 These constraints are too specific and will exclude valid solutions that are not in the training set.
-On the other hand, if the chosen limits are too small, the necessary constraints will not be found.
+However, if the chosen limits are too small, the necessary constraints will not be found.
 
@@
 
+\newpage
 %
@@
 %
+\bibliography{Bibliography}
 \bibliographystyle{aaai}
-\bibliography{Bibliography}
 \end{document}

### 2015-10-31 Paper.tex
Final version

@@
 \section{Introduction}
-\sam{it might be good to state somewhere early that you are in the "learning from interpretations" setting, but additionally have preferences}
-\sam{Structure}
 An important class of declarative models is concerned with (discrete) constrained optimization problems.

### 2015-12-14 Paper.tex
Working to incorporate changes

@@
 
+\begin{itemize}
+  \item \sam{Why interesting, connection to Campignotto, Domingos, Kok, Why preferences useful?}
+  \item \sam{Real tasks with preferences? Small scale evaluation}
+  \item \sam{Lot of background knowledge assumed (vs. constraint programming, probabilistic programming, demo model and need for learning)}
+\end{itemize}
+
 \section{Introduction}
@@
 
-Learning constraints has been addressed in constraint programming (e.g. \cite{Beldiceanu:ModelSeeker,bessiere2013constraint}) and inductive logic programming (e.g. \cite{DeRaedt:ClausalDiscovery}).
+Constraint learning is the task of discovering constraints from data.
+This task has been addressed in constraint programming (e.g. \cite{Beldiceanu:ModelSeeker,bessiere2013constraint}) and inductive logic programming (e.g. \cite{DeRaedt:ClausalDiscovery}).
 Recent efforts \cite{Lallouet:LearningCP} have attempted to combine these two fields.
@@
 
-We use weighted first-order logical theories to represent constraint optimization problems.
+We use weighted first-order logical theories to represent constrained optimization problems.
 The weights capture the optimization function, and (first-order) weighted MAX-SAT can be used to determine the \emph{best} solutions.
@@
 This setting extends traditional constraint learning and ILP approaches with the ability to learn soft constraints and integrates constraint satisfaction and optimization.
-It also extends the approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than examples with target values.
+It also extends the propositional approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than examples with target values.
+Preferences are useful in situation where the score of an example is hard to quantify and in interactive settings because it is often easier for humans to provide preferences over few examples at a time than absolute scores.
 
@@
 However, our work differs from the SRL and MLN approaches in that one learns from preferences rather than from examples.
+
+% We use preferences for learning as it is often easier for humans to provide preferences over few examples at a time than absolute scores.
+% This makes them useful in interactive settings and in situations where an examples absolute score is hard to quantify.
+
+% One of the main advantages of using preferences for learning is that preferences are often easier to provide for humans and only consider few examples at a time.
+% Consequently, they are useful in interactive settings and in situations where an examples absolute score is hard to quantify.
 
@@
 There are several opportunities for future work.
+In the current approach weight learning is performed using preferences and a learn-to-rank SVM.
+It would, however, be easy to use another type of input (e.g. absolute scores) and a corresponding SVM.
+
 It would be interesting to adapt the number of variables and literals per clause dynamically, as they have a strong influence on the efficiency and accuracy of the clause learning system.
@@
 The author thanks his promoters Luc De Raedt and Anton Dries.
+Furthermore, he is is thankful for the help provided by Bart Bogaerts, Jesse Davis, Marc Denecker, Vladimir Dzyuba, Angelika Kimmig and Jan Tobias M\"uhlberg.
 Samuel Kolb is supported by the Research Foundation - Flanders (FWO).
@@
 \newpage
-%
-% ---- Bibliography ----
-%
+
+% -------------------------------------------------
+% Bibliography
+% -------------------------------------------------
+
 \bibliography{Bibliography}

### 2015-12-15 Paper.tex
Final content edits

@@
 
-\begin{itemize}
-  \item \sam{Why interesting, connection to Campignotto, Domingos, Kok, Why preferences useful?}
-  \item \sam{Real tasks with preferences? Small scale evaluation}
-  \item \sam{Lot of background knowledge assumed (vs. constraint programming, probabilistic programming, demo model and need for learning)}
-\end{itemize}
+% \begin{itemize}
+%   \item \sam{Why interesting, connection to Campignotto, Domingos, Kok, Why preferences useful?}
+%   \item \sam{Real tasks with preferences? Small scale evaluation}
+%   \item \sam{Lot of background knowledge assumed (vs. constraint programming, probabilistic programming, demo model and need for learning!)}
+% \end{itemize}
 
@@
 
-We use weighted first-order logical theories to represent constrained optimization problems.
+We use weighted first-order logical theories to represent constrained optimization problems (as illustrated in example~1).
 The weights capture the optimization function, and (first-order) weighted MAX-SAT can be used to determine the \emph{best} solutions.
@@
 It also extends the propositional approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than examples with target values.
-Preferences are useful in situation where the score of an example is hard to quantify and in interactive settings because it is often easier for humans to provide preferences over few examples at a time than absolute scores.
+Preferences are useful in situations where the score of an example is hard to quantify (e.g.
+comparing songs or designs) and it is often easier for humans to provide preferences over few examples at a time than absolute scores.
 
@@
 Starting with the empty clause ($\square$), clauses are refined until they cover enough examples.
-If a clause covers enough examples it is added to the result set, provided it is not entailed by other clauses already in the result set.
+If a clause covers enough examples it is added to the result set, unless it is entailed by clauses already in the result set.
 In the refinement step, a clause $c = \{l_1, l_2, ..., l_k\}$ is extended (refined) by adding a new literal $l$.
@@
 The weights that it assigns to the features can be used directly as weights for the clauses.
-While the current implementation uses \svm{}, other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.
+Other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.
 
@@
 These constraints are too specific and will exclude valid solutions that are not in the training set.
-However, if the chosen limits are too small, the necessary constraints will not be found.
-
-Over-fitting can be addressed by removing constraints that exclude valid solutions (manually or automatically) as well as providing more (or larger) training examples.
-Negative examples can be used to detect under-fitting, indicating that the parameter values are too small.
+Over-fitting can be addressed by removing invalid constraints (manually or automatically) and providing more (or larger) training examples.
+On the other hand, if the chosen limits are too small, the necessary constraints will not be found.
+Under-fitting can be detected using negative examples.
 
@@
 Increasing the number of variables or literals per clause negatively impacts the efficiency.
-Especially allowing both more variables \emph{and} more literals can steeply increase the execution time.
-Including additional examples, however, only increases the execution time by a constant factor.
-Since adding more examples can help improve the accuracy, this trade-off is often worthwhile.
+Especially increasing both parameters can steeply increase the execution time.
+The current implementation may still need significant time to solve large problems that include many predicates.
+However, including additional or larger examples usually only increases the execution time by a constant factor.
 
@@
 We presented an approach to automatically acquire constraints and optimization criteria from examples and preferences.
-Implementations for both clause discovery and clausal optimization have been provided that accomplish these tasks using first-order logic clauses.
+Implementations for both clause discovery and clausal optimization were provided that accomplish these tasks using first-order logic clauses.
 
